### 2021-03-29 outputs/exploratory.Rmd
add some of the exploratory work

---
title: "data exploration"
author: "Amy"
date: "29/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

```{r}
isrdata <- read.dta13("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta")
```

```{r}
expdata <- isrdata %>%
  filter(use == 1)
```

```{r}
colSums(is.na(expdata))
```

```{r}
expdata %>%
  select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7) %>%
  unique() %>%
  summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
            black = sum(t1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
            asian = sum(t1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(t1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(t1_race_6, na.rm = TRUE)/n(),
            other = sum(t1_race_7, na.rm = TRUE)/n(),
            total = n())
```

```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
            black = sum(s1_race_2, na.rm = TRUE)/n(),
            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
            asian = sum(s1_race_4, na.rm = TRUE)/n(),
            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
            other = sum(s1_race_7, na.rm = TRUE)/n(),
            total = n())
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarize(meanage = mean(s1_age, na.rm = TRUE))
```
```{r}
expdata %>%
  group_by(treatment) %>%
  summarise(prestudentsimilarity = mean(s1_sim, na.rm = TRUE),
            prestudentISR = mean(s1_tsr),
            poststudentsimilarity = mean(s2_sim, na.rm = TRUE),
            poststudentISR = mean(s2_tsr, na.rm = TRUE),
            postinstructorsimilarity = mean(t2_sim1, na.rm = TRUE),
            postinstructorISR = mean(t2_tsr, na.rm = TRUE),
            standardgrade = mean(grade_sd, na.rm = TRUE))
```

### 2021-03-29 scripts/01_import_data.R

@@ -18,9 +18,6 @@ library(labelled)
 # this import shows the codes for answers, not just numbers
 isrdata <- read.dta13("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta")
 
-# save as a csv
-write_csv(isrdata, "inputs/data/isrdata.csv")
-
 ### Make the codebook ###
 
 # this import shows the variable labels, but the answers are numerical

### 2021-04-09 outputs/exploratory.Rmd
add models

@@
 library(here)
+library(readstata13)
+library(arsenal)
+library(Hmisc)
+library(MASS)
+library(kableExtra)
+library(apaTables)
 ```
@@
 ```{r}
-isrdata <- read.dta13("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta")
+codebook <- read_csv(here("inputs/data/codebook.csv"))
+```
+
+```{r, warning=FALSE}
+isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
 ```
@@
   filter(use == 1)
+```
+
+```{r}
+summary(tableby(treatment ~  s1_sim + s1_tsr + s2_sim + s2_tsr + t2_sim1 + t2_tsr + grade + std_grade + t2_finalexam + f17_enrolled + ir_f16_gpa + s_female,
+                data = isrdata)) %>%
+  kable()
+```
+
+```{r}
+rcorr(as.matrix(isrdata %>% 
+                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa)))
+```
+
+```{r}
+apa.cor.table(isrdata %>% 
+                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa))
 ```
@@
 ```
+
+```{r}
+summary(lm(data = isrdata, s1_sim ~ treatment))
+```
+```{r}
+summary(lm(data = isrdata, s2_sim ~ treatment))
+```
+```{r}
+summary(lm(data = isrdata, s2_tsr ~ treatment + s1_tsr))
+```
+```{r}
+summary(polr(formula = as.factor(t2_sim1) ~ treatment, data = isrdata, Hess = TRUE))
+```
+
+```{r}
+summary(lm(data = isrdata, s2_tsr ~ treatment))
+```
+
+```{r}
+summary(lm(data = isrdata, grade ~ treatment + s1_female + ir_f16_gpa))
+```
+
+```{r}
+summary(lm(data = isrdata %>% filter(obj_exam == 1), t2_finalexam ~ treatment + s1_female + ir_f16_gpa))
+```
+
+```{r}
+summary(glm(data = isrdata, f17_enrolled ~ treatment + s1_female + ir_f16_gpa, family = "binomial"))
+```

### 2021-04-09 outputs/paper/paper.Rmd
add models

---
title: "TODO TITLE"
author: "Amy Farrow"
date: "April 9th, 2021"
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Reproducing ‘Taking It to the Next Level’"
abstract: "TODO ABSTRACT"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table

# NOTE: script 01_ must have been run already

# functions from https://stefaneng.github.io/apa_correlation_table/

apply_if <- function(mat, p, f) {
  # Fill NA with FALSE
  p[is.na(p)] <- FALSE
  mat[p] <- f(mat[p])
  mat
}

apaCorr <- function(mat, corrtype = "pearson") {
  matCorr <- mat
  if (class(matCorr) != "rcorr") {
    matCorr <- rcorr(mat, type = corrtype)
  }

  # Add one star for each p < 0.05, 0.01, 0.001
  stars <- apply_if(round(matCorr$r, 2), matCorr$P < 0.05, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.01, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.001, function(x) paste0(x, "*"))
  # Put - on diagonal and blank on upper diagonal
  stars[upper.tri(stars, diag = T)] <- "-"
  stars[upper.tri(stars, diag = F)] <- ""
  n <- length(stars[1,])
  #colnames(stars) <- 1:n
  # Remove _ and convert to title case
  #row.names(stars) <- tools::toTitleCase(sapply(row.names(stars), gsub, pattern="_", replacement = " "))
  # Add index number to row names
  #row.names(stars) <- paste(paste0(1:n,"."), row.names(stars))
  stars
}
```

# Introduction

TODO intro

# Data

```{r, warning=FALSE, message=FALSE, echo=FALSE}
isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
expdata <- isrdata %>%
  filter(use == 1) %>%
  mutate(across(where(is.numeric), ~na_if(., -99)))
```

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by Robinson et al., including missing values. The variables are as follows:

* s1_sim: Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5. These responses were averaged to create a similarity scale. 
  - Overall, how similar to your instructor's values do you think your values are?
  - How similar are your goals for the course and your instructor's goals?
  - In general, how similar do you think your views about the course content and your instructor's are?
  - How much do you think you have in common with your instructor?
  - How similar do you think your personality is compared to your instructor's?
  - Overall, how similar do you think you and your instructor are?
* s1_tsr: Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create an ISR scale.
  - How much do you think you will enjoy learning from this instructor?
  - How friendly do you think this instructor will be towards you?
  - How encouraging do you think this instructor will be towards you?
  - If you came back to visit this instructor three years from now, how excited do you think they would be?
  - How motivating do you think you will find this instructor's class?
  - How caring do you think this instructor will be towards you?
  - Overall, how much do you think you will learn from this instructor?
* s2_sim: At the end of the semester, students answered the above six similarity scale questions again.
* s2_tsr: At the end of the semester, students answered the above seven ISR scale questions again.
* t2_sim1: At the end of the semester, instructors answered only one quesiton about similarity with the student, on a scale of 1 to 5.
  - Overall, how similar do you think you and STUDENTNAME are?
* t2_tsr: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create an ISR scale.
  - How much did you enjoy helping STUDENTNAME learn?
  - How caring was STUDENTNAME towards you?
  - How often did you say something encouraging to STUDENTNAME?
  - How friendly was STUDENTNAME towards you?
  - If this student came back to visit you three years from now, how excited would you be?
  - How motivating did STUDENTNAME find the activities that you plan for class?
  - Overall, how much did STUDENTNAME learn from you?
* grade: The final grade that the student received in the course. These were not reported by the student.
* std_grade: The student's final grade, standardized against other grades in the course.
* t2_finalexam: Instructors were asked to report the student's grade on their final exam, paper, or project.
* ir_f16_gpa: The student's cumulative GPA after the Fall 2016 term.
* f17_enrolled: The student's status as of Fall term 2017: not enrolled or enrolled.
* s1_female: The student's gender.

```{r summarytable, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>% 
  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, s1_female) %>%
  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "ir_f16_gpa", "f17_enrolled", "s1_female"),
                     p = FALSE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N") %>%
  knitr::kable(caption = "Descriptive statistics for treatment and control groups",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

# Models

## Reproduction models

$treatment_{i}$ is the indicator that treatment was given.

$X_{1i}$ is a vector of student-level covariates (pre-intervention measures included).

$X_{2j}$ is a vector of instructor-level covariates.

$\epsilon_{ij}$ is a clustered residual.

$\beta_{0}$, $beta_{1}$, $\Gamma_{1}$, $\Gamma_{2}$, and $a_{k}$ are coefficients on the resulting models.

### Linear models

Equation \@ref(eq:linmodel) is Robinson et al.'s linear model, used for continuous outcomes, which include immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), end of semester instructor ISR rating (t2_tsr), course grade (grade), and objectively graded exam grade (t2_finalexam).

\begin{equation}
(\#eq:linmodel)
Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij}
\end{equation}

### Ordinal logistic models

Equation \@ref(eq:ordlogmodel) is Robinson et al.'s ordinal logistic model, used for ordinal outcomes, which include end of semester instructor similarity rating (t2_sim1).

\begin{equation}
(\#eq:ordlogmodel)
prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij} > k
\end{equation}

### Logistic models

Equation \@ref(eq:logmodel) is Robinson et al.'s logistic model, used for binary outcomes, which include enrollment in Fall term 2017 (f17_enrolled).

\begin{equation}
(\#eq:logmodel)
prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij} > k
\end{equation}

## Additional models

# Results

## Reproduction results

Table \@ref(tab:correlationtable)

```{r correlationtable, echo=FALSE, warning=FALSE, message=FALSE}
summary <- expdata %>% 
  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, s1_female) %>%
  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "ir_f16_gpa"),
                     p = FALSE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = FALSE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N (%) = ") %>%
  select("Total N", "Missing N", "Total")

matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa)))) 
summary %>%
  bind_cols(matrix) %>%
  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N") %>%
  kable(caption = "Descriptive statistics and correlation matrix for key continuous variables",
               booktabs = TRUE, linesep = "") %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

### Linear models

Table \@ref(tab:studentfirstsimilarity)

```{r studentfirstsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(s1_sim)) %>%
  finalfit("s1_sim", "treatment") %>%
  knitr::kable(caption = "Effect of treatment on immediate student similarity rating",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

Table \@ref(tab:studentsecondsimilarity)

```{r studentsecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(s2_sim)) %>%
  finalfit("s2_sim", "treatment") %>%
  knitr::kable(caption = "Effect of treatment on end of semester student similarity rating",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

Table \@ref(tab:studentsecondrelationship)

```{r studentsecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(s2_tsr) & !is.na(s1_tsr)) %>%
  finalfit("s2_tsr", c("treatment", "s1_tsr")) %>%
  knitr::kable(caption = "Effect of treatment on immediate student ISR rating, controlling for anticipated ISR",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

Table \@ref(tab:teachersecondrelationship)

```{r teachersecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(t2_tsr)) %>%
  finalfit("t2_tsr", c("treatment")) %>%
  knitr::kable(caption = "Effect of treatment on end of semester instructor ISR rating",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

Table \@ref(tab:grade)

```{r grade, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99) %>%
  finalfit("grade", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
  knitr::kable(caption = "Effect of treatment on course grade, controlling for student gender and prior GPA",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

Table \@ref(tab:examgrade)

```{r examgrade, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(obj_exam == 1) %>%
  filter(!is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99) %>%
  finalfit("t2_finalexam", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
  knitr::kable(caption = "Effect of treatment on objectively graded exam grade, controlling for student gender and prior GPA",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

### Ordinal logistic models

Table \@ref(tab:teachersecondsimilarity)

```{r teachersecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  filter(!is.na(t2_sim1)) %>%
  mutate(t2_sim1 = as.factor(t2_sim1)) %>%
  finalfit("t2_sim1", "treatment") %>%
  knitr::kable(caption = "Effect of treatment on end of semester instructor similarity rating",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

### Logistic models

Table \@ref(tab:enrollmentfall)

```{r enrollmentfall, echo=FALSE, warning=FALSE, message=FALSE}
expdata %>%
  mutate(f17_enrolled = as.factor(f17_enrolled)) %>%
  filter(!is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99) %>%
  finalfit("f17_enrolled", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
  knitr::kable(caption = "Effect of treatment on enrollment in next Fall 2017, controlling for student gender and prior GPA",
               booktabs = TRUE, linesep = "") %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
```

## Additional results

# Discussion

### 2021-04-10 outputs/exploratory.Rmd
draft for initial submission

@@
 ```
+
+```{r studentfirstsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(s1_sim)) %>%
+  finalfit("s1_sim", "treatment") %>%
+  knitr::kable(caption = "Effect of treatment on immediate student similarity rating",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r studentsecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(s2_sim)) %>%
+  finalfit("s2_sim", "treatment") %>%
+  knitr::kable(caption = "Effect of treatment on end of semester student similarity rating",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r studentsecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(s2_tsr) & !is.na(s1_tsr)) %>%
+  finalfit("s2_tsr", c("treatment", "s1_tsr")) %>%
+  knitr::kable(caption = "Effect of treatment on end of semester student ISR rating, controlling for anticipated ISR",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r teachersecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(t2_tsr)) %>%
+  finalfit("t2_tsr", c("treatment")) %>%
+  knitr::kable(caption = "Effect of treatment on end of semester instructor ISR rating",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r grade, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99) %>%
+  dplyr::select(s1_female, ir_f16_gpa, grade, s1_female, treatment) %>%
+  finalfit("grade", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
+  knitr::kable(caption = "Effect of treatment on course grade, controlling for student gender and prior GPA",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r examgrade, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(obj_exam == 1) %>%
+  filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99) %>%
+  finalfit("t2_finalexam", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
+  knitr::kable(caption = "Effect of treatment on objectively graded exam grade, controlling for student gender and prior GPA",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r teachersecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  filter(!is.na(t2_sim1)) %>%
+  mutate(t2_sim1 = as.factor(t2_sim1)) %>%
+  finalfit("t2_sim1", "treatment") %>%
+  knitr::kable(caption = "Effect of treatment on end of semester instructor similarity rating",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r enrollmentfall, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>%
+  mutate(f17_enrolled = as.factor(f17_enrolled)) %>%
+  filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99) %>%
+  finalfit("f17_enrolled", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
+  knitr::kable(caption = "Effect of treatment on enrollment in next Fall 2017, controlling for student gender and prior GPA",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r}
+expdata %>% filter(!is.na(s1_sim)) %>%
+  ggplot(aes(x = s1_sim, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(!is.na(s2_sim)) %>%
+  ggplot(aes(x = s2_sim, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(!is.na(s2_tsr) & !is.na(s1_tsr)) %>%
+  ggplot(aes(x = s2_tsr, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(!is.na(t2_tsr)) %>%
+  ggplot(aes(x = t2_tsr, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99) %>%
+  ggplot(aes(x = grade, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(obj_exam == 1) %>% filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99) %>%
+  ggplot(aes(x = t2_finalexam, fill = treatment)) +
+  geom_histogram(position = "dodge")
+```
+
+```{r}
+expdata %>% filter(!is.na(t2_sim1)) %>%
+  ggplot(aes(x = t2_sim1, fill = treatment)) +
+  geom_bar(position = "dodge")
+```
+
+```{r}
+expdata %>% 
+  mutate(f17_enrolled = as.factor(f17_enrolled)) %>% 
+  filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99) %>%
+  ggplot(aes(x = f17_enrolled, fill = treatment)) +
+  geom_bar(position = "dodge")
+```

### 2021-04-10 outputs/paper/paper.Rmd
draft for initial submission

@@
 ---
-title: "TODO TITLE"
+title: "The inefficacy of superficial similarities for establishing instructor-student relationships"
 author: "Amy Farrow"
 date: "April 9th, 2021"
+header-includes:
+  \usepackage{dcolumn}
+  \usepackage{rotating}
 output:
@@
 subtitle: "Reproducing ‘Taking It to the Next Level’"
-abstract: "TODO ABSTRACT"
+abstract: "This paper reproduces the 2019 paper 'Taking It to the Next Level', which evaluates an intervention to improve instructor-student relationship.
+The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment.
+While the scalability and affordability of the intervention are desirable, there are no results in any of the targeted measures: those that affect and reflect college retention.
+These results are consistent with the original paper.
+\\par \\textbf{Keywords:} instructor-student relationship, college, reproduction study"
 thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
@@
 library(finalfit) # for summary table
+library(stargazer) # for model table
+library(MASS) # for ordinal logistic model
 
@@
 
-TODO intro
+College is perceived to be a meritocratic tool for social mobility and career success [@college].
+Unfortunately, retention is a large problem in contemporary American colleges, and many students begin degrees without completing them; six-year completion rates for full-time first-time students range from 51% to 86%, depending on school [@finishline].
+Even when controlling for pre-college test scores and initial enrollments, completion rate disparities exist based on parental education, socio-economic status, and race/ethnicity [@finishline].
+Disparities in college completion lead to further entrenchment of long-standing inequalities [@finishline].
+Thus, measures to help students persist in college completion are desirable to reduce wasted resources, and to increase societal equity.
 
-# Data
+Many interventions to increase completion rates have been suggested and tested.
+For example, @citecasemanagement studied case managment techniques; @citerecruiting investigated the effect of high school recruitment; @citenearbies identify the importance of interpersonal relationships; @citecertificate argue that certificates awarded prior to degrees can ameliorate the 'college-completion crisis'.
+These are only a small sample of the interventions suggested.
+Implementing an effective program is difficult due to complicated causes of attrition, embedded social inequalities, and expenses.
 
-```{r, warning=FALSE, message=FALSE, echo=FALSE}
-isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
-```
+@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests one intervention.
+In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships and measures of student success.
+Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
 
-```{r, warning=FALSE, message=FALSE, echo=FALSE}
-expdata <- isrdata %>%
-  filter(use == 1) %>%
-  mutate(across(where(is.numeric), ~na_if(., -99)))
-```
+This paper reproduces @citeoriginal's original study, using anonymized data provided by the authors.
+Linear, logistic, and ordinal logistic models are used to predict outcomes including perceived ISR, grade, and re-enrollment.
+These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grade, exam grade, or future enrollment.
+In addition to the models in the original paper, this paper explores models using different controls and models that consider specific subgroups of the study participants.
 
-Table \@ref(tab:summarytable) shows summary statistics for key variables identified by Robinson et al., including missing values. The variables are as follows:
+# Methods
+
+@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on perceived similarity, instructor-student relationship, course grade, and re-enrollment.
+
+## Participants
+
+The study took place in the 2017 spring semester at a large Californian University.
+The study included 120 instructors and their 2,749 students.
+The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized.Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors.
+
+## Treatment and control
+
+Participating students were randomly assigned to either treatment or control.
+At the beginning of the term, all participating students and instructors were given "get to know you" surveys.
+Using those responses, for each student in the treatment group, seven commonalities were identified between student and instructor (for example, perhaps both student and instructor binge-watch TV to relieve stress, or appreciate loyalty as the most important friend quality), and both student and instructor were informed of these commonalities.
+They completed a few questions about the similarities and were reminded of them through the semester to ensure they were internalized.
+Students in the control group were informed about similarities they shared with students in another part of the country, and instructors were told nothing about these students.
+
+## Procedures
+
+All students participated in a survey immediately following the treatment or the placebo. They were surveyed again at the end of the course. Instructors were surveyed only at the end of the course.
+
+## Measures
+
+@citeoriginal identify key measures. Some are extracted from the student survey at the beginning of the term:
 
@@
   - Overall, how much do you think you will learn from this instructor?
+* s1_female: The student's gender.
+
+Others are extracted from the student survey at the end of the term:
+
 * s2_sim: At the end of the semester, students answered the above six similarity scale questions again.
 * s2_tsr: At the end of the semester, students answered the above seven ISR scale questions again.
+
+Others come from the instructor survey at the end of the term:
+
 * t2_sim1: At the end of the semester, instructors answered only one quesiton about similarity with the student, on a scale of 1 to 5.
@@
   - Overall, how much did STUDENTNAME learn from you?
-* grade: The final grade that the student received in the course. These were not reported by the student.
+* t2_finalexam: Instructors were asked to report the student's grade on their final exam, paper, or project.
+
+Finally, some are extracted from the university's internal records:
+
+* grade: The final grade that the student received in the course.
 * std_grade: The student's final grade, standardized against other grades in the course.
-* t2_finalexam: Instructors were asked to report the student's grade on their final exam, paper, or project.
 * ir_f16_gpa: The student's cumulative GPA after the Fall 2016 term.
 * f17_enrolled: The student's status as of Fall term 2017: not enrolled or enrolled.
-* s1_female: The student's gender.
+
+TODO: evaluate scales' validity
+
+# Data
+
+TODO: add all citations for R packages
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
+isrdata <- read.dta13(here("inputs/data/Instructor-Student Relationships Experiment Data_Anonymous.dta"))
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
+expdata <- isrdata %>%
+  filter(use == 1) %>%
+  mutate(across(where(is.numeric), ~na_if(., -99)))
+```
+
+Table \@ref(tab:demotable) shows the demographics for the treatment and control samples. There are no significant differences between the treatment and control groups.
+
+However, there are differences between the students and instructors.
+Notably, the student sample is 60.3% female, while the instructor sample is 78.7% female.
+The student sample is 21.5% White/Caucasian and 51.5% Hispanic American or Latino/a, while the instructor sample is 67.0% White/Caucasian and 6.0% Hispanic American or Latino/a.
+43.1% of the students are first generation university attendees, while only 23.3% of the instructors are.
+
+```{r demotable, echo=FALSE, warning=FALSE, message=FALSE}
+expdata %>% 
+  dplyr::select(treatment, s1_sim, s_female, t_female, s_race, t_race, s1_age, t_age, s_firstgen, t_firstgen, ir_f16_gpa, ir_t_s17enrltot) %>%
+  mutate(s_race = as.factor(s_race), t_race = as.factor(t_race)) %>%
+  summary_factorlist("treatment", c("s_female", "t_female", "s_race", "t_race", "s1_age", "t_age", "s_firstgen", "t_firstgen", "ir_f16_gpa", "ir_t_s17enrltot"),
+                     p = TRUE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = TRUE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N(%) = ") %>%
+  rename(N = "Total N", Missing = "Missing N") %>%
+  knitr::kable(caption = "Demographics for treatment and control groups",
+               booktabs = TRUE, linesep = "") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"), stripe_index =c(2:3,6:13,20,22:23,26))
+```
+
+TODO: graphs showing instructor vs. student demographics. Implications.
+
+Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values.
 
@@
 expdata %>% 
-  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, s1_female) %>%
-  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "ir_f16_gpa", "f17_enrolled", "s1_female"),
+  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, f17_enrolled, s1_female) %>%
+  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "f17_enrolled", "s1_female"),
                      p = FALSE, 
@@
                booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"), stripe_index =c(2,4,6,8,10,13:14))
 ```
@@
 
-Equation \@ref(eq:linmodel) is Robinson et al.'s linear model, used for continuous outcomes, which include immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), end of semester instructor ISR rating (t2_tsr), course grade (grade), and objectively graded exam grade (t2_finalexam).
+Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes, which include complete scale outcomes (immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), and end of semester instructor ISR rating (t2_tsr)) and grade-based outcomes, measured in GPA (course grade (grade) and objectively graded exam grade (t2_finalexam)).
 
@@
 
-Equation \@ref(eq:ordlogmodel) is Robinson et al.'s ordinal logistic model, used for ordinal outcomes, which include end of semester instructor similarity rating (t2_sim1).
+Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end of semester instructor similarity rating (t2_sim1).
 
@@
 
-Equation \@ref(eq:logmodel) is Robinson et al.'s logistic model, used for binary outcomes, which include enrollment in Fall term 2017 (f17_enrolled).
+Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017 (f17_enrolled).
 
@@
 
+TODO: additional models
+Consider:
+More covariates: race, class size, first generation
+New covariates: instructor-student commonalities (gender, race)
+Models looking at subgroups of the sample: Only those with matching traits, only small classes
+Focus on grade as the outcome.
+
 # Results
@@
 
-Table \@ref(tab:correlationtable)
+Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as significant.
+Some measures, unsurprisingly, are highly correlated: grade and standardized grade, and grade and final exam grade.
+Others are moderately correlated: student initial perceptions of similarity and instructor-student relationship, student end-of-term perceptions of similarity and instructor-student relationship, and instructor end-of-term perceptions of similarity and instructor-student relationship.
+These last three pairs could indicate that the similarity perception scales and instructor-student relationship perception scales may not be measuring distinct concepts, or that feelings of similarity and positivity in relationships are strongly associated in the classroom.
 
@@
                      col_totals_prefix = "N (%) = ") %>%
-  select("Total N", "Missing N", "Total")
+  dplyr::select("Total N", "Missing N", "Total")
 
@@
                booktabs = TRUE, linesep = "") %>%
-  kable_styling(latex_options = c("striped", "scale_down"))
+  kable_styling(latex_options = c("scale_down"))
 ```
 
-### Linear models
-
-Table \@ref(tab:studentfirstsimilarity)
-
-```{r studentfirstsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(s1_sim)) %>%
-  finalfit("s1_sim", "treatment") %>%
-  knitr::kable(caption = "Effect of treatment on immediate student similarity rating",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+studentfirstsimilaritymodel <- lm(data = expdata %>%
+                                    filter(!is.na(s1_sim)),
+                                  s1_sim ~ treatment)
+studentsecondsimilaritymodel <- lm(data = expdata %>% 
+                                     filter(!is.na(s2_sim)),
+                                   s2_sim ~ treatment)
+studentsecondrelationshipmodel <- lm(data = expdata %>%
+                                       filter(!is.na(s2_tsr) & !is.na(s1_tsr)),
+                                     s2_tsr ~ treatment + s1_tsr)
+teachersecondrelationshipmodel <- lm(data = expdata %>%
+                                       filter(!is.na(t2_tsr)),
+                                     t2_tsr ~ treatment)
+grademodel <- lm(data = expdata %>%
+                   filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99),
+                 grade ~ treatment + s1_female + ir_f16_gpa)
+examgrademodel <- lm(data = expdata %>%
+                       filter(obj_exam == 1) %>%
+                       filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99),
+                     t2_finalexam ~ treatment + s1_female + ir_f16_gpa)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment,
+                                data = expdata %>%
+                                  filter(!is.na(t2_sim1)),
+                                Hess = TRUE)
+enrollmentfallmodel <- glm(data = expdata %>%
+                             mutate(f17_enrolled = as.factor(f17_enrolled)) %>%
+                             filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99),
+                           f17_enrolled ~ treatment + s1_female + ir_f16_gpa,
+                           family = "binomial")
 ```
 
-Table \@ref(tab:studentsecondsimilarity)
-
-```{r studentsecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(s2_sim)) %>%
-  finalfit("s2_sim", "treatment") %>%
-  knitr::kable(caption = "Effect of treatment on end of semester student similarity rating",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
+```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+stargazer(studentfirstsimilaritymodel,
+          studentsecondsimilaritymodel,
+          studentsecondrelationshipmodel,
+          teachersecondrelationshipmodel,
+          grademodel,
+          examgrademodel,
+          teachersecondsimilarity,
+          enrollmentfallmodel,
+          title = "Model results",
+          align = TRUE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltable",
+          type = "latex",
+          header = FALSE)
 ```
 
-Table \@ref(tab:studentsecondrelationship)
+In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and student initial similarity perception.
+However, the expected increase is only 0.155 on a scale of 1 to 5.
+Model 2 shows a statistically significant relationship between treatment and student end-of-semester similarity perception.
+However, the expected increase is only 0.095 on a scale of 1 to 5.
 
-```{r studentsecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(s2_tsr) & !is.na(s1_tsr)) %>%
-  finalfit("s2_tsr", c("treatment", "s1_tsr")) %>%
-  knitr::kable(caption = "Effect of treatment on immediate student ISR rating, controlling for anticipated ISR",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
+Model 3 shows no significant relationship between treatment and student end-of-semester ISR perception.
+However, it does show a significant correlation between anticipated student ISR and student end-of-semester ISR.
+For every 1 point increase in anticipated ISR on a 1 to 5 scale, the expected increase in end-of-semester ISR is 0.674.
 
-Table \@ref(tab:teachersecondrelationship)
+Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-semester ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent semester.
 
-```{r teachersecondrelationship, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(t2_tsr)) %>%
-  finalfit("t2_tsr", c("treatment")) %>%
-  knitr::kable(caption = "Effect of treatment on end of semester instructor ISR rating",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-Table \@ref(tab:grade)
-
-```{r grade, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99) %>%
-  finalfit("grade", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
-  knitr::kable(caption = "Effect of treatment on course grade, controlling for student gender and prior GPA",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-Table \@ref(tab:examgrade)
-
-```{r examgrade, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(obj_exam == 1) %>%
-  filter(!is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99) %>%
-  finalfit("t2_finalexam", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
-  knitr::kable(caption = "Effect of treatment on objectively graded exam grade, controlling for student gender and prior GPA",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-### Ordinal logistic models
-
-Table \@ref(tab:teachersecondsimilarity)
-
-```{r teachersecondsimilarity, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  filter(!is.na(t2_sim1)) %>%
-  mutate(t2_sim1 = as.factor(t2_sim1)) %>%
-  finalfit("t2_sim1", "treatment") %>%
-  knitr::kable(caption = "Effect of treatment on end of semester instructor similarity rating",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-### Logistic models
-
-Table \@ref(tab:enrollmentfall)
-
-```{r enrollmentfall, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>%
-  mutate(f17_enrolled = as.factor(f17_enrolled)) %>%
-  filter(!is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99) %>%
-  finalfit("f17_enrolled", c("treatment",  "s1_female", "ir_f16_gpa")) %>%
-  knitr::kable(caption = "Effect of treatment on enrollment in next Fall 2017, controlling for student gender and prior GPA",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "scale_down"))
-```
+TODO: clustered residuals
 
@@
 
+TODO: additional results
+
 # Discussion
+
+TODO: discussion
+Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
+Who perceives similarity
+What does it mean for relationships and therefore academic success to be built on similarity.
+
+\newpage
+
+# References

### 2021-04-13 outputs/paper/paper.Rmd
update paper with TODOs from peer review

@@
 ---
-title: "The inefficacy of superficial similarities for establishing instructor-student relationships"
+title: "The inefficacy of superficial similarities for improving instructor-student relationships"
 author: "Amy Farrow"
-date: "April 9th, 2021"
+date: "April 13th, 2021"
 header-includes:
@@
     toc: yes
-subtitle: "Reproducing ‘Taking It to the Next Level’"
-abstract: "This paper reproduces the 2019 paper 'Taking It to the Next Level', which evaluates an intervention to improve instructor-student relationship.
+subtitle: "Replicating ‘Taking It to the Next Level’"
+abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationship.
 The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment.
@@
 These results are consistent with the original paper.
-\\par \\textbf{Keywords:} instructor-student relationship, college, reproduction study"
+TODO: revise to clearliy indicate my own findings \\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
 thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
@@
 
-# functions from https://stefaneng.github.io/apa_correlation_table/
+# functions to create a correlation matrix from https://stefaneng.github.io/apa_correlation_table/ TODO cite
 
@@
 
+@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance.
+In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships (ISRs) and measures of student success.
+Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
+
+TODO: add a bit more about the original experiment.
+
+This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors.
+Linear, logistic, and ordinal logistic models are used to predict outcomes including perceived ISR, grade, and re-enrollment.
+These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grade, exam grade, or future enrollment.
+In addition to the models in the original paper, this paper explores models using different controls and models that consider specific subgroups of the study participants.
+
+TODO: expand on my own findings.
+
+TODO: clearly outline the structure of the whole paper.
+
+## Literature review
+
+TODO: expand this paragraph about interventions, preferably targeting similar ideas to the @citeoriginal paper.
+What has worked?
+Draw attention to results.
+Make sure to refer to any studies that @citeoriginal say are especially influential for their work.
+
 Many interventions to increase completion rates have been suggested and tested.
@@
 
-@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests one intervention.
-In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships and measures of student success.
-Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
+TODO: write a paragraph about research on why similarity builds relationships (similarity -> relationships)
 
-This paper reproduces @citeoriginal's original study, using anonymized data provided by the authors.
-Linear, logistic, and ordinal logistic models are used to predict outcomes including perceived ISR, grade, and re-enrollment.
-These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grade, exam grade, or future enrollment.
-In addition to the models in the original paper, this paper explores models using different controls and models that consider specific subgroups of the study participants.
+TODO: write a paragraph about research on why relationships matter for scholastic performance (relationships -> grades)
 
-# Methods
+TODO: write a paragraph about research on why relationships matter for persistence (relationships -> persistence)
+
+# Methodology
 
@@
 The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized.Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors.
+TODO: what kind of sampling is this?
+What are the implications?
 
@@
 
+TODO: figure out which parts of this list can go to the appendix
+
 @citeoriginal identify key measures. Some are extracted from the student survey at the beginning of the term:
 
-* s1_sim: Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
+* Student similarity perception scale (s1_sim): Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
 These responses were averaged to create a similarity scale.
@@
   - Overall, how similar do you think you and your instructor are?
-* s1_tsr: Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+* Student ISR perception scale (s1_tsr): Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
 These responses were averaged to create an ISR scale.
@@
   - Overall, how much do you think you will learn from this instructor?
-* s1_female: The student's gender.
+* Student gender (s1_female): The student's gender.
 
@@
 
-* s2_sim: At the end of the semester, students answered the above six similarity scale questions again.
-* s2_tsr: At the end of the semester, students answered the above seven ISR scale questions again.
+* Student similarity perception scale (s2_sim): At the end of the semester, students answered the above six similarity scale questions again.
+* Student ISR perception scale (s2_tsr): At the end of the semester, students answered the above seven ISR scale questions again.
 
@@
 
-* t2_sim1: At the end of the semester, instructors answered only one quesiton about similarity with the student, on a scale of 1 to 5.
+* Instructor similarity perception (t2_sim1): At the end of the semester, instructors answered only one quesiton about similarity with the student, on a scale of 1 to 5.
   - Overall, how similar do you think you and STUDENTNAME are?
-* t2_tsr: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+* Instructor ISR perception scale (t2_tsr): At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
 These responses were averaged to create an ISR scale.
@@
   - Overall, how much did STUDENTNAME learn from you?
-* t2_finalexam: Instructors were asked to report the student's grade on their final exam, paper, or project.
+* Final grade (t2_finalexam): Instructors were asked to report the student's grade on their final exam, paper, or project.
 
@@
 
-* grade: The final grade that the student received in the course.
-* std_grade: The student's final grade, standardized against other grades in the course.
-* ir_f16_gpa: The student's cumulative GPA after the Fall 2016 term.
-* f17_enrolled: The student's status as of Fall term 2017: not enrolled or enrolled.
+* Course grade (grade): The final grade that the student received in the course.
+* Standardized course grade (std_grade): The student's final grade, standardized against other grades in the course.
+* CGPA (ir_f16_gpa): The student's cumulative GPA after the Fall 2016 term.
+* Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
 
-TODO: evaluate scales' validity
+TODO: evaluate scales' validity?
 
@@
 
-TODO: add all citations for R packages
+TODO: add all citations for R packages, R, and R markdown
+
+TODO: note where the data is available and describe the dataset basics (cases, variables). Cite the dataset correctly, add to .bib
 
@@
 expdata <- isrdata %>%
-  filter(use == 1) %>%
-  mutate(across(where(is.numeric), ~na_if(., -99)))
+  filter(use == 1) %>% # Select only the cases the original study used.
+  mutate(across(where(is.numeric), ~na_if(., -99))) %>% # Some NAs are shown as -99. Replace with NA.
+  mutate(s_race = as.factor(s_race), # make nominal variables into factors.
+         t_race = as.factor(t_race),
+         teacherid = as.factor(teacherid),
+         f17_enrolled = as.factor(f17_enrolled))
 ```
 
-Table \@ref(tab:demotable) shows the demographics for the treatment and control samples. There are no significant differences between the treatment and control groups.
+## Demographics
+
+Table \@ref(tab:demotable1) shows the nominal covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
 
@@
 
-```{r demotable, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>% 
-  dplyr::select(treatment, s1_sim, s_female, t_female, s_race, t_race, s1_age, t_age, s_firstgen, t_firstgen, ir_f16_gpa, ir_t_s17enrltot) %>%
-  mutate(s_race = as.factor(s_race), t_race = as.factor(t_race)) %>%
-  summary_factorlist("treatment", c("s_female", "t_female", "s_race", "t_race", "s1_age", "t_age", "s_firstgen", "t_firstgen", "ir_f16_gpa", "ir_t_s17enrltot"),
+Table \@ref(tab:demotable2) shows the continuous covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
+
+```{r demotable1, echo=FALSE, warning=FALSE, message=FALSE}
+# Make a summary table of nominal variables, by treatment group.
+demotab <- expdata %>% 
+  dplyr::select(treatment, 
+                s_female, 
+                t_female, 
+                s_race, 
+                t_race, 
+                s_firstgen, 
+                t_firstgen) %>%
+  mutate(treatment = ff_label(treatment, "Group"),
+         s_female = ff_label(s_female, "Student gender"), 
+         t_female = ff_label(t_female, "Instructor gender"), 
+         s_race = ff_label(s_race, "Student race"), 
+         t_race = ff_label(t_race, "Instructor race"), 
+         s_firstgen = ff_label(s_firstgen, "Student first-gen status"), 
+         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status")) %>%
+  summary_factorlist("treatment", 
+                     c("s_female", "t_female", "s_race", "t_race", "s_firstgen", "t_firstgen"),
                      p = TRUE, 
@@
                      col_totals_rowname = "",
-                     total_col = TRUE,
-                     col_totals_prefix = "N(%) = ") %>%
-  rename(N = "Total N", Missing = "Missing N") %>%
-  knitr::kable(caption = "Demographics for treatment and control groups",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("scale_down", "striped"), stripe_index =c(2:3,6:13,20,22:23,26))
+                     total_col = TRUE) %>%
+  rename(N = "Total N", Missing = "Missing N")
+
+# Replace variables that are encoded with their true meanings
+demotab[,4] <- c("","Male", "Female", "Male", "Female",
+                 "White/Caucasian", "Black or African American", "Hispanic American or Latino/a",
+                 "Asian or Pacific Islander", "American Indian or Alaskan", "Middle Eastern",
+                 "Other", "Multi", "White/Caucasian", "Black or African American",
+                 "Hispanic American or Latino/a", "Asian or Pacific Islander",
+                 "Middle Eastern", "Multi", "No", "Yes", "No", "Yes")
+
+# Put it all in a nice table.
+demotab %>%
+  knitr::kable(caption = "Demographic statistics for treatment and control groups, nominal variables",
+               booktabs = TRUE, linesep = "",
+               col.names = c("", "N", "Missing", "", "Control N(%)", "Treatment N(%)", "Total  N(%)", "p")) %>%
+  column_spec(5, width = "5em") %>%
+  column_spec(6, width = "5em") %>%
+  column_spec(7, width = "5em") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
+                            stripe_index =c(2:3, 6:13, 20:21, 26))
+
+```
+
+```{r demotable2, echo=FALSE, warning=FALSE, message=FALSE}
+# make a summary table of continuous variables
+demotab2 <- expdata %>% 
+  dplyr::select(treatment, s1_age, t_age, ir_f16_gpa, n_course) %>%
+  mutate(treatment = ff_label(treatment, "Group"),
+         s1_age = ff_label(s1_age, "Student age"), 
+         t_age = ff_label(t_age, "Instructor age"),  
+         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
+         n_course = ff_label(n_course, "Course size")) %>%
+  summary_factorlist("treatment",
+                     c("s1_age", "t_age", "ir_f16_gpa", "n_course"),
+                     p = TRUE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE) %>%
+  rename(N = "Total N", Missing = "Missing N")
+
+# put in nice table
+demotab2 %>%
+  knitr::kable(caption = "Demographic statistics for treatment and control groups, continuous variables",
+               booktabs = TRUE, linesep = "",
+               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total", "p")) %>%
+  kableExtra::kable_styling(latex_options = c("striped"), 
+                            stripe_index =c(2, 4))
 ```
@@
 
+TODO: graphs showing class size. Implications.
+
+## Outcomes of interest
+
 Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values.
@@
 ```{r summarytable, echo=FALSE, warning=FALSE, message=FALSE}
-expdata %>% 
-  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, f17_enrolled, s1_female) %>%
-  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "f17_enrolled", "s1_female"),
+# Summarize key outcome variables by treatment group
+sumtable <- expdata %>% 
+  dplyr::select(treatment, 
+                s1_sim,
+                s1_tsr,
+                s2_sim,
+                s2_tsr,
+                t2_sim1,
+                t2_tsr, grade,
+                std_grade,
+                t2_finalexam,
+                f17_enrolled) %>%
+  mutate(t2_sim1 = as.factor(t2_sim1)) %>%
+  mutate(treatment = ff_label(treatment, "Group"), 
+         s1_sim = ff_label(s1_sim, "Initial student similarity perception scale"), 
+         s1_tsr = ff_label(s1_tsr, "Initial student ISR perception scale"), 
+         s2_sim = ff_label(s2_sim, "End-of-term student similarity perception scale"), 
+         s2_tsr = ff_label(s2_tsr, "End-of-term student ISR perception scale"), 
+         t2_sim1 = ff_label(t2_sim1, "End-of-term instructor similarity perception"), 
+         t2_tsr = ff_label(t2_tsr, "End-of-term instructor ISR perception scale"), 
+         grade = ff_label(grade, "Course grade"), 
+         std_grade = ff_label(std_grade, "Standardized course grade"), 
+         t2_finalexam = ff_label(t2_finalexam, "Final grade"), 
+         f17_enrolled = ff_label(f17_enrolled, "Persistence")) %>%
+  summary_factorlist("treatment",
+                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr",
+                       "grade", "std_grade", "t2_finalexam", "f17_enrolled"),
                      p = FALSE, 
@@
                      col_totals_prefix = "N(%) = ") %>%
-  rename(N = "Total N", Missing = "Missing N") %>%
-  knitr::kable(caption = "Descriptive statistics for treatment and control groups",
-               booktabs = TRUE, linesep = "") %>%
-  kableExtra::kable_styling(latex_options = c("scale_down", "striped"), stripe_index =c(2,4,6,8,10,13:14))
+  rename(N = "Total N", Missing = "Missing N")
+
+# Rename variables for readability
+sumtable[,1] <- c("", "Similarity", "ISR", " Similarity", " ISR", "Similarity ", "", "", "", "", "ISR ", "Course grade", "Stand. grade", "Final grade", "Peristence", "")
+# replace nominal encodings with true values
+sumtable[15,4] <- "No"
+sumtable[16,4] <- "Yes"
+
+# put in table.
+sumtable %>%
+  knitr::kable(caption = "Outcomes of interest for treatment and control groups",
+               booktabs = TRUE, linesep = "",
+               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total")) %>%
+  pack_rows("Initial student perception", 2, 3) %>%
+  pack_rows("End-of-term student perception", 4, 5) %>%
+  pack_rows("End-of-term instructor perception", 6, 11) %>%
+  pack_rows("Student outcomes", 12, 15) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
+                            stripe_index =c(2,4,6,8,10,13:14))
 ```
+
+Replicating @citeoriginal's study, Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as significant.
+Some measures, unsurprisingly, are highly correlated: grade and standardized grade, and grade and final exam grade.
+Others are moderately correlated: student initial perceptions of similarity and instructor-student relationship, student end-of-term perceptions of similarity and instructor-student relationship, and instructor end-of-term perceptions of similarity and instructor-student relationship.
+These last three pairs could indicate that the similarity perception scales and instructor-student relationship perception scales may not be measuring distinct concepts, or that feelings of similarity and positivity in relationships are strongly associated in the classroom.
+
+```{r correlationtable, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
+                grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, 
+                s1_female, n_course) %>%
+  mutate(treatment = ff_label(treatment, "Group"), 
+         s1_sim = ff_label(s1_sim, "Initial student similarity perception scale"), 
+         s1_tsr = ff_label(s1_tsr, "Initial student ISR perception scale"), 
+         s2_sim = ff_label(s2_sim, "End-of-term student similarity perception scale"), 
+         s2_tsr = ff_label(s2_tsr, "End-of-term student ISR perception scale"), 
+         t2_sim1 = ff_label(t2_sim1, "End-of-term instructor similarity perception"), 
+         t2_tsr = ff_label(t2_tsr, "End-of-term instructor ISR perception scale"), 
+         grade = ff_label(grade, "Course grade"), 
+         std_grade = ff_label(std_grade, "Standardized course grade"), 
+         t2_finalexam = ff_label(t2_finalexam, "Final grade"),
+         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
+         f17_enrolled = ff_label(f17_enrolled, "Persistence"),
+         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
+         n_course = ff_label(n_course, "Course size")) %>%
+  summary_factorlist("treatment", 
+                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade",
+                       "std_grade", "t2_finalexam", "ir_f16_gpa", "n_course"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, n_course))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# rename the variables for table readability
+rownames(table) <- c("Similarity", "ISR", " Similarity", " ISR", "Similarity ", "ISR ", "Course", "Stand. course", "Final", "CGPA", "Size")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for continuous and ordinal key variables and outcomes",
+        booktabs = TRUE,
+        col.names = c("N", "Missing", "Mean (SD)", "Similarity", "ISR", "Similarity","ISR",
+                      "Similarity", "ISR", "Course", "Stand. course", "Final", "CGPA", "Size")) %>%
+  column_spec(1, width = "8em") %>%
+  column_spec(4, width = "4em") %>%
+  column_spec(5, width = "4em") %>%
+  column_spec(6, width = "4em") %>%
+  column_spec(7, width = "4em") %>%
+  column_spec(8, width = "4em") %>%
+  column_spec(9, width = "4em") %>%
+  column_spec(10, width = "4em") %>%
+  column_spec(11, width = "4em") %>%
+  column_spec(12, width = "4em") %>%
+  column_spec(13, width = "2.5em") %>%
+  column_spec(14, width = "2.5em") %>%
+  pack_rows("Initial student perception", 1, 2) %>%
+  pack_rows("End-of-term student perception", 3, 4) %>%
+  pack_rows("End-of-term instructor perception", 5, 6) %>%
+  pack_rows("Grades", 7, 10) %>%
+  pack_rows("Course", 11, 11) %>%
+  add_header_above(c(" " = 4, "Initial student" = 2, "End-of-term student" = 2,
+                     "End-of-term instructor" = 2, "Grades" = 4, "Course" = 1), 
+                   align = "l") %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+TODO: graphs that show distributions of key scales (s1_sim, s2_sim, t2_sim1, s1_tsr, s2_tsr, t2_tsr, grade, t2_finalexam) between treatment and control.
+These will be boring and should be very small.
+Maybe set up at 2x3 grid for the sims and tsrs.
+Maybe push to appendix.
+
+## Missing data
+
+TODO: any patterns in missing data.
+
+TODO: did attrition follow any particular pattern? Did some types of students stop responding in between the first and second surveys? Non-response.
+
+## Data selection
+
+TODO: what data was excluded and why?
 
@@
 
-## Reproduction models
+## Replication models
+
+TODO: why did @citeoriginal pick these? What do I think about the selection?
 
@@
 
-TODO: additional models
-Consider:
-More covariates: race, class size, first generation
-New covariates: instructor-student commonalities (gender, race)
-Models looking at subgroups of the sample: Only those with matching traits, only small classes
+TODO: additional models.
+Consider: More covariates (race, first generation, matching gender and race); Models looking at subgroups of the sample (Only those with matching traits, only small classes).
 Focus on grade as the outcome.
@@
 
-## Reproduction results
-
-Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as significant.
-Some measures, unsurprisingly, are highly correlated: grade and standardized grade, and grade and final exam grade.
-Others are moderately correlated: student initial perceptions of similarity and instructor-student relationship, student end-of-term perceptions of similarity and instructor-student relationship, and instructor end-of-term perceptions of similarity and instructor-student relationship.
-These last three pairs could indicate that the similarity perception scales and instructor-student relationship perception scales may not be measuring distinct concepts, or that feelings of similarity and positivity in relationships are strongly associated in the classroom.
-
-```{r correlationtable, echo=FALSE, warning=FALSE, message=FALSE}
-summary <- expdata %>% 
-  dplyr::select(treatment, s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, s1_female) %>%
-  summary_factorlist("treatment", c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade", "std_grade", "t2_finalexam", "ir_f16_gpa"),
-                     p = FALSE, 
-                     add_dependent_label = TRUE, 
-                     dependent_label_prefix = "",
-                     add_col_totals = FALSE,
-                     add_row_totals = TRUE,
-                     include_row_missing_col = TRUE,
-                     col_totals_rowname = "",
-                     total_col = TRUE,
-                     col_totals_prefix = "N (%) = ") %>%
-  dplyr::select("Total N", "Missing N", "Total")
-
-matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
-                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa)))) 
-summary %>%
-  bind_cols(matrix) %>%
-  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N") %>%
-  kable(caption = "Descriptive statistics and correlation matrix for key continuous variables",
-               booktabs = TRUE, linesep = "") %>%
-  kable_styling(latex_options = c("scale_down"))
-```
+## Replication results
 
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
 studentfirstsimilaritymodel <- lm(data = expdata %>%
                                     filter(!is.na(s1_sim)),
-                                  s1_sim ~ treatment)
+                                  s1_sim ~ treatment + n_course + teacherid)
 studentsecondsimilaritymodel <- lm(data = expdata %>% 
                                      filter(!is.na(s2_sim)),
-                                   s2_sim ~ treatment)
+                                   s2_sim ~ treatment + n_course + teacherid)
 studentsecondrelationshipmodel <- lm(data = expdata %>%
                                        filter(!is.na(s2_tsr) & !is.na(s1_tsr)),
-                                     s2_tsr ~ treatment + s1_tsr)
+                                     s2_tsr ~ treatment + s1_tsr + n_course + teacherid)
 teachersecondrelationshipmodel <- lm(data = expdata %>%
                                        filter(!is.na(t2_tsr)),
-                                     t2_tsr ~ treatment)
+                                     t2_tsr ~ treatment + n_course + teacherid)
 grademodel <- lm(data = expdata %>%
                    filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99),
-                 grade ~ treatment + s1_female + ir_f16_gpa)
+                 grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid)
 examgrademodel <- lm(data = expdata %>%
@@
                        filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99),
-                     t2_finalexam ~ treatment + s1_female + ir_f16_gpa)
-teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment,
+                     t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
                                 data = expdata %>%
@@
 enrollmentfallmodel <- glm(data = expdata %>%
-                             mutate(f17_enrolled = as.factor(f17_enrolled)) %>%
                              filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99),
-                           f17_enrolled ~ treatment + s1_female + ir_f16_gpa,
+                           f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
                            family = "binomial")
@@
 ```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
 stargazer(studentfirstsimilaritymodel,
@@
           type = "latex",
-          header = FALSE)
+          header = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Teacher fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Instructor sim. 2",
+                             "Persistence"),
+          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA", "Course size"))
 ```
 
+TODO: reread paper & Stata code and ensure I caught all models that they report.
+
 In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and student initial similarity perception.
-However, the expected increase is only 0.155 on a scale of 1 to 5.
+However, the expected increase is only 0.158 on a scale of 1 to 5.
 Model 2 shows a statistically significant relationship between treatment and student end-of-semester similarity perception.
-However, the expected increase is only 0.095 on a scale of 1 to 5.
+However, the expected increase is only 0.101 on a scale of 1 to 5.
 
@@
 
-TODO: clustered residuals
+TODO: any problems with the models. General performance.
+
+TODO: clustered residuals (low priority)
 
@@
 
+TODO: Shiny.
+
 # Discussion
 
-TODO: discussion
-Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
-Who perceives similarity
-What does it mean for relationships and therefore academic success to be built on similarity.
+TODO: does this study have external validity?
+
+TODO: Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
+
+TODO: Who perceives similarity
+
+TODO: What does it mean for relationships and therefore academic success to be built on similarity.
 

### 2021-04-17 outputs/exploratory.Rmd
add graphs; fix instructor covariate tables; add my own data selection methods.

@@
 ```{r}
-expdata %>%
-  select(teacherid,
-         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7) %>%
+instructorrace <- expdata %>%
+  dplyr::select(teacherid,
+         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
   unique() %>%
-  summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
+  dplyr::summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
             black = sum(t1_race_2, na.rm = TRUE)/n(),
@@
             other = sum(t1_race_7, na.rm = TRUE)/n(),
+            mixed = sum(t_multi_race > 1, na.rm = TRUE)/n(),
+            total = n())
+```
+```{r}
+studentrace <- expdata %>%
+  dplyr::select(id,
+         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
+  unique() %>%
+  dplyr::summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
+            black = sum(s1_race_2, na.rm = TRUE)/n(),
+            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
+            asian = sum(s1_race_4, na.rm = TRUE)/n(),
+            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
+            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
+            other = sum(s1_race_7, na.rm = TRUE)/n(),
+            mixed = sum(s_multi_race > 1, na.rm = TRUE)/n(),
             total = n())
@@
   geom_histogram(position = "dodge")
+
+theirs <- isrdata %>%
+  filter(use == 1) %>%
+  dplyr::select(teacherid) %>%
+  unique()
+
+mine <- expdata %>%
+  dplyr::select(teacherid) %>%
+  unique()
 ```

### 2021-04-17 outputs/paper/paper.Rmd
add graphs; fix instructor covariate tables; add my own data selection methods.

@@
 library(MASS) # for ordinal logistic model
+library(cowplot) # for putting plots side by side
 
 # NOTE: script 01_ must have been run already
-
-# functions to create a correlation matrix from https://stefaneng.github.io/apa_correlation_table/ TODO cite
 
@@
 TODO: write a paragraph about research on why relationships matter for scholastic performance (relationships -> grades)
+@Abrami_Mizener_1985's study of college students found a .
+One study of Taiwanese high school students explored the link between shared education and life values and student performance [@Lai_2015].
+While they found a significant, but small, positive relationship between life values and student performance on analytic tests, education values gave mixed results depending on the natures of the education value and the analytic test in question [@Lai_2015].
 
@@
 
-TODO: figure out which parts of this list can go to the appendix
+@citeoriginal identify key measures. Full descriptions of all measures are available in [Appendix A].
 
-@citeoriginal identify key measures. Some are extracted from the student survey at the beginning of the term:
+Some are extracted from the student survey at the beginning of the term:
 
-* Student similarity perception scale (s1_sim): Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
-These responses were averaged to create a similarity scale.
-  - Overall, how similar to your instructor's values do you think your values are?
-  - How similar are your goals for the course and your instructor's goals?
-  - In general, how similar do you think your views about the course content and your instructor's are?
-  - How much do you think you have in common with your instructor?
-  - How similar do you think your personality is compared to your instructor's?
-  - Overall, how similar do you think you and your instructor are?
-* Student ISR perception scale (s1_tsr): Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create an ISR scale.
-  - How much do you think you will enjoy learning from this instructor?
-  - How friendly do you think this instructor will be towards you?
-  - How encouraging do you think this instructor will be towards you?
-  - If you came back to visit this instructor three years from now, how excited do you think they would be?
-  - How motivating do you think you will find this instructor's class?
-  - How caring do you think this instructor will be towards you?
-  - Overall, how much do you think you will learn from this instructor?
-* Student gender (s1_female): The student's gender.
+  1.
+Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
+These responses were averaged to create a student similarity perception scale.
+  2.
+Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create a student ISR perception scale.
+  3. Student gender
 
@@
 
-* Student similarity perception scale (s2_sim): At the end of the semester, students answered the above six similarity scale questions again.
-* Student ISR perception scale (s2_tsr): At the end of the semester, students answered the above seven ISR scale questions again.
+  4. At the end of the semester, students answered the student similarity perception scale questions again.
+  5. At the end of the semester, students answered the student ISR perception scale questions again.
 
@@
 
-* Instructor similarity perception (t2_sim1): At the end of the semester, instructors answered only one quesiton about similarity with the student, on a scale of 1 to 5.
+  6. Instructor similarity perception: At the end of the semester, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
       - Overall, how similar do you think you and STUDENTNAME are?
-* Instructor ISR perception scale (t2_tsr): At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+  7.
+Instructor ISR perception scale: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
 These responses were averaged to create an ISR scale.
-  - How much did you enjoy helping STUDENTNAME learn?
-  - How caring was STUDENTNAME towards you?
-  - How often did you say something encouraging to STUDENTNAME?
-  - How friendly was STUDENTNAME towards you?
-  - If this student came back to visit you three years from now, how excited would you be?
-  - How motivating did STUDENTNAME find the activities that you plan for class?
-  - Overall, how much did STUDENTNAME learn from you?
-* Final grade (t2_finalexam): Instructors were asked to report the student's grade on their final exam, paper, or project.
+  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
 
@@
 
-* Course grade (grade): The final grade that the student received in the course.
-* Standardized course grade (std_grade): The student's final grade, standardized against other grades in the course.
-* CGPA (ir_f16_gpa): The student's cumulative GPA after the Fall 2016 term.
-* Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
+  9. The final grade that the student received in the course.
+  10. The student's final grade, standardized against other grades in the course.
+  11. The student's cumulative GPA (CGPA) after the Fall 2016 term.
+  12. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
 
@@
 TODO: add all citations for R packages, R, and R markdown
+The correlation matrix was made using functions by @apacorrcode.
 
@@
 expdata <- isrdata %>%
-  filter(use == 1) %>% # Select only the cases the original study used.
+  filter(t1_consent == 1 # Only include if instructor consented to first survey
+         & t1_complete == 1 # Only included if instructor completed first survey
+         & t1_timer_consent_3 > 1 # Only include if instructor read consent
+         & s1_consent == 1 # Only include if student consented to first survey
+         & s1_timer_consent_3 > 1 # Only include if student read consent
+         & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
+         & t2_complete == 1 # Only include if the instructor completed the whole second survey
+         # The rest of the lines make sure that the survey taker looked at each page for more than three seconds
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & s1_timer_gtky1_3 > 10
+         & s1_timer_gtky2_3 > 10
+         & s1_timer_gtky3_3 > 10
+         & s1_timerpre1_3 > 10
+         & s1_timerpre2_3 > 10
+         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
+         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
+         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
+         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
+         & s1_timerpost1_3 > 10
+         & s1_timerpost2_3 > 10
+         & s1_timerdemo1_3 > 10
+         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
+         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
+         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
+         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
+         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
+         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
+         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))
+         & t2_timer > 10
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         )
+
+# Alternatively: expdata <- isrdata %>% filter(use == 1) %>% # Select only the cases the original study used.
+
+expdata <- expdata %>%
   mutate(across(where(is.numeric), ~na_if(., -99))) %>% # Some NAs are shown as -99. Replace with NA.
@@
 
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
+# The original coding of race contained errors -- ex. a professor who selected White and Asian in the survey,
+# and was marked as Asian for most cases, but white for one random case.
+# This remakes the variable t_race, for consistency
+expdata <- expdata %>%
+  mutate(t1_race_1 = replace_na(as.numeric(t1_race_1), 0),
+         t1_race_2 = replace_na(as.numeric(t1_race_2), 0),
+         t1_race_3 = replace_na(as.numeric(t1_race_3), 0),
+         t1_race_4 = replace_na(as.numeric(t1_race_4), 0),
+         t1_race_5 = replace_na(as.numeric(t1_race_5), 0),
+         t1_race_6 = replace_na(as.numeric(t1_race_6), 0),
+         t1_race_7 = replace_na(as.numeric(t1_race_7), 0)) %>%
+  mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
+                         ifelse(t1_race_1 == 1, ifelse(t1_race_2 == 1 & t1_race_3 != 1 & t1_race_4 != 1 & t1_race_5 != 1 & t1_race_6 != 1,
+                                                       "Black and White",
+                                                       ifelse(t1_race_2 != 1 & t1_race_3 == 1 & t1_race_4 != 1 & t1_race_5 != 1 & t1_race_6 != 1,
+                                                              "White Hispanic", "Other multi")), 
+                                ifelse(t1_race_2 == 1 & t1_race_3 == 1, "Black Hispanic", "Other multi")), 
+                         ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
+                                ifelse(t1_race_1 == 1, "White",
+                                       ifelse(t1_race_2 == 1, "Black",
+                                              ifelse(t1_race_3 == 1, "Hispanic",
+                                                     ifelse(t1_race_4, "Asian",
+                                                            ifelse(t1_race_5, "American Indian",
+                                                                   ifelse(t1_race_6, "Middle Eastern",
+                                                                          "Other")))))))))
+
+expdata <- expdata %>%
+  mutate(s1_race_1 = replace_na(as.numeric(s1_race_1), 0),
+         s1_race_2 = replace_na(as.numeric(s1_race_2), 0),
+         s1_race_3 = replace_na(as.numeric(s1_race_3), 0),
+         s1_race_4 = replace_na(as.numeric(s1_race_4), 0),
+         s1_race_5 = replace_na(as.numeric(s1_race_5), 0),
+         s1_race_6 = replace_na(as.numeric(s1_race_6), 0),
+         s1_race_7 = replace_na(as.numeric(s1_race_7), 0)) %>%
+  mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
+                         ifelse(s1_race_1 == 1, ifelse(s1_race_2 == 1 & s1_race_3 != 1 & s1_race_4 != 1 & s1_race_5 != 1 & s1_race_6 != 1,
+                                                       "Black and White",
+                                                       ifelse(s1_race_2 != 1 & s1_race_3 == 1 & s1_race_4 != 1 & s1_race_5 != 1 & s1_race_6 != 1,
+                                                              "White Hispanic", "Other multi")), 
+                                ifelse(s1_race_2 == 1 & s1_race_3 == 1, "Black Hispanic", "Other multi")), 
+                         ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
+                                ifelse(s1_race_1 == 1, "White",
+                                       ifelse(s1_race_2 == 1, "Black",
+                                              ifelse(s1_race_3 == 1, "Hispanic",
+                                                     ifelse(s1_race_4, "Asian",
+                                                            ifelse(s1_race_5, "American Indian",
+                                                                   ifelse(s1_race_6, "Middle Eastern",
+                                                                          "Other")))))))))
+```
+
 ## Demographics
 
-Table \@ref(tab:demotable1) shows the nominal covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
+Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
 
-However, there are differences between the students and instructors.
-Notably, the student sample is 60.3% female, while the instructor sample is 78.7% female.
-The student sample is 21.5% White/Caucasian and 51.5% Hispanic American or Latino/a, while the instructor sample is 67.0% White/Caucasian and 6.0% Hispanic American or Latino/a.
-43.1% of the students are first generation university attendees, while only 23.3% of the instructors are.
+Table \@ref(tab:demotable2) shows the instructor covariates.
+Because instructors are counted multiple times, for each student in the treatment and control groups that is in their course, only the totals are shown.
+The same covariates, presented by treatment and control groups, are shown in [Appendix B].
+
+There are differences between the students and instructors.
+Notably, the student sample is 60.3% female, while the instructor sample is 77.5% female.
+The student sample is 18.6% White, while the instructor sample is 39.2% White.
+The student sample is 43.5% Hispanic, while the instructor sample is only 3.3%.
+Finally, the student sample is 9.1% multiracial, but the instructor sample is 51.7% multiracial.
+Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classed as multiracial.
+For this reason, it is very difficult to draw conclusions.
 
@@
                 s_female, 
-                t_female, 
                 s_race, 
-                t_race, 
                 s_firstgen,
-                t_firstgen) %>%
+                s1_age,
+                ir_f16_gpa,
+                year) %>%
+  mutate(year = as.factor(year)) %>%
   mutate(treatment = ff_label(treatment, "Group"),
          s_female = ff_label(s_female, "Student gender"), 
-         t_female = ff_label(t_female, "Instructor gender"), 
+         s1_age = ff_label(s1_age, "Student age"),
          s_race = ff_label(s_race, "Student race"), 
-         t_race = ff_label(t_race, "Instructor race"), 
+         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
          s_firstgen = ff_label(s_firstgen, "Student first-gen status"), 
-         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status")) %>%
+         year = ff_label(year, "Year")) %>%
   summary_factorlist("treatment", 
-                     c("s_female", "t_female", "s_race", "t_race", "s_firstgen", "t_firstgen"),
+                     c("s_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
                      p = TRUE, 
@@
                      col_totals_rowname = "",
-                     total_col = TRUE) %>%
+                     total_col = TRUE,
+                     col_totals_prefix = "N(%) = ") %>%
   rename(N = "Total N", Missing = "Missing N")
@@
 # Replace variables that are encoded with their true meanings
-demotab[,4] <- c("","Male", "Female", "Male", "Female",
-                 "White/Caucasian", "Black or African American", "Hispanic American or Latino/a",
-                 "Asian or Pacific Islander", "American Indian or Alaskan", "Middle Eastern",
-                 "Other", "Multi", "White/Caucasian", "Black or African American",
-                 "Hispanic American or Latino/a", "Asian or Pacific Islander",
-                 "Middle Eastern", "Multi", "No", "Yes", "No", "Yes")
+demotab[2:3,4] <- c("Male", "Female")
+demotab[16:17,4] <- c("No", "Yes")
 
@@
 demotab %>%
-  knitr::kable(caption = "Demographic statistics for treatment and control groups, nominal variables",
+  knitr::kable(caption = "Student covariates for treatment and control groups",
                booktabs = TRUE, linesep = "",
-               col.names = c("", "N", "Missing", "", "Control N(%)", "Treatment N(%)", "Total  N(%)", "p")) %>%
+               ) %>%
   column_spec(5, width = "5em") %>%
@@
   kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
-                            stripe_index =c(2:3, 6:13, 20:21, 26))
+                            stripe_index =c(2:3, 16:17, 22))
 
@@
 demotab2 <- expdata %>% 
-  dplyr::select(treatment, s1_age, t_age, ir_f16_gpa, n_course) %>%
-  mutate(treatment = ff_label(treatment, "Group"),
-         s1_age = ff_label(s1_age, "Student age"), 
-         t_age = ff_label(t_age, "Instructor age"),  
-         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
+  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
+  dplyr::select(teacherid, t_age, t_firstgen, t_race, t_female, n_course) %>%
+  unique() %>%
+  mutate(t_age = ff_label(t_age, "Instructor age"), 
+         t_female = ff_label(t_female, "Instructor gender"),
+         t_race = ff_label(t_race, "Instructor race"),
+         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"),
          n_course = ff_label(n_course, "Course size")) %>%
-  summary_factorlist("treatment",
-                     c("s1_age", "t_age", "ir_f16_gpa", "n_course"),
+  summary_factorlist("teacherid",
+                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
                      p = TRUE, 
@@
                      col_totals_rowname = "",
-                     total_col = TRUE) %>%
+                     total_col = TRUE,
+                     add_col_totals = TRUE,
+                     col_totals_prefix = "N = ") %>%
   rename(N = "Total N", Missing = "Missing N")
+
+demotab2[2:3,4] <- c("Male", "Female")
+demotab2[13:14,4] <- c("No", "Yes")
 
@@
 demotab2 %>%
-  knitr::kable(caption = "Demographic statistics for treatment and control groups, continuous variables",
+  dplyr::select(-"0", -"1", -"p") %>%
+  knitr::kable(caption = "Teacher covariates",
                booktabs = TRUE, linesep = "",
-               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total", "p")) %>%
+               col.names = c("", "N", "Missing", "", "")) %>%
   kableExtra::kable_styling(latex_options = c("striped"), 
-                            stripe_index =c(2, 4))
+                            stripe_index =c(2:3, 13:14, 16))
 ```
 
-TODO: graphs showing instructor vs. student demographics. Implications.
+In Figure \@ref(fig:racegraph), we can see that more students identified as Hispanic, while more instructors identified as White and multi-racial.
 
-TODO: graphs showing class size. Implications.
+```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.cap="Bar chart illustrating racial identification proprtions of participants"}
+instructorrace <- expdata %>%
+  dplyr::select(teacherid,
+         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
+  unique() %>%
+  dplyr::summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
+            black = sum(t1_race_2, na.rm = TRUE)/n(),
+            hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
+            asian = sum(t1_race_4, na.rm = TRUE)/n(),
+            americanindian = sum(t1_race_5, na.rm = TRUE)/n(),
+            middleeastern = sum(t1_race_6, na.rm = TRUE)/n(),
+            other = sum(t1_race_7, na.rm = TRUE)/n(),
+            mixed = sum(t_multi_race > 1, na.rm = TRUE)/n())
+studentrace <- expdata %>%
+  dplyr::select(id,
+         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
+  unique() %>%
+  dplyr::summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
+            black = sum(s1_race_2, na.rm = TRUE)/n(),
+            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
+            asian = sum(s1_race_4, na.rm = TRUE)/n(),
+            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
+            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
+            other = sum(s1_race_7, na.rm = TRUE)/n(),
+            mixed = sum(s_multi_race > 1, na.rm = TRUE)/n())
+
+bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
+          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
+  dplyr::select(-nms...3) %>%
+  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
+  ggplot() +
+  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
+  scale_x_discrete(labels = c("American Indian", "Asian", "Black", "Hispanic", "Middle Eastern", "Multiple", "Other", "White")) +
+  scale_fill_discrete(labels = c("Instructors", "Students")) +
+  labs(x = "Self-identified race", y = "Proportion of sample", fill = "Group", title = "Self-identified racial proportions different for instructors and students") +
+  theme_minimal()
+```
+
+In Figure \@ref(fig:classsizegraph), we can see that most class sizes are between 25 and 50 students, but there are some that are much larger.
+
+```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height = 3, fig.cap="Histogram illustrating class sizes"}
+expdata %>%
+  dplyr::select(teacherid, n_course) %>%
+  unique() %>%
+  ggplot(aes(x = n_course)) +
+  geom_histogram(binwidth = 5) +
+  labs(x = "Class size", y = "Frequency", title = "Some class sizes are very large") +
+  theme_minimal()
+```
 
@@
                             stripe_index =c(2,4,6,8,10,13:14))
+```
+
+Figure \@ref(fig:continuousoutcomes) shows key continuous outcome distributions for both treatment and control.
+The only visible differences are in the initial student similarity perception and end-of-term student similarity perception.
+
+```{r continuousoutcomes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height= 10, fig.cap="Histograms illustrating continous and ordinal key outcomes"}
+s1_simplot <- expdata %>% 
+  ggplot(aes(x = s1_sim, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "Initial student similarity", y = "Frequency", fill = "Group") +
+  theme_minimal()
+
+s1_tsrplot <- expdata %>% 
+  ggplot(aes(x = s1_tsr, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "Initial student ISR", y = "Frequency") +
+  theme_minimal()
+
+s2_simplot <- expdata %>% 
+  ggplot(aes(x = s2_sim, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "End-of-term student similarity", y = "Frequency") +
+  theme_minimal()
+
+s2_tsrplot <- expdata %>% 
+  ggplot(aes(x = s2_tsr, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "End-of-term student ISR", y = "Frequency") +
+  theme_minimal()
+
+t2_sim1plot <- expdata %>% 
+  ggplot(aes(x = t2_sim1, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "End-of-term instructor similarity", y = "Frequency") +
+  theme_minimal()
+
+t2_tsrplot <- expdata %>% 
+  ggplot(aes(x = t2_tsr, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "End-of-term instructor ISR", y = "Frequency") +
+  theme_minimal()
+
+gradeplot <- expdata %>% 
+  ggplot(aes(x = grade, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "Course grade", y = "Frequency") +
+  theme_minimal()
+
+std_gradeplot <- expdata %>% 
+  ggplot(aes(x = std_grade, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "Standardized course grade", y = "Frequency") +
+  theme_minimal()
+
+t2_finalexamplot <- expdata %>% 
+  ggplot(aes(x = t2_finalexam, fill = treatment)) +
+  geom_histogram(position = "dodge") +
+  labs(x = "Final grade", y = "Frequency") +
+  theme_minimal()
+
+combined <- plot_grid(s1_simplot + theme(legend.position = "none"), s1_tsrplot + theme(legend.position = "none"),
+          s2_simplot + theme(legend.position = "none"), s2_tsrplot + theme(legend.position = "none"),
+          t2_sim1plot + theme(legend.position = "none"), t2_tsrplot + theme(legend.position = "none"),
+          std_gradeplot + theme(legend.position = "none"), t2_finalexamplot + theme(legend.position = "none"),
+          ncol = 2)
+
+legend <- get_legend(s1_simplot + 
+                       guides(color = guide_legend(nrow = 1)) +
+                       theme(legend.position = "bottom"))
+
+title <- ggdraw() + 
+  draw_label(
+    "Most key outcomes unaffected by treatment",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+plot_grid(title,
+          combined,
+          legend,
+          ncol = 1,
+          rel_heights = c(.1, 2, .1))
 ```
@@
 ```
-
-TODO: graphs that show distributions of key scales (s1_sim, s2_sim, t2_sim1, s1_tsr, s2_tsr, t2_tsr, grade, t2_finalexam) between treatment and control.
-These will be boring and should be very small.
-Maybe set up at 2x3 grid for the sims and tsrs.
-Maybe push to appendix.
 
@@
 
+# (APPENDIX) Appendix {-}
+
+# Appendix A
+
+Complete list of key measures
+
+* Initial student survey:
+  1. Student similarity perception scale (s1_sim)
+      - Overall, how similar to your instructor's values do you think your values are?
+      - How similar are your goals for the course and your instructor's goals?
+      - In general, how similar do you think your views about the course content and your instructor's are?
+      - How much do you think you have in common with your instructor?
+      - How similar do you think your personality is compared to your instructor's?
+      - Overall, how similar do you think you and your instructor are?
+  2. Student ISR perception scale (s1_tsr)
+      - How much do you think you will enjoy learning from this instructor?
+      - How friendly do you think this instructor will be towards you?
+      - How encouraging do you think this instructor will be towards you?
+      - If you came back to visit this instructor three years from now, how excited do you think they would be?
+      - How motivating do you think you will find this instructor's class?
+      - How caring do you think this instructor will be towards you?
+      - Overall, how much do you think you will learn from this instructor?
+  3. Student gender (s1_female)
+* End of term student survey:
+  4. Student similarity perception scale (s2_sim)
+  5. Student ISR perception scale (s2_tsr)
+* End of term instructor survey:
+  6. Instructor similarity perception (t2_sim1):
+      - Overall, how similar do you think you and STUDENTNAME are?
+  7. Instructor ISR perception scale (t2_tsr):
+      - How much did you enjoy helping STUDENTNAME learn?
+      - How caring was STUDENTNAME towards you?
+      - How often did you say something encouraging to STUDENTNAME?
+      - How friendly was STUDENTNAME towards you?
+      - If this student came back to visit you three years from now, how excited would you be?
+      - How motivating did STUDENTNAME find the activities that you plan for class?
+      - Overall, how much did STUDENTNAME learn from you?
+  8. Final grade (t2_finalexam): Instructors were asked to report the student's grade on their final exam, paper, or project.
+* University internal records:
+  9. Course grade (grade): The final grade that the student received in the course.
+  10. Standardized course grade (std_grade): The student's final grade, standardized against other grades in the course.
+  11. CGPA (ir_f16_gpa): The student's cumulative GPA after the Fall 2016 term.
+  12. Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
+
+# Appendix B
+
+```{r demotable3, echo=FALSE, warning=FALSE, message=FALSE}
+# Make a summary table of nominal variables, by treatment group.
+demotab3 <- expdata %>% 
+  dplyr::select(treatment, 
+                t_female, 
+                t_race, 
+                t_firstgen,
+                t_age,
+                n_course) %>%
+  mutate(treatment = ff_label(treatment, "Group"),
+         t_female = ff_label(t_female, "Teacher gender"), 
+         t_age = ff_label(t_age, "Teacher age"),
+         t_race = ff_label(t_race, "Teacher race"), 
+         t_firstgen = ff_label(t_firstgen, "Teacher first-gen status"), 
+         n_course = ff_label(n_course, "Course size")) %>%
+  summary_factorlist("treatment", 
+                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
+                     p = TRUE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = TRUE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N(%) = ") %>%
+  rename(N = "Total N", Missing = "Missing N")
+
+# Replace variables that are encoded with their true meanings
+demotab3[2:3,4] <- c("Male", "Female")
+demotab3[13:14,4] <- c("No", "Yes")
+
+# Put it all in a nice table.
+demotab3 %>%
+  knitr::kable(caption = "Teacher covariates for treatment and control groups",
+               booktabs = TRUE, linesep = "",
+               ) %>%
+  column_spec(5, width = "5em") %>%
+  column_spec(6, width = "5em") %>%
+  column_spec(7, width = "5em") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
+                            stripe_index =c(2:3, 13:14, 16))
+
+```
+
+\newpage
+
 # References

### 2021-04-17 outputs/paper/paper.Rmd
update bibtex, intro, replication results. add scale evaluation, missing data, attrition, data selection, additional models, exploratory model replications, additional results, missing data matrix.

@@
 library(cowplot) # for putting plots side by side
+library(ltm) # for checking scale validity
 
@@
 Disparities in college completion lead to further entrenchment of long-standing inequalities [@finishline].
-Thus, measures to help students persist in college completion are desirable to reduce wasted resources, and to increase societal equity.
+Thus, measures to help students persist in college completion are desirable to reduce wasted resources and to increase societal equity.
 
@@
 Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
-
-TODO: add a bit more about the original experiment.
+The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not.
+Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.
 
 This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors.
-Linear, logistic, and ordinal logistic models are used to predict outcomes including perceived ISR, grade, and re-enrollment.
+First, methods are discussed, and the measures and scales created by @citeoriginal are evaluated.
+Second, different demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated.
+Third, @citeoriginal's models are replicated, and additional models are explored.
+Linear, logistic, and ordinal logistic models are replicated to predict outcomes including perceived ISR, grade, and persistence.
 These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grade, exam grade, or future enrollment.
-In addition to the models in the original paper, this paper explores models using different controls and models that consider specific subgroups of the study participants.
-
-TODO: expand on my own findings.
-
-TODO: clearly outline the structure of the whole paper.
+Replicating exploratory models, the relationship between similarity and relationship, as well as between relationship and student outcomes, are modeled.
+Additional models use different controls; specifically, variables indicating shared racial and gender self-identification.
+As with the replication models, treatment has no significant impact on ISR perception or student outcomes.
 
@@
 # Methodology
-
-@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on perceived similarity, instructor-student relationship, course grade, and re-enrollment.
-
-## Participants
-
-The study took place in the 2017 spring semester at a large Californian University.
-The study included 120 instructors and their 2,749 students.
-The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized.Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors.
-TODO: what kind of sampling is this?
-What are the implications?
-
-## Treatment and control
-
-Participating students were randomly assigned to either treatment or control.
-At the beginning of the term, all participating students and instructors were given "get to know you" surveys.
-Using those responses, for each student in the treatment group, seven commonalities were identified between student and instructor (for example, perhaps both student and instructor binge-watch TV to relieve stress, or appreciate loyalty as the most important friend quality), and both student and instructor were informed of these commonalities.
-They completed a few questions about the similarities and were reminded of them through the semester to ensure they were internalized.
-Students in the control group were informed about similarities they shared with students in another part of the country, and instructors were told nothing about these students.
-
-## Procedures
-
-All students participated in a survey immediately following the treatment or the placebo. They were surveyed again at the end of the course. Instructors were surveyed only at the end of the course.
-
-## Measures
-
-@citeoriginal identify key measures. Full descriptions of all measures are available in [Appendix A].
-
-Some are extracted from the student survey at the beginning of the term:
-
-  1.
-Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
-These responses were averaged to create a student similarity perception scale.
-  2.
-Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create a student ISR perception scale.
-  3. Student gender
-
-Others are extracted from the student survey at the end of the term:
-
-  4. At the end of the semester, students answered the student similarity perception scale questions again.
-  5. At the end of the semester, students answered the student ISR perception scale questions again.
-
-Others come from the instructor survey at the end of the term:
-
-  6. Instructor similarity perception: At the end of the semester, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
-      - Overall, how similar do you think you and STUDENTNAME are?
-  7.
-Instructor ISR perception scale: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create an ISR scale.
-  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
-
-Finally, some are extracted from the university's internal records:
-
-  9. The final grade that the student received in the course.
-  10. The student's final grade, standardized against other grades in the course.
-  11. The student's cumulative GPA (CGPA) after the Fall 2016 term.
-  12. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
-
-TODO: evaluate scales' validity?
-
-# Data
-
-TODO: add all citations for R packages, R, and R markdown
-The correlation matrix was made using functions by @apacorrcode.
-
-TODO: note where the data is available and describe the dataset basics (cases, variables). Cite the dataset correctly, add to .bib
 
@@
 
+@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on perceived similarity, instructor-student relationship, course grade, and re-enrollment.
+
+## Participants
+
+The study took place in the 2017 spring semester at a large Californian University.
+The study included 120 instructors and their 2,749 students.
+The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized.
+Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors.
+The initial sample of 147 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements.
+145 instructors consented and took the initial survey, resulting in an corresponding sample of 3,352 students.
+Once errors of administration, missing responses, and inadequate time spent on surveys were considered, the sample consisted of only 119 instructors and 2,273 students.
+
+## Treatment and control
+
+Participating students were randomly assigned to either treatment or control.
+At the beginning of the term, all participating students and instructors were given "get to know you" surveys.
+Using those responses, for each student in the treatment group, seven commonalities were identified between student and instructor (for example, perhaps both student and instructor binge-watch TV to relieve stress, or appreciate loyalty as the most important friend quality), and both student and instructor were informed of these commonalities.
+They completed a few questions about the similarities and were reminded of them through the semester to ensure they were internalized.
+Students in the control group were informed about similarities they shared with students in another part of the country, and instructors were told nothing about these students.
+
+## Procedures
+
+All students participated in a survey immediately following the treatment or the placebo. They were surveyed again at the end of the course. Instructors were surveyed only at the end of the course.
+
+## Measures
+
+@citeoriginal identify key measures. Full descriptions of all measures are available in [Appendix A].
+
+Some are extracted from the student survey at the beginning of the term:
+
+  1.
+Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
+These responses were averaged to create a student similarity perception scale.
+  2.
+Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create a student ISR perception scale.
+  3. Student gender
+
+Others are extracted from the student survey at the end of the term:
+
+  4. At the end of the semester, students answered the student similarity perception scale questions again.
+  5.
+At the end of the semester, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create a student ISR perception scale.
+
+Others come from the instructor survey at the end of the term:
+
+  6. Instructor similarity perception: At the end of the semester, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
+      - Overall, how similar do you think you and STUDENTNAME are?
+  7.
+Instructor ISR perception scale: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create an ISR scale.
+  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
+
+Finally, some are extracted from the university's internal records:
+
+  9. The final grade that the student received in the course.
+  10. The student's final grade, standardized against other grades in the course.
+  11. The student's cumulative GPA (CGPA) after the Fall 2016 term.
+  12. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
+
+### Scale Reliability
+
+A scale is externally reliable if it gives the same results across different testings [@Coolican_2014].
+Test-retest reliability, where a scale is given to a group at different times, is a common way to assess this [@Coolican_2014].
+In our data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.
+
+A scale is internally reliable if it is consistent within itself [@Coolican_2014].
+Cronbach's alpha was used to assess the reliability of the similarity and ISR scales.
+This statistic evaluates how much participants vary on individual items, compared to how they vary overall [@Coolican_2014].
+Good reliability is indicatedby an alpha between .75 and 1 [@Coolican_2014].
+For the initial student similarity scale, initial student ISR scale, end-of-term student similarity scale, end-of-term student ISR scale, and end-of-term student ISR scale, the alpha values were 0.892, 0.872, 0.91, 0.905, and 0.926, respectively, indicating that the scales are reliable.
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+s1_sim_scale <- expdata %>%
+  dplyr::select(s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6) %>%
+  drop_na()
+cronbach.alpha(s1_sim_scale)
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+s1_tsr_scale <- expdata %>%
+  dplyr::select(s1_tsr1, s1_tsr2, s1_tsr3, s1_tsr4, s1_tsr5, s1_tsr6, s1_tsr7) %>%
+  drop_na()
+cronbach.alpha(s1_tsr_scale)
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+s2_sim_scale <- expdata %>%
+  dplyr::select(s2_sim1, s2_sim2, s2_sim3, s2_sim4, s2_sim5, s2_sim6) %>%
+  drop_na()
+cronbach.alpha(s2_sim_scale)
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+s2_tsr_scale <- expdata %>%
+  dplyr::select(s2_tsr1, s2_tsr2, s2_tsr3, s2_tsr4, s2_tsr5, s2_tsr6, s2_tsr7) %>%
+  drop_na()
+cronbach.alpha(s2_tsr_scale)
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+t2_tsr_scale <- expdata %>%
+  dplyr::select(t2_tsr1, t2_tsr2, t2_tsr3, t2_tsr4, t2_tsr5, t2_tsr6, t2_tsr7) %>%
+  drop_na()
+cronbach.alpha(t2_tsr_scale)
+```
+
+### Scale Validity
+
+A scale is valid if it measures what it is supposed to measure [@Coolican_2014].
+In this case, the scales should measure perceived similarity and perceived instructor-student relationship (ISR).
+@citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, where they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers.
+Given the fact that these scales were developed for children, it is worth considering what the scales actually measure.
+To do this, we can look at the specific questions in the three distinct scales.
+
+To assess similarity, students were asked about values, course goals, views on course content, general commonalities, personality, and overall similarity.
+By this standard, high similarity is when a student shares values, course goals, and views on course content, and the student and instructor have similar personalities, high commonality, and high general similarity.
+Questions ask about course-specific measures (content and goals), personal traits (values and personality), and non-specific traits (commonalities and general similarity), but do not ask about life experience, abilities, approaches, amoungst other ways that two people can theoretically be similar.
+
+TODO: ^ link to any other definitions/studies of similarity
+
+To assess ISR, students were asked about enjoyment of learning, friendliness, encouragement, excitement, motivation, caring, and overall learning.
+By this standard, a high-quality ISR is one where a student enjoys learning from, is motivated by, and learns a great deal from a friendly, encouraging, caring instructor, who would be excited to see them in three years time.
+Questions do not ask about respect, reliability, admiration, challenge, personal growth, frequency of contact, or many other aspects of how a student perceives an instructor-student relationship, especially at the college level.
+
+To assess ISR, instructors were asked about enjoyment of aiding learning, caring, frequency of encouragement, friendliness, excitement, motivation, and overall learning.
+By this standard, a high-quality ISR is one where an instructor enjoys helping and frequently encourages a friendly, caring student who is motivated and learns a great deal from this instructor, and the instructor would be excited to see the student in three years.
+Questions do not ask about respectfulness, interest, potential, frequency of contact, tone of that contact, or other aspects of instructor perception of an instructor-student relationship.
+
+TODO: ^^ link to any other definitions of ISR
+
+# Data
+
+TODO: add all citations for R packages, R, and R markdown
+The correlation matrix was made using functions by @apacorrcode.
+
+The data, in a .dta format, and Stata code for the @citeoriginal paper are available on the Inter-University Consortium for Political and Social Research [@Robinson_Scott_Gottfried_2020].
+The dataset contains 36838 observations of 653 variables.
+
 ## Demographics
@@
 Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
-
-Table \@ref(tab:demotable2) shows the instructor covariates.
-Because instructors are counted multiple times, for each student in the treatment and control groups that is in their course, only the totals are shown.
-The same covariates, presented by treatment and control groups, are shown in [Appendix B].
-
-There are differences between the students and instructors.
-Notably, the student sample is 60.3% female, while the instructor sample is 77.5% female.
-The student sample is 18.6% White, while the instructor sample is 39.2% White.
-The student sample is 43.5% Hispanic, while the instructor sample is only 3.3%.
-Finally, the student sample is 9.1% multiracial, but the instructor sample is 51.7% multiracial.
-Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classed as multiracial.
-For this reason, it is very difficult to draw conclusions.
-
-Table \@ref(tab:demotable2) shows the continuous covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
 
@@
 
+Table \@ref(tab:demotable2) shows the instructor covariates.
+Because instructors are counted multiple times, for each student in the treatment and control groups that is in their course, only the totals are shown.
+The same covariates, presented by treatment and control groups, are shown in [Appendix B].
+
 ```{r demotable2, echo=FALSE, warning=FALSE, message=FALSE}
@@
 ```
+
+There are differences between the students and instructors.
+Notably, the student sample is 62.7% female, while the instructor sample is 78.2% female.
+The student sample is 19.8% White, while the instructor sample is 38.7% White.
+The student sample is 47% Hispanic, while the instructor sample is only 3.4%.
+Finally, the student sample is 9.8% multiracial, but the instructor sample is 52.1% multiracial.
+Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classed differently as a result.
+For this reason, it is very difficult to draw conclusions.
+Additionally, students are 43.5% first-generation college students, while only 25.2% of instructors are.
 
@@
 
-TODO: any patterns in missing data.
-
-TODO: did attrition follow any particular pattern? Did some types of students stop responding in between the first and second surveys? Non-response.
+There are no significant patterns in missing data. [Appendix C] shows the distributions of key variables relative to missing data in other key variables.
+
+## Attrition
+
+129 eligible instructors completed the first survey, but 6 of those instructors did not complete the second survey.
+Corresponding to the remaining 123 instructors, 2,801 students completed the first survey (1,392 in the treatment group, 1,409 in the control group), but 682 of those students did not complete the second survey (328 in the treatment group, 354 in the control group).
+The attrition in the treatment and control groups was comparable (24% in the treatment group, 25% in the control group).
 
@@
 
-TODO: what data was excluded and why?
+From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis.
+Units of study were excluded because the instructor did not participate in the study (33,486); the student did not consent (123); the student did not spend more than a second reading the consent page (1); the initial student survey, initial instructor survey, or end-of-term instructor survey was not complete (428); the instructor mistakenly administered the wrong survey (30); the course was online (20); the course was for graduate students (50); or a participant did not spend more than ten seconds reading and answering a page with five or more questions (427).
+The ten-second time limit is somewhat arbitrary, and a different choice could lead to different study results.
 
@@
 
-TODO: why did @citeoriginal pick these? What do I think about the selection?
-
 $treatment_{i}$ is the indicator that treatment was given.
 
-$X_{1i}$ is a vector of student-level covariates (pre-intervention measures included).
-
-$X_{2j}$ is a vector of instructor-level covariates.
+$X_{1i}$ is a vector of student-level covariates (anticipated student ISR, gender, CGPA).
+Anticipated ISR is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes.
+@citeoriginal state this is because females earn higher grades generally.
+
+$X_{2j}$ is a vector of instructor-level covariates (course size, teacher ID).
 
@@
 
-Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes, which include complete scale outcomes (immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), and end of semester instructor ISR rating (t2_tsr)) and grade-based outcomes, measured in GPA (course grade (grade) and objectively graded exam grade (t2_finalexam)).
+Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes.
+This include complete scale outcomes: immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), and end of semester instructor ISR rating (t2_tsr).
+Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, they can be treated as continuous variables and modeled using linear functions.
+It also includes grade-based outcomes: course grade (grade) and objectively graded exam grade (t2_finalexam), both of which are shown on a 4.0 GPA scale.
 
@@
 
-TODO: additional models.
-Consider: More covariates (race, first generation, matching gender and race); Models looking at subgroups of the sample (Only those with matching traits, only small classes).
-Focus on grade as the outcome.
+When creating additional models, I focused on three key outcomes: end-of-term student similarity perception scale, end-of-term student ISR perception scale, and standardized course grade.
+I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance) and perceptions that could improve student outcomes (like student ISR perception), and this particular intervention tries to accomplish these goals by improving perceived similarity.
+
+\begin{equation}
+(\#eq:mylinmodel)
+Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + X_{3jk}\Gamma_{3} + \epsilon_{ij}
+\end{equation}
+
+Equation \@ref(eq:mylinmodel) shows the linear model used. $X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).
 
@@
 # Make all replication models
-studentfirstsimilaritymodel <- lm(data = expdata %>%
-                                    filter(!is.na(s1_sim)),
-                                  s1_sim ~ treatment + n_course + teacherid)
-studentsecondsimilaritymodel <- lm(data = expdata %>% 
-                                     filter(!is.na(s2_sim)),
-                                   s2_sim ~ treatment + n_course + teacherid)
-studentsecondrelationshipmodel <- lm(data = expdata %>%
-                                       filter(!is.na(s2_tsr) & !is.na(s1_tsr)),
-                                     s2_tsr ~ treatment + s1_tsr + n_course + teacherid)
-teachersecondrelationshipmodel <- lm(data = expdata %>%
-                                       filter(!is.na(t2_tsr)),
-                                     t2_tsr ~ treatment + n_course + teacherid)
-grademodel <- lm(data = expdata %>%
-                   filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(grade) & s1_female > -99),
-                 grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid)
-examgrademodel <- lm(data = expdata %>%
-                       filter(obj_exam == 1) %>%
-                       filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(t2_finalexam) & s1_female > -99),
-                     t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid)
+studentfirstsimilaritymodel <- lm(data = expdata,
+                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondsimilaritymodel <- lm(data = expdata,
+                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodel <- lm(data = expdata,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel <- lm(data = expdata,
+                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+grademodel <- lm(data = expdata,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
 teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
-                                data = expdata %>%
-                                  filter(!is.na(t2_sim1)),
-                                Hess = TRUE)
-enrollmentfallmodel <- glm(data = expdata %>%
-                             filter(!is.na(s1_female) & !is.na(ir_f16_gpa) & !is.na(f17_enrolled) & s1_female > -99),
-                           f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
-                           family = "binomial")
+                                data = expdata,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+enrollmentfallmodel <- glm(data = expdata,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           family = "binomial",
+                                     na.action = na.omit)
 ```
@@
           enrollmentfallmodel,
-          title = "Model results",
+          title = "Replication model results",
           align = TRUE,
@@
 
-TODO: reread paper & Stata code and ensure I caught all models that they report.
-
 In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and student initial similarity perception.
@@
 
-TODO: any problems with the models. General performance.
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+studentsecondrelationshipmodel_ <- lm(data = expdata,
+                                     formula = s2_tsr ~ treatment + s2_sim + t2_sim1 + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel_ <- lm(data = expdata,
+                                     formula = t2_tsr ~ treatment + s2_sim + t2_sim1 + n_course + teacherid,
+                                     na.action = na.omit)
+grademodel_ <- lm(data = expdata,
+                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                                     na.action = na.omit)
+examgrademodel_ <- lm(data = expdata %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                                     na.action = na.omit)
+enrollmentfallmodel_ <- glm(data = expdata,
+                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+```{r noncausaltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentsecondrelationshipmodel_,
+          teachersecondrelationshipmodel_,
+          grademodel_,
+          examgrademodel_,
+          enrollmentfallmodel_,
+          title = "Replication model results 2",
+          align = TRUE,
+          #float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:noncausaltable",
+          type = "latex",
+          header = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Teacher fixed effect"),
+          dep.var.labels = c("Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Persistence"),
+          covariate.labels = c("Treatment",
+                             "Student sim. 2",
+                             "Instructor sim. 2",
+                             "Course size",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Student female",
+                             "CGPA"))
+```
+
+Table \@ref(tab:noncausaltable) considers the relationship between perceived similarity and perceived ISR and the relationship between perceived ISR and student outcomes.
+Because only the treatment was randomized, we cannot make causal inferences.
+
+In Models 1 and 2, we can see that student and instructor perceived similarity are both significantly related to student and instructor perceived ISR.
+Unsurprisingly, student perceived similarity is a better predictor of student perceived ISR, and instructor perceived similarity is a better predictor of instructor perceived ISR.
+Additionally, course size is negatively related to student ISR perception.
+
+Models 3 and 4 show that student ISR perception and instructor ISR perception are both significantly related to course grade and final grade.
+Instructor perception is a stronger predictor than student perception.
+
+Model 5 indicates that ISR perception is not significantly related to persistence.
 
@@
 
-TODO: additional results
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+expdata <- expdata %>%
+  mutate(race_match = ifelse(t1_race_1 == s1_race_1
+                             & t1_race_2 == s1_race_2 
+                             & t1_race_3 == s1_race_3 
+                             & t1_race_4 == s1_race_4 
+                             & t1_race_5 == s1_race_5 
+                             & t1_race_6 == s1_race_6 
+                             & t1_race_7 == s1_race_7,
+                             1,
+                             0)) %>%
+  mutate(gend_match = ifelse(t_female == s_female,
+                             1,
+                             0)) %>%
+  mutate(age_dif = t_age - s1_age)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+mystudentsecondsimilaritymodel <- lm(data = expdata,
+                                     formula = s2_sim ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+mystudentsecondrelationshipmodel <- lm(data = expdata,
+                                     formula = s2_tsr ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+mystdgrademodel <- lm(data = expdata,
+                      formula = grade ~ treatment + ir_f16_gpa + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+```
+
+```{r mymodeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(mystudentsecondsimilaritymodel,
+          mystudentsecondrelationshipmodel,
+          mystdgrademodel,
+          title = "Additional model results",
+          align = TRUE,
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:mymodeltable",
+          type = "latex",
+          header = FALSE,
+          dep.var.labels = c("Student sim. 2",
+                             "Student ISR 2",
+                             "Std. course grade"),
+          covariate.labels = c("Treatment",
+                               "CGPA", 
+                               "Course size",
+                               "Matching racial self-ID", "Matching gender self-ID", "Age difference")
+          )
+```
+
+Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
+As with the replication results, there is a statistically significant relationship between treatment and student initial similarity perception.
+However, the expected increase is only 0.108 on a scale of 1 to 5, which is even lower than in the replication models.
+There is still no significant relationship between treatment and student end-of-semester ISR perception.
+Additionally, there is no significant relationship between treatment and standardized course grade.
 
@@
 # Discussion
+
+TODO: why did @citeoriginal pick their models? What do I think about the selection?
 
@@
   5. Student ISR perception scale (s2_tsr)
+      - How much do you enjoy learning from this professor?
+      - How friendly do you think this professor is towards you?
+      - If you came back to visit this professor three years from now, how excited do you think they would be?
+      - How motivating do you find this professor's class?
+      - How caring do you think this professor is towards you?
+      - How encouraging do you think this professor is towards you?
+      - Overall, how much do you think you have learned from this professor?
 * End of term instructor survey:
@@
   column_spec(7, width = "5em") %>%
-  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped", "hold_position"),
                             stripe_index =c(2:3, 13:14, 16))
@@
 
+# Appendix C
+
+```{r missingdata, echo=FALSE, warning=FALSE, message=FALSE, fig.height=9, fig.width=9}
+explanatory = c("s2_tsr", "s1_tsr", "ir_f16_gpa", "n_course", "s1_female")
+dependent = "treatment"
+
+expdata %>%
+  dplyr::select(treatment, 
+                s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
+                grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, 
+                s1_female, n_course) %>%
+  missing_pairs(dependent, explanatory)
+```
+
 \newpage

### 2021-04-19 outputs/exploratory.Rmd
Complete lit review section, scale validity section. Refine graphs and tables. Expand outcomes of interest section. Add coefficents and CIs to in-text model results. Add replications for first gen only, H/B only, impact of race on ISR, and relationship between similarity, ISR, and outcomes.

@@
 ```
+
+```{r s1simcorrelation, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6) %>%
+  summary_factorlist("treatment", 
+                     c("s1_sim1", "s1_sim2", "s1_sim3", "s1_sim4", "s1_sim5", "s1_sim6"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for initial student similarity scale",
+        booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r s1tsrcorrelation, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                s1_tsr1, s1_tsr2, s1_tsr3, s1_tsr4, s1_tsr5, s1_tsr6, s1_tsr7) %>%
+  summary_factorlist("treatment", 
+                     c("s1_tsr1", "s1_tsr2", "s1_tsr3", "s1_tsr4", "s1_tsr5", "s1_tsr6", "s1_tsr7"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(s1_tsr1, s1_tsr2, s1_tsr3, s1_tsr4, s1_tsr5, s1_tsr6, s1_tsr7))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for initial student similarity scale",
+        booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r s2simcorrelation, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                s2_sim1, s2_sim2, s2_sim3, s2_sim4, s2_sim5, s2_sim6) %>%
+  summary_factorlist("treatment", 
+                     c("s2_sim1", "s2_sim2", "s2_sim3", "s2_sim4", "s2_sim5", "s2_sim6"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(s2_sim1, s2_sim2, s2_sim3, s2_sim4, s2_sim5, s2_sim6))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for initial student similarity scale",
+        booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r s2tsrcorrelation, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                s2_tsr1, s2_tsr2, s2_tsr3, s2_tsr4, s2_tsr5, s2_tsr6, s2_tsr7) %>%
+  summary_factorlist("treatment", 
+                     c("s2_tsr1", "s2_tsr2", "s2_tsr3", "s2_tsr4", "s2_tsr5", "s2_tsr6", "s2_tsr7"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(s2_tsr1, s2_tsr2, s2_tsr3, s2_tsr4, s2_tsr5, s2_tsr6, s2_tsr7))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for initial student similarity scale",
+        booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r t2tsrcorrelation, echo=FALSE, warning=FALSE, message=FALSE}
+# Make the summary stats part of the table
+summary <- expdata %>% 
+  dplyr::select(treatment, 
+                t2_tsr1, t2_tsr2, t2_tsr3, t2_tsr4, t2_tsr5, t2_tsr6, t2_tsr7) %>%
+  summary_factorlist("treatment", 
+                     c("t2_tsr1", "t2_tsr2", "t2_tsr3", "t2_tsr4", "t2_tsr5", "t2_tsr6", "t2_tsr7"),
+                     p = FALSE, 
+                     add_dependent_label = TRUE, 
+                     dependent_label_prefix = "",
+                     add_col_totals = FALSE,
+                     add_row_totals = TRUE,
+                     include_row_missing_col = TRUE,
+                     col_totals_rowname = "",
+                     total_col = TRUE,
+                     col_totals_prefix = "N (%) = ") %>%
+  dplyr::select("Total N", "Missing N", "Total")
+
+# Make the correlation matrix part of the table
+matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
+                  dplyr::select(t2_tsr1, t2_tsr2, t2_tsr3, t2_tsr4, t2_tsr5, t2_tsr6, t2_tsr7))))
+
+# Combine the summary and correlation tables
+table <- summary %>%
+  bind_cols(matrix) %>%
+  rename("Mean (SD)" = Total, N = "Total N", Missing = "Missing N")
+
+# Put in a table.
+table %>%
+  kable(caption = "Correlation matrix for initial student similarity scale",
+        booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+```
+
+```{r}
+expdata %>%
+  dplyr::select(treatment, 
+                s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
+                grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, 
+                s1_female, n_course) %>%
+  missing_pattern()
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+substudentsecondsimilaritymodel <- lm(data = expdata %>% filter(s_firstgen == 1),
+                                     formula = s2_sim ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+substudentsecondrelationshipmodel <- lm(data = expdata %>% filter(s_firstgen == 1),
+                                     formula = s2_tsr ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+substdgrademodel <- lm(data = expdata %>% filter(s_firstgen == 1),
+                      formula = grade ~ treatment + ir_f16_gpa + n_course + race_match + gend_match + age_dif,
+                                     na.action = na.omit)
+```
+
+```{r submodeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(substudentsecondsimilaritymodel,
+          substudentsecondrelationshipmodel,
+          substdgrademodel,
+          title = "Additional model results",
+          align = TRUE,
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltable",
+          type = "latex",
+          header = FALSE,
+          dep.var.labels = c("Student Similarity 2",
+                             "Student ISR 2",
+                             "Standardized course grade"),
+          covariate.labels = c("Treatment",
+                               "CGPA", 
+                               "Course size",
+                               "Matching racial self-ID", "Matching gender self-ID", "Age difference")
+          )
+```

### 2021-04-19 outputs/paper/paper.Rmd
Complete lit review section, scale validity section. Refine graphs and tables. Expand outcomes of interest section. Add coefficents and CIs to in-text model results. Add replications for first gen only, H/B only, impact of race on ISR, and relationship between similarity, ISR, and outcomes.

@@
 author: "Amy Farrow"
-date: "April 13th, 2021"
+date: "April 18th, 2021"
 header-includes:
@@
 
+TODO: one more paragraph for the introduction - quick summary of significance from lit review, also look at rubric.
+
+@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance.
+In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships (ISRs) and measures of student success.
+Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
+The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not.
+Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.
+
+This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors.
+First, methods are discussed, and the measures and scales created by @citeoriginal are evaluated.
+Second, different demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated.
+Third, @citeoriginal's models are replicated, and additional models are explored.
+Linear, logistic, and ordinal logistic models are replicated to predict outcomes including ISR perception, grade, and persistence.
+These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence.
+Replicating exploratory models, the relationship between similarity and relationship, as well as between relationship and student outcomes, are modeled.
+Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.
+As with the replication models, treatment has no significant impact on ISR perception or student outcomes.
+
+## Literature review
+
 College is perceived to be a meritocratic tool for social mobility and career success [@college].
@@
 
-@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance.
-In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships (ISRs) and measures of student success.
-Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
-The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not.
-Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.
-
-This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors.
-First, methods are discussed, and the measures and scales created by @citeoriginal are evaluated.
-Second, different demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated.
-Third, @citeoriginal's models are replicated, and additional models are explored.
-Linear, logistic, and ordinal logistic models are replicated to predict outcomes including perceived ISR, grade, and persistence.
-These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grade, exam grade, or future enrollment.
-Replicating exploratory models, the relationship between similarity and relationship, as well as between relationship and student outcomes, are modeled.
-Additional models use different controls; specifically, variables indicating shared racial and gender self-identification.
-As with the replication models, treatment has no significant impact on ISR perception or student outcomes.
-
-## Literature review
-
-TODO: expand this paragraph about interventions, preferably targeting similar ideas to the @citeoriginal paper.
-What has worked?
-Draw attention to results.
-Make sure to refer to any studies that @citeoriginal say are especially influential for their work.
-
-Many interventions to increase completion rates have been suggested and tested.
-For example, @citecasemanagement studied case managment techniques; @citerecruiting investigated the effect of high school recruitment; @citenearbies identify the importance of interpersonal relationships; @citecertificate argue that certificates awarded prior to degrees can ameliorate the 'college-completion crisis'.
-These are only a small sample of the interventions suggested.
-Implementing an effective program is difficult due to complicated causes of attrition, embedded social inequalities, and expenses.
-
-TODO: write a paragraph about research on why similarity builds relationships (similarity -> relationships)
-
-TODO: write a paragraph about research on why relationships matter for scholastic performance (relationships -> grades)
-@Abrami_Mizener_1985's study of college students found a .
+Many causes, and corresponding solutions, have been theorized for the 'college completion crisis'.
+In his examination of persistence research, @Reason_2009 considers student precollege characteristics (sociodemographic traits, academic preparation and performance, and dispositions), organizational factors (structural-demographic characteristics and organizational behavior dimensions), student peer environment (campus racial and academic climates), and individual student experiences (curricular, classroom, and out-of-class experiences).
+Such reviews indicate the sheer complexity involved.
+Implementing an effective intervention is difficult due to complicated causes of attrition, embedded social inequalities, and expenses [@Tinto_2006].
+
+Interventions targeting non-academic factors are popular.
+For example, using a randomized controlled trial, @citecasemanagement implemented a case management intervention, where social workers helped students with financial assistance, course selection, finding childcare, and accessing social services.
+The case management intervention significantly improved persistence rates [@citecasemanagement].
+First year programs that are separate from academic faculty are popular, but increasingly there is a focus on the classroom's importance for retention [@Tinto_2006].
+@Tinto_1997 evaluated the impact of a college program designed for communal learning, which aims to engage students more than traditional programs.
+They found that students in the communal program participated in more academic activities, reported more positive views of the college and their role in it, and persisted at higher rates than the control [@Tinto_1997].
+
+Especially at the elementary and secondary level, many interventions to improve student outcomes do so by targeting ISRs.
+@Roorda_Koomen_Spilt_Oort_2011 performed a meta-analysis of 99 studies on teacher-student relationships and student engagement and achievement.
+Effect sizes were larger for engagement than for achievement, and the effect sizes varied widely based on the specific measurements being used [@Roorda_Koomen_Spilt_Oort_2011].
+They found stronger effects of teacher-student relationships in the upper secondary level, as compared to the lower elementary level [@Roorda_Koomen_Spilt_Oort_2011].
+At the college level, @Tinto_2006 argues that the instructor is a key player in student retention.
+@Creasey_Jarvis_Gadke_2009 found that students who had connected, nonthreatening relationships with their instructors had more positive feelings about their own abilities and expectations for the course.
+
+ISRs are built on a range of social behaviors and experiences, but similarity's effect on ISRs is of particular interest for this study.
+This is a specific case of a wider field of study: how does similarity affect all types of relationships?
+In general, people are more likely to persist in building relationships with acquaintances that they share similarities with, including prejudices, behaviors, personality traits, attitudes, demographics, and activities [@Bahns_Crandall_Gillath_Preacher_2017].
+In mentoring relationships, perceived deep similarity contributes to information sharing behavior and positive reception, which then is positively associated with mentee adjustment [@Zheng_Zheng_Wu_Yao_Wang_2021].
+
+Other studies consider similarity and student performance, but do not directly address ISRs.
 One study of Taiwanese high school students explored the link between shared education and life values and student performance [@Lai_2015].
 While they found a significant, but small, positive relationship between life values and student performance on analytic tests, education values gave mixed results depending on the natures of the education value and the analytic test in question [@Lai_2015].
-
-TODO: write a paragraph about research on why relationships matter for persistence (relationships -> persistence)
+At the college level, @Abrami_Mizener_1985's study found a small correlation between students' perceiving instructors' attitudes to be similar to their own and the grades of those students, but the correlation became insignificant when instructor fixed effects were considered.
+In the college context, the link between similarity and grades is not well-established.
+
+Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, @citeoriginal was directly inspired by two previous works in middle and high schools.
+@Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships are associated with changes in student perceived self-efficacy, but not scholastic performance.
+Using a randomized controlled trial and surveys, @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015 found that when high school students were presented with similarities they shared with their teachers, they subsequently perceived greater similarity with their teachers.
+When teachers were presented with similarities, they perceived better relationships with those students, and those students subsequently performed better academically.
+For students, however, awareness of similarities did not effectively improve their perception of the student-teacher relationship.
 
@@
   mutate(across(where(is.numeric), ~na_if(., -99))) %>% # Some NAs are shown as -99. Replace with NA.
-  mutate(s_race = as.factor(s_race), # make nominal variables into factors.
-         t_race = as.factor(t_race),
-         teacherid = as.factor(teacherid),
+  mutate(teacherid = as.factor(teacherid),
          f17_enrolled = as.factor(f17_enrolled))
@@
   mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
-                         ifelse(t1_race_1 == 1, ifelse(t1_race_2 == 1 & t1_race_3 != 1 & t1_race_4 != 1 & t1_race_5 != 1 & t1_race_6 != 1,
-                                                       "Black and White",
-                                                       ifelse(t1_race_2 != 1 & t1_race_3 == 1 & t1_race_4 != 1 & t1_race_5 != 1 & t1_race_6 != 1,
-                                                              "White Hispanic", "Other multi")), 
-                                ifelse(t1_race_2 == 1 & t1_race_3 == 1, "Black Hispanic", "Other multi")), 
+                         "Mixed race", 
                          ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
@@
   mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
-                         ifelse(s1_race_1 == 1, ifelse(s1_race_2 == 1 & s1_race_3 != 1 & s1_race_4 != 1 & s1_race_5 != 1 & s1_race_6 != 1,
-                                                       "Black and White",
-                                                       ifelse(s1_race_2 != 1 & s1_race_3 == 1 & s1_race_4 != 1 & s1_race_5 != 1 & s1_race_6 != 1,
-                                                              "White Hispanic", "Other multi")), 
-                                ifelse(s1_race_2 == 1 & s1_race_3 == 1, "Black Hispanic", "Other multi")), 
+                         "Mixed race", 
                          ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
@@
                                                                    ifelse(s1_race_6, "Middle Eastern",
-                                                                          "Other")))))))))
-```
-
-@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on perceived similarity, instructor-student relationship, course grade, and re-enrollment.
+                                                                          "Other"))))))))) %>%
+  mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))
+
+expdata <- expdata %>%
+  mutate(t2_finalexamobj = ifelse(obj_exam == 1, t2_finalexam, NA))
+```
+
+@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship, course grade, and re-enrollment.
 
@@
 
-The study took place in the 2017 spring semester at a large Californian University.
+The study took place in the 2017 spring term at a large Californian University.
 The study included 120 instructors and their 2,749 students.
@@
 Using those responses, for each student in the treatment group, seven commonalities were identified between student and instructor (for example, perhaps both student and instructor binge-watch TV to relieve stress, or appreciate loyalty as the most important friend quality), and both student and instructor were informed of these commonalities.
-They completed a few questions about the similarities and were reminded of them through the semester to ensure they were internalized.
+They completed a few questions about the similarities and were reminded of them through the term to ensure they were internalized.
 Students in the control group were informed about similarities they shared with students in another part of the country, and instructors were told nothing about these students.
@@
 Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
-These responses were averaged to create a student similarity perception scale.
+These responses were averaged to create a student perceived similarity scale.
   2.
-Immediately after the treatment or placebo, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create a student ISR perception scale.
-  3. Student gender
+Immediately after the treatment or placebo, students answered seven questions about their anticipated instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create a student anticipated ISR scale.
 
@@
 
-  4. At the end of the semester, students answered the student similarity perception scale questions again.
-  5.
-At the end of the semester, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create a student ISR perception scale.
+  3. At the end of the term, students answered the student perceived similarity scale questions again.
+  4.
+At the end of the term, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create a student perceived ISR scale.
 
@@
 
-  6. Instructor similarity perception: At the end of the semester, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
-      - Overall, how similar do you think you and STUDENTNAME are?
-  7.
-Instructor ISR perception scale: At the end of the semester, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+  5. Instructor similarity perception: At the end of the term, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
+  6.
+Instructor perceived ISR scale: At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
 These responses were averaged to create an ISR scale.
-  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
+  7. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
 
@@
 
-  9. The final grade that the student received in the course.
-  10. The student's final grade, standardized against other grades in the course.
-  11. The student's cumulative GPA (CGPA) after the Fall 2016 term.
-  12. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
+  8. The grade that the student received in the course.
+  9. The student's course grade, standardized against other grades in the course.
+  10. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
 
@@
 This statistic evaluates how much participants vary on individual items, compared to how they vary overall [@Coolican_2014].
-Good reliability is indicatedby an alpha between .75 and 1 [@Coolican_2014].
-For the initial student similarity scale, initial student ISR scale, end-of-term student similarity scale, end-of-term student ISR scale, and end-of-term student ISR scale, the alpha values were 0.892, 0.872, 0.91, 0.905, and 0.926, respectively, indicating that the scales are reliable.
+Good reliability is indicated by an alpha between .75 and 1 [@Coolican_2014].
+For the initial student perceived similarity scale, student anticipated ISR scale, end-of-term student  perceived similarity scale, end-of-term student perceived ISR scale, and end-of-term instructor perceived ISR scale, $\alpha$ = 0.892, 0.872, 0.91, 0.905, and 0.926, respectively, indicating that the scales are reliable.
 
@@
 cronbach.alpha(s1_sim_scale)
-```
-
-```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
 s1_tsr_scale <- expdata %>%
@@
 cronbach.alpha(s1_tsr_scale)
-```
-
-```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
 s2_sim_scale <- expdata %>%
@@
 cronbach.alpha(s2_sim_scale)
-```
-
-```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
 s2_tsr_scale <- expdata %>%
@@
 cronbach.alpha(s2_tsr_scale)
-```
-
-```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
 t2_tsr_scale <- expdata %>%
@@
 A scale is valid if it measures what it is supposed to measure [@Coolican_2014].
-In this case, the scales should measure perceived similarity and perceived instructor-student relationship (ISR).
-@citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, where they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers.
+In this case, the scales should measure similarity perception and instructor-student relationship (ISR) perception.
+@citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers.
 Given the fact that these scales were developed for children, it is worth considering what the scales actually measure.
@@
 By this standard, high similarity is when a student shares values, course goals, and views on course content, and the student and instructor have similar personalities, high commonality, and high general similarity.
-Questions ask about course-specific measures (content and goals), personal traits (values and personality), and non-specific traits (commonalities and general similarity), but do not ask about life experience, abilities, approaches, amoungst other ways that two people can theoretically be similar.
-
-TODO: ^ link to any other definitions/studies of similarity
+Questions ask about course-specific measures (content and goals), personal traits (values and personality), and non-specific traits (commonalities and general similarity), but do not ask about life experience, abilities, or approaches, amoungst other ways that two people can theoretically be similar.
 
@@
 By this standard, a high-quality ISR is one where a student enjoys learning from, is motivated by, and learns a great deal from a friendly, encouraging, caring instructor, who would be excited to see them in three years time.
-Questions do not ask about respect, reliability, admiration, challenge, personal growth, frequency of contact, or many other aspects of how a student perceives an instructor-student relationship, especially at the college level.
-
 To assess ISR, instructors were asked about enjoyment of aiding learning, caring, frequency of encouragement, friendliness, excitement, motivation, and overall learning.
 By this standard, a high-quality ISR is one where an instructor enjoys helping and frequently encourages a friendly, caring student who is motivated and learns a great deal from this instructor, and the instructor would be excited to see the student in three years.
-Questions do not ask about respectfulness, interest, potential, frequency of contact, tone of that contact, or other aspects of instructor perception of an instructor-student relationship.
-
-TODO: ^^ link to any other definitions of ISR
+
+As @Brinkworth_McIntyre_Juraschek_Gehlbach_2018 argue, the instructor-student relationship is conceptualized, and therefore measured, in a variety of ways.
+Other constructions of ISR scales may measure negative aspects, respect, liking, responsiveness, engagement, trust, honesty, humour, interest, approachability, clarity, confidence, fairness, patience, and consistency [@Brinkworth_McIntyre_Juraschek_Gehlbach_2018].
+@citeoriginal's ISR scale is measuring only one concept of an ISR, and that particular concept needs to be considered when analyzing the results of this experiment.
+Like most research about ISRs, @citeoriginal's scale originates in elementary and secondary schools, and the concept of "good ISR" may not translate easily to a post-secondary context.
 
@@
 demotab[2:3,4] <- c("Male", "Female")
-demotab[16:17,4] <- c("No", "Yes")
+demotab[13:14,4] <- c("No", "Yes")
 
@@
 demotab2 <- expdata %>%
-  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
   dplyr::select(teacherid, t_age, t_firstgen, t_race, t_female, n_course) %>%
   unique() %>%
+  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
   mutate(t_age = ff_label(t_age, "Instructor age"), 
@@
 demotab2[2:3,4] <- c("Male", "Female")
-demotab2[13:14,4] <- c("No", "Yes")
+demotab2[10:11,4] <- c("No", "Yes")
 
@@
 
-In Figure \@ref(fig:racegraph), we can see that more students identified as Hispanic, while more instructors identified as White and multi-racial.
-
-```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.cap="Bar chart illustrating racial identification proprtions of participants"}
+In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories.
+More students identified as Hispanic, while more instructors identified as White and multi-racial.
+
+```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8,4), fig.cap="Bar chart illustrating racial identification proprtions of participants"}
 instructorrace <- expdata %>%
@@
   unique() %>%
-  dplyr::summarize(white = sum(t1_race_1, na.rm = TRUE)/n(),
-            black = sum(t1_race_2, na.rm = TRUE)/n(),
-            hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
-            asian = sum(t1_race_4, na.rm = TRUE)/n(),
-            americanindian = sum(t1_race_5, na.rm = TRUE)/n(),
-            middleeastern = sum(t1_race_6, na.rm = TRUE)/n(),
-            other = sum(t1_race_7, na.rm = TRUE)/n(),
-            mixed = sum(t_multi_race > 1, na.rm = TRUE)/n())
+  dplyr::summarize(White = sum(t1_race_1, na.rm = TRUE)/n(),
+            Black = sum(t1_race_2, na.rm = TRUE)/n(),
+            Hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
+            Asian = sum(t1_race_4, na.rm = TRUE)/n(),
+            "American Indian" = sum(t1_race_5, na.rm = TRUE)/n(),
+            "Middle Eastern" = sum(t1_race_6, na.rm = TRUE)/n(),
+            Other = sum(t1_race_7, na.rm = TRUE)/n(),
+            Multiple = sum(t_multi_race > 1, na.rm = TRUE)/n())
 studentrace <- expdata %>%
@@
   unique() %>%
-  dplyr::summarize(white = sum(s1_race_1, na.rm = TRUE)/n(),
-            black = sum(s1_race_2, na.rm = TRUE)/n(),
-            hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
-            asian = sum(s1_race_4, na.rm = TRUE)/n(),
-            americanindian = sum(s1_race_5, na.rm = TRUE)/n(),
-            middleeastern = sum(s1_race_6, na.rm = TRUE)/n(),
-            other = sum(s1_race_7, na.rm = TRUE)/n(),
-            mixed = sum(s_multi_race > 1, na.rm = TRUE)/n())
+  dplyr::summarize(White = sum(s1_race_1, na.rm = TRUE)/n(),
+            Black = sum(s1_race_2, na.rm = TRUE)/n(),
+            Hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
+            Asian = sum(s1_race_4, na.rm = TRUE)/n(),
+            "American Indian" = sum(s1_race_5, na.rm = TRUE)/n(),
+            "Middle Eastern" = sum(s1_race_6, na.rm = TRUE)/n(),
+            Other = sum(s1_race_7, na.rm = TRUE)/n(),
+            Multiple = sum(s_multi_race > 1, na.rm = TRUE)/n())
 
@@
   pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
+  mutate(nms...1 = fct_relevel(nms...1, 
+                               c("Hispanic", "White", "Asian",  "Middle Eastern",
+                                 "Black", "American Indian", "Other", "Multiple")),
+         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
   ggplot() +
   geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
-  scale_x_discrete(labels = c("American Indian", "Asian", "Black", "Hispanic", "Middle Eastern", "Multiple", "Other", "White")) +
-  scale_fill_discrete(labels = c("Instructors", "Students")) +
-  labs(x = "Self-identified race", y = "Proportion of sample", fill = "Group", title = "Self-identified racial proportions different for instructors and students") +
+  scale_fill_discrete(labels = c("Students", "Instructors")) +
+  labs(x = "Self-identified race", y = "Proportion of sample", 
+       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
   theme_minimal()
@@
 
-In Figure \@ref(fig:classsizegraph), we can see that most class sizes are between 25 and 50 students, but there are some that are much larger.
-
-```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height = 3, fig.cap="Histogram illustrating class sizes"}
+Given the potential significance of class size for instructor-student relationships, Figure \@ref(fig:classsizegraph) shows the distribution of class sizes of instructors that participated in the study.
+Most class sizes are between 25 and 50 students, but there are some that are much larger.
+
+```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(5,3), fig.cap="Histogram illustrating class sizes", out.width="60%", fig.align="center"}
 expdata %>%
@@
 Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values.
+There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception.
+Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 and 3.5 versus 3.6).
 
@@
   mutate(t2_sim1 = as.factor(t2_sim1)) %>%
-  mutate(treatment = ff_label(treatment, "Group"), 
-         s1_sim = ff_label(s1_sim, "Initial student similarity perception scale"), 
-         s1_tsr = ff_label(s1_tsr, "Initial student ISR perception scale"), 
-         s2_sim = ff_label(s2_sim, "End-of-term student similarity perception scale"), 
-         s2_tsr = ff_label(s2_tsr, "End-of-term student ISR perception scale"), 
-         t2_sim1 = ff_label(t2_sim1, "End-of-term instructor similarity perception"), 
-         t2_tsr = ff_label(t2_tsr, "End-of-term instructor ISR perception scale"), 
-         grade = ff_label(grade, "Course grade"), 
-         std_grade = ff_label(std_grade, "Standardized course grade"), 
-         t2_finalexam = ff_label(t2_finalexam, "Final grade"), 
-         f17_enrolled = ff_label(f17_enrolled, "Persistence")) %>%
   summary_factorlist("treatment",
@@
                        "grade", "std_grade", "t2_finalexam", "f17_enrolled"),
-                     p = FALSE, 
+                     p = TRUE, 
                      add_dependent_label = TRUE, 
@@
                booktabs = TRUE, linesep = "",
-               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total")) %>%
+               col.names = c("", "N", "Missing", "", "Control", "Treatment", "Total", "p")) %>%
   pack_rows("Initial student perception", 2, 3) %>%
@@
 Figure \@ref(fig:continuousoutcomes) shows key continuous outcome distributions for both treatment and control.
-The only visible differences are in the initial student similarity perception and end-of-term student similarity perception.
+Corresponding to Table \@ref(tab:summarytable), the only visible differences in distributions are in the initial student similarity perception and end-of-term student similarity perception.
 
@@
   geom_histogram(position = "dodge") +
-  labs(x = "Initial student similarity", y = "Frequency", fill = "Group") +
+  labs(x = "Initial student similarity perception", y = "Frequency", fill = "Group") +
   theme_minimal()
@@
   geom_histogram(position = "dodge") +
-  labs(x = "Initial student ISR", y = "Frequency") +
+  labs(x = "Student ISR anticipation", y = "Frequency") +
   theme_minimal()
@@
   geom_histogram(position = "dodge") +
-  labs(x = "End-of-term student similarity", y = "Frequency") +
+  labs(x = "End-of-term student similarity perception", y = "Frequency") +
   theme_minimal()
@@
   geom_histogram(position = "dodge") +
-  labs(x = "End-of-term student ISR", y = "Frequency") +
+  labs(x = "End-of-term student ISR perception", y = "Frequency") +
   theme_minimal()
@@
   geom_histogram(position = "dodge") +
-  labs(x = "End-of-term instructor similarity", y = "Frequency") +
+  labs(x = "End-of-term instructor similarity perception", y = "Frequency") +
   theme_minimal()
@@
   geom_histogram(position = "dodge") +
-  labs(x = "End-of-term instructor ISR", y = "Frequency") +
+  labs(x = "End-of-term instructor ISR perception", y = "Frequency") +
   theme_minimal()
@@
 
-Replicating @citeoriginal's study, Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as significant.
-Some measures, unsurprisingly, are highly correlated: grade and standardized grade, and grade and final exam grade.
-Others are moderately correlated: student initial perceptions of similarity and instructor-student relationship, student end-of-term perceptions of similarity and instructor-student relationship, and instructor end-of-term perceptions of similarity and instructor-student relationship.
-These last three pairs could indicate that the similarity perception scales and instructor-student relationship perception scales may not be measuring distinct concepts, or that feelings of similarity and positivity in relationships are strongly associated in the classroom.
+Replicating @citeoriginal's study, Table \@ref(tab:correlationtable) displays a correlation matrix for the measures that @citeoriginal identified as key.
+The very strong correlations (course grade and the  standardized version, course grade and final grade, and course grade and objectively graded final grade) and some of the strong correlations (standardized course grade with final grade and objectively grade final grade) are entirely expected.
+
+Other strong correlations are more interesting:  initial student similarity and instructor-student relationship perceptions, end-of-term student similarity and instructor-student relationship perceptions, and end-of-term instructor similarity and instructor-student relationship perceptions.
+At a given moment in time, a participant's assessment of similarity and ISR are highly related.
+This could indicate that the two perception scales may not be measuring psychologically distinct concepts, or that feelings of similarity and good relationships are strongly associated in the classroom.
+
+Moderate correlations exist between initial and end-of-term student similarity perceptions as well as initial and end-of-term student ISR perceptions, indicating that student perceptions change somewhat but not entirely through the term.
+There are some moderate correlations between instructor perceptions of similarity or ISR and course grade, final grade, and objectively graded final grade, perhaps indicating that instructor opinions of students are linked to student performance in some way.
 
@@
                 s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, 
-                grade, std_grade, t2_finalexam, ir_f16_gpa, f17_enrolled, 
-                s1_female, n_course) %>%
-  mutate(treatment = ff_label(treatment, "Group"), 
-         s1_sim = ff_label(s1_sim, "Initial student similarity perception scale"), 
-         s1_tsr = ff_label(s1_tsr, "Initial student ISR perception scale"), 
-         s2_sim = ff_label(s2_sim, "End-of-term student similarity perception scale"), 
-         s2_tsr = ff_label(s2_tsr, "End-of-term student ISR perception scale"), 
-         t2_sim1 = ff_label(t2_sim1, "End-of-term instructor similarity perception"), 
-         t2_tsr = ff_label(t2_tsr, "End-of-term instructor ISR perception scale"), 
-         grade = ff_label(grade, "Course grade"), 
-         std_grade = ff_label(std_grade, "Standardized course grade"), 
-         t2_finalexam = ff_label(t2_finalexam, "Final grade"),
-         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
-         f17_enrolled = ff_label(f17_enrolled, "Persistence"),
-         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
-         n_course = ff_label(n_course, "Course size")) %>%
+                grade, std_grade, t2_finalexam, t2_finalexamobj) %>%
   summary_factorlist("treatment", 
-                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr", "grade",
-                       "std_grade", "t2_finalexam", "ir_f16_gpa", "n_course"),
+                     c("s1_sim", "s1_tsr", "s2_sim", "s2_tsr", "t2_sim1", "t2_tsr",
+                       "grade", "std_grade", "t2_finalexam", "t2_finalexamobj"),
                      p = FALSE, 
@@
 matrix <- data.frame(apaCorr(as.matrix(expdata %>% 
-                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr, grade, std_grade, t2_finalexam, ir_f16_gpa, n_course))))
+                  dplyr::select(s1_sim, s1_tsr, s2_sim, s2_tsr, t2_sim1, t2_tsr,
+                                grade, std_grade, t2_finalexam, t2_finalexamobj))))
 
@@
 # rename the variables for table readability
-rownames(table) <- c("Similarity", "ISR", " Similarity", " ISR", "Similarity ", "ISR ", "Course", "Stand. course", "Final", "CGPA", "Size")
+rownames(table) <- c("Similarity", "ISR", " Similarity", " ISR", "Similarity ", "ISR ", "Course", "Stand. course", "Final", "Obj. Final")
 
 # Put in a table.
-table %>%
+table[,1:11] %>%
   kable(caption = "Correlation matrix for continuous and ordinal key variables and outcomes",
@@
         col.names = c("N", "Missing", "Mean (SD)", "Similarity", "ISR", "Similarity","ISR",
-                      "Similarity", "ISR", "Course", "Stand. course", "Final", "CGPA", "Size")) %>%
-  column_spec(1, width = "8em") %>%
+                      "Similarity", "ISR", "Course", "Stand. course")) %>%
+  column_spec(1, width = "7em") %>%
   column_spec(4, width = "4em") %>%
@@
   column_spec(11, width = "4em") %>%
-  column_spec(12, width = "4em") %>%
-  column_spec(13, width = "2.5em") %>%
-  column_spec(14, width = "2.5em") %>%
   pack_rows("Initial student perception", 1, 2) %>%
@@
   pack_rows("Grades", 7, 10) %>%
-  pack_rows("Course", 11, 11) %>%
   add_header_above(c(" " = 4, "Initial student" = 2, "End-of-term student" = 2,
-                     "End-of-term instructor" = 2, "Grades" = 4, "Course" = 1), 
+                     "End-of-term instructor" = 2, "Grades" = 2), 
                    align = "l") %>%
@@
 
-$X_{1i}$ is a vector of student-level covariates (anticipated student ISR, gender, CGPA).
-Anticipated ISR is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes.
+$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA).
+ISR anticipation is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes.
 @citeoriginal state this is because females earn higher grades generally.
@@
 Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes.
-This include complete scale outcomes: immediate student similarity rating (s1_sim), end of semester student similarity rating (s2_sim), end of semester student ISR rating (s2_tsr), and end of semester instructor ISR rating (t2_tsr).
+This include complete scale outcomes: initial student similarity perception (s1_sim), end-of-term student similarity perception (s2_sim), end-of-term student ISR perception (s2_tsr), and end-of-term instructor ISR perception (t2_tsr).
 Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, they can be treated as continuous variables and modeled using linear functions.
@@
 
-Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end of semester instructor similarity rating (t2_sim1).
+Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception (t2_sim1).
 
@@
 
-When creating additional models, I focused on three key outcomes: end-of-term student similarity perception scale, end-of-term student ISR perception scale, and standardized course grade.
-I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance) and perceptions that could improve student outcomes (like student ISR perception), and this particular intervention tries to accomplish these goals by improving perceived similarity.
+When creating additional models, I focused on three key outcomes: end-of-term student perceived similarity scale, end-of-term student perceived ISR scale, and standardized course grade.
+I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance) and perceptions that could improve student outcomes (like student ISR perception), and this particular intervention tries to accomplish these goals by improving similarity perception.
+Equation \@ref(eq:mylinmodel) shows the linear model used.
+$X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).
 
@@
 
-Equation \@ref(eq:mylinmodel) shows the linear model used. $X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).
-
 # Results
@@
 ```
+
+In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.19, 95% CI[0.135,0.246]).
+Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.118, 95% CI[0.051,0.184]).
+However, both coefficients are minimal on a scale of 1 to 5.
+
+Model 3 shows no significant relationship between treatment and student end-of-term ISR perception.
+However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.565, 95% CI[0.519,0.611]).
+For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.565.
+
+Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-term ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent term.
 
@@
           align = TRUE,
+          ci = TRUE,
           float.env = "sidewaystable",
@@
 
-In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and student initial similarity perception.
-However, the expected increase is only 0.158 on a scale of 1 to 5.
-Model 2 shows a statistically significant relationship between treatment and student end-of-semester similarity perception.
-However, the expected increase is only 0.101 on a scale of 1 to 5.
-
-Model 3 shows no significant relationship between treatment and student end-of-semester ISR perception.
-However, it does show a significant correlation between anticipated student ISR and student end-of-semester ISR.
-For every 1 point increase in anticipated ISR on a 1 to 5 scale, the expected increase in end-of-semester ISR is 0.674.
-
-Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-semester ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent semester.
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+
+expdatahb <- expdata %>%
+  filter(s_hisp_black == 1)
+
+studentfirstsimilaritymodelhb <- lm(data = expdatahb,
+                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondsimilaritymodelhb <- lm(data = expdatahb,
+                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodelhb <- lm(data = expdatahb,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodelhb <- lm(data = expdatahb,
+                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+grademodelhb <- lm(data = expdatahb,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
+                                data = expdatahb,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+enrollmentfallmodelhb <- glm(data = expdatahb,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+Table \@ref(tab:modeltablehb) shows the same models as in Table \@ref(tab:modeltable), created only using the subset of Hispanic and/or Black college student participants.
+While the effect of treatment on student similarity perception is larger than for the general student population ($\beta$ = 0.21, 95% CI[0.137,0.282] for initial perception; $\beta$ = 0.148, 95% CI[0.06,0.237] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+
+```{r modeltablehb, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstsimilaritymodelhb,
+          studentsecondsimilaritymodelhb,
+          studentsecondrelationshipmodelhb,
+          teachersecondrelationshipmodelhb,
+          grademodelhb,
+          examgrademodelhb,
+          teachersecondsimilarityhb,
+          enrollmentfallmodelhb,
+          title = "Replication model results: Hispanic and Black students only",
+          align = TRUE,
+          ci = TRUE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltablehb",
+          type = "latex",
+          header = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Teacher fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Instructor sim. 2",
+                             "Persistence"),
+          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA", "Course size"))
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+
+expdata1g <- expdata %>%
+  filter(s_firstgen == 1)
+
+studentfirstsimilaritymodel1g <- lm(data = expdata1g,
+                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondsimilaritymodel1g <- lm(data = expdata1g,
+                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodel1g <- lm(data = expdata1g,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel1g <- lm(data = expdata1g,
+                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+grademodel1g <- lm(data = expdata1g,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
+                                data = expdata1g,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+enrollmentfallmodel1g <- glm(data = expdata1g,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+Table \@ref(tab:modeltable1g) shows the same models as in Table \@ref(tab:modeltable), created only using the subset of first-generation college student participants.
+While the effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.272, 95% CI[0.188,0.357] for initial perception; $\beta$ = 0.165, 95% CI[0.062,0.268] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+
+```{r modeltable1g, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstsimilaritymodel1g,
+          studentsecondsimilaritymodel1g,
+          studentsecondrelationshipmodel1g,
+          teachersecondrelationshipmodel1g,
+          grademodel1g,
+          examgrademodel1g,
+          teachersecondsimilarity1g,
+          enrollmentfallmodel1g,
+          title = "Replication model results: first-generation students only",
+          align = TRUE,
+          ci = TRUE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltable1g",
+          type = "latex",
+          header = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Teacher fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Instructor sim. 2",
+                             "Persistence"),
+          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA", "Course size"))
+```
+
+### Exploratory replication results
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+
+studentfirstisrmodel1g <- lm(data = expdata,
+                                  formula = s1_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodel1g <- lm(data = expdata,
+                                     formula = s2_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel1g <- lm(data = expdata,
+                                     formula = t2_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+studentfirstisrmodelhb <- lm(data = expdata,
+                                  formula = s1_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodelhb <- lm(data = expdata,
+                                     formula = s2_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodelhb <- lm(data = expdata,
+                                     formula = t2_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
+                                     na.action = na.omit)
+
+```
+
+Table \@ref(tab:modeltablerel) disregards the treatment condition, instead considering the relationship between student identity and similarity/ISR.
+While we cannot make causal inferences, we can explore relationships between the variables in this particular study.
+If a student identifies as first generation, or as Hispanic and/or Black, can we anticipate a different student or instructor ISR perception?
+There is a significant negative relationship between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.1, 95% CI[-0.158,-0.041]).
+
+```{r modeltablerel, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstisrmodel1g,
+          studentsecondrelationshipmodel1g,
+          teachersecondrelationshipmodel1g,
+          studentfirstisrmodelhb,
+          studentsecondrelationshipmodelhb,
+          teachersecondrelationshipmodelhb,
+          title = "Replication model results: first-generation status, Hispanic and/or Black students, and ISR",
+          align = TRUE,
+          ci = TRUE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltablerel",
+          type = "latex",
+          header = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Teacher fixed effect"),
+          dep.var.labels = c("Student ISR 1",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Student ISR 1",
+                             "Student ISR 2",
+                             "Instructor ISR 2"),
+          covariate.labels = c("Student first gen.", "Student Hispanic and/or Black", "CGPA", "Course grade"))
+```
 
@@
 ```
+
+Table \@ref(tab:noncausaltable) considers the relationship between similarity perception and ISR perception and the relationship between ISR perception and student outcomes.
+Again, because only the treatment was randomized, we cannot make causal inferences.
+
+In Models 1 and 2, we can see that student and instructor similarity perception are both significantly related to student and instructor ISR perception.
+
+Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.66, 95% CI[0.631,0.688]), and instructor similiarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.502, 95% CI[0.472,0.533]).
+Additionally, course size is negatively related to student ISR perception ($\beta$ = -0.102, 95% CI[-0.178,-0.026]).
+
+Models 3 and 4 show that student ISR perception and instructor ISR perception are both significantly related to course grade and final grade.
+Instructor perception is a stronger predictor ($\beta$ = 0.395, 95% CI[0.343,0.447] for course grade; $\beta$ = 0.432, 95% CI[0.299,0.565] for final grade) than student perception ($\beta$ = 0.076, 95% CI[0.023,0.129] for course grade; $\beta$ = 0.144, 95% CI[0.017,0.271] for final grade).
+
+Model 5 indicates that ISR perception is not significantly related to persistence.
 
@@
           enrollmentfallmodel_,
-          title = "Replication model results 2",
+          title = "Replication model results: Similarity, ISR, and student outcomes",
           align = TRUE,
+          ci = TRUE,
           #float.env = "sidewaystable",
@@
 
-Table \@ref(tab:noncausaltable) considers the relationship between perceived similarity and perceived ISR and the relationship between perceived ISR and student outcomes.
-Because only the treatment was randomized, we cannot make causal inferences.
-
-In Models 1 and 2, we can see that student and instructor perceived similarity are both significantly related to student and instructor perceived ISR.
-Unsurprisingly, student perceived similarity is a better predictor of student perceived ISR, and instructor perceived similarity is a better predictor of instructor perceived ISR.
-Additionally, course size is negatively related to student ISR perception.
-
-Models 3 and 4 show that student ISR perception and instructor ISR perception are both significantly related to course grade and final grade.
-Instructor perception is a stronger predictor than student perception.
-
-Model 5 indicates that ISR perception is not significantly related to persistence.
-
 TODO: clustered residuals (low priority)
@@
 
+Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
+As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception.
+However, the expected increase is only 0.107 on a scale of 1 to 5, which is even lower than in the replication models.
+There is still no significant relationship between treatment and student end-of-term ISR perception.
+Additionally, there is no significant relationship between treatment and standardized course grade.
+
 ```{r mymodeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
@@
           align = TRUE,
+          ci = TRUE,
           no.space = TRUE,
@@
 
-Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
-As with the replication results, there is a statistically significant relationship between treatment and student initial similarity perception.
-However, the expected increase is only 0.108 on a scale of 1 to 5, which is even lower than in the replication models.
-There is still no significant relationship between treatment and student end-of-semester ISR perception.
-Additionally, there is no significant relationship between treatment and standardized course grade.
-
 TODO: Shiny.
@@
 
+TODO: causal chain: intervention -> increased similarity perception -> improved ISR/ISR perception -> improved grades/persistence
+
+TODO: Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
+
+TODO: Who perceives similarity?
+
+TODO: What does it mean for relationships and therefore academic success to be built on similarity.
+
+## Limitations
+
 TODO: does this study have external validity?
 
-TODO: Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
-
-TODO: Who perceives similarity
-
-TODO: What does it mean for relationships and therefore academic success to be built on similarity.
+TODO: very long surveys pose limitations on accuracy
 
@@
 * Initial student survey:
-  1. Student similarity perception scale (s1_sim)
+  1. Student perceived similarity scale (s1_sim)
       - Overall, how similar to your instructor's values do you think your values are?
@@
       - Overall, how similar do you think you and your instructor are?
-  2. Student ISR perception scale (s1_tsr)
+  2. Student anticipated ISR scale (s1_tsr)
       - How much do you think you will enjoy learning from this instructor?
@@
       - Overall, how much do you think you will learn from this instructor?
-  3. Student gender (s1_female)
 * End of term student survey:
-  4. Student similarity perception scale (s2_sim)
-  5. Student ISR perception scale (s2_tsr)
+  3. Student perceived similarity scale (s2_sim)
+  4. Student perceived ISR scale (s2_tsr)
       - How much do you enjoy learning from this professor?
@@
 * End of term instructor survey:
-  6. Instructor similarity perception (t2_sim1):
+  5. Instructor similarity perception (t2_sim1):
       - Overall, how similar do you think you and STUDENTNAME are?
-  7. Instructor ISR perception scale (t2_tsr):
+  6. Instructor perceived ISR scale (t2_tsr):
       - How much did you enjoy helping STUDENTNAME learn?
@@
   10. Standardized course grade (std_grade): The student's final grade, standardized against other grades in the course.
-  11. CGPA (ir_f16_gpa): The student's cumulative GPA after the Fall 2016 term.
-  12. Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
+  11. Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
+
+\newpage
 
@@
 
+\newpage
+
 # Appendix C
@@
 # References
+
+TODO: look at the references that bookdown is generating and edit bibtex accordingly.

### 2021-04-20 outputs/exploratory.Rmd
first draft complete

@@
 ```
+
+```{r causaldiagram2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Diagram of desired causal sequence"}
+mermaid("
+graph LR
+A{intervention}-->|1|B[student similarity perception]
+A-->|2|C[instructor similarity perception]
+B-->|3|D[ISR]
+C-->|4|D[ISR]
+D-->|5|E(grades)
+D-->|6|F(persistence)
+E-->|7|F
+")
+```
+
+```{r}
+studentfirstsimilaritymodel <- lm(data = expdata,
+                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondsimilaritymodel <- lm(data = expdata,
+                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+studentsecondrelationshipmodel <- lm(data = expdata,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel <- lm(data = expdata,
+                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     na.action = na.omit)
+grademodel <- lm(data = expdata,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
+                                data = expdata,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+
+stargazer(studentfirstsimilaritymodel,
+          studentsecondsimilaritymodel,
+          studentsecondrelationshipmodel,
+          teachersecondrelationshipmodel,
+          grademodel,
+          examgrademodel,
+          teachersecondsimilarity,
+          omit = c("teacherid"),
+          omit.labels = c("Instructor fixed effect"))
+```

### 2021-04-20 outputs/paper/paper.Rmd
first draft complete

@@
 author: "Amy Farrow"
-date: "April 18th, 2021"
+date: "April 19th, 2021"
 header-includes:
@@
 While the scalability and affordability of the intervention are desirable, there are no results in any of the targeted measures: those that affect and reflect college retention.
-These results are consistent with the original paper.
-TODO: revise to clearliy indicate my own findings \\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
+These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
 thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
@@
 library(ltm) # for checking scale validity
+library(DiagrammeR) # for causal diagram
 
@@
 
+TODO: Shiny.
+Recreate Figure \@ref(fig:continuousoutcomes)(add graphs for standardized grade and persistence).
+Make an interactive element where you can filter the graph for a subset of the data -- specific racial s_race, firstgen/not, specific year of study.
+
 # Introduction
 
-TODO: one more paragraph for the introduction - quick summary of significance from lit review, also look at rubric.
+Despite the growing importance of college degrees for career success, college completion rates are a continuing problem.
+The causes are complex: factors that affect student persistence include student characteristics, organizational factors, peer environment, and student experiences.
+Interventions to improve student retention can target any of these aspects.
 
@@
 This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors.
-First, methods are discussed, and the measures and scales created by @citeoriginal are evaluated.
-Second, different demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated.
+First, @citeoriginal's methods are discussed, and the measures and scales are evaluated.
+Second, demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated.
 Third, @citeoriginal's models are replicated, and additional models are explored.
 Linear, logistic, and ordinal logistic models are replicated to predict outcomes including ISR perception, grade, and persistence.
+Further replicated exploratory models consider the relationship between similarity and instructor-student relationship, as well as between instructor-student relationship and student outcomes.
+Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.
+Finally,
+
 These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence.
-Replicating exploratory models, the relationship between similarity and relationship, as well as between relationship and student outcomes, are modeled.
-Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.
-As with the replication models, treatment has no significant impact on ISR perception or student outcomes.
+These conclusions stand when matching instructor-student traits are considered, and when vulnerable student groups are evaluated independent of the wider student population.
+The lack of significant effects reflects the complicated task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.
 
@@
 expdata <- expdata %>%
-  mutate(across(where(is.numeric), ~na_if(., -99))) %>% # Some NAs are shown as -99. Replace with NA.
+  mutate(s1_female = ifelse(s1_female == -99, NA, s1_female),
+         t_female = ifelse(t_female == -99, NA, t_female)) %>% # Some NAs are shown as -99. Replace with NA.
   mutate(teacherid = as.factor(teacherid),
-         f17_enrolled = as.factor(f17_enrolled))
+         f17_enrolled = as.factor(f17_enrolled),
+         s1_female = as.factor(s1_female),
+         t_female = as.factor(t_female))
 ```
@@
 
-TODO: add all citations for R packages, R, and R markdown
-The correlation matrix was made using functions by @apacorrcode.
+This paper uses the R statistical language [@citeR] and `tidyverse` packages [@citetidyverse].
+`here` was used to manage the RProject [@citehere].
+Stata files were handled using `readstata13` [@citereadstata13].
+The report was made using RMarkdown [@citermarkdown] and `bookdown` [@citebookdown].
+Tables were made using `kableExtra` [@citekableExtra], `finalfit` [@citefinalfit], and `stargazer` [@citestargazer].
+The correlation matrix was made using functions by @apacorrcode using the `Hmisc` package [@citeHmisc].
+`MASS` provided additional models [@citeMASS].
+Plots used `cowplot` [@citecowplot].
+`ltm` was used to check scale validity [@citeltm].
+Finally, the diagram was made using `DiagrammeR` [@citeDiagrammeR].
 
 The data, in a .dta format, and Stata code for the @citeoriginal paper are available on the Inter-University Consortium for Political and Social Research [@Robinson_Scott_Gottfried_2020].
-The dataset contains 36838 observations of 653 variables.
+The dataset contains 36,838 observations of 653 variables.
 
@@
   kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
-                            stripe_index =c(2:3, 16:17, 22))
+                            stripe_index =c(2:3, 13:14, 20),
+                            stripe_color = "#f6f0fb")
 
@@
   kableExtra::kable_styling(latex_options = c("striped"), 
-                            stripe_index =c(2:3, 13:14, 16))
+                            stripe_index =c(2:3, 10:11, 13),
+                            stripe_color = "#f6f0fb")
 ```
@@
   geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
-  scale_fill_discrete(labels = c("Students", "Instructors")) +
   labs(x = "Self-identified race", y = "Proportion of sample", 
        fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
+  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
   theme_minimal()
@@
   ggplot(aes(x = n_course)) +
-  geom_histogram(binwidth = 5) +
+  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
   labs(x = "Class size", y = "Frequency", title = "Some class sizes are very large") +
@@
   kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
-                            stripe_index =c(2,4,6,8,10,13:14))
+                            stripe_index =c(2,4,6:10,12,14),
+                            stripe_color = "#f6f0fb")
 ```
@@
   labs(x = "Initial student similarity perception", y = "Frequency", fill = "Group") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "Student ISR anticipation", y = "Frequency") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "End-of-term student similarity perception", y = "Frequency") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "End-of-term student ISR perception", y = "Frequency") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "End-of-term instructor similarity perception", y = "Frequency") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "End-of-term instructor ISR perception", y = "Frequency") +
+  scale_x_continuous(limits = c(.8,5.2)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "Course grade", y = "Frequency") +
+  scale_x_continuous(limits = c(-.1,4.5)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "Standardized course grade", y = "Frequency") +
+  scale_x_continuous(limits = c(-4.5,3)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
   labs(x = "Final grade", y = "Frequency") +
+  scale_x_continuous(limits = c(-.1,4.5)) +
+  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
   theme_minimal()
@@
 Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes.
-This include complete scale outcomes: initial student similarity perception (s1_sim), end-of-term student similarity perception (s2_sim), end-of-term student ISR perception (s2_tsr), and end-of-term instructor ISR perception (t2_tsr).
+This include complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception.
 Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, they can be treated as continuous variables and modeled using linear functions.
-It also includes grade-based outcomes: course grade (grade) and objectively graded exam grade (t2_finalexam), both of which are shown on a 4.0 GPA scale.
+It also includes grade-based outcomes: course grade and objectively graded exam grade, both of which are shown on a 4.0 GPA scale.
 
@@
 studentfirstsimilaritymodel <- lm(data = expdata,
-                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                  formula = s1_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondsimilaritymodel <- lm(data = expdata,
-                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                   formula = s2_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondrelationshipmodel <- lm(data = expdata,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
                                      na.action = na.omit)
 teachersecondrelationshipmodel <- lm(data = expdata,
-                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     formula = t2_tsr ~ treatment + teacherid,
                                      na.action = na.omit)
 grademodel <- lm(data = expdata,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
                                      na.action = na.omit)
 examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
                                 data = expdata,
@@
 enrollmentfallmodel <- glm(data = expdata,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
                            family = "binomial",
@@
 
-In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.19, 95% CI[0.135,0.246]).
+In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.189, 95% CI[0.134,0.245]).
 Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.118, 95% CI[0.051,0.184]).
@@
 Model 3 shows no significant relationship between treatment and student end-of-term ISR perception.
-However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.565, 95% CI[0.519,0.611]).
+However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.566, 95% CI[0.52,0.612]).
 For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.565.
@@
           no.space = TRUE,
-          omit.stat=c("ser","f"),
+          keep.stat = c("rsq", "adj.rsq", "aic", "lr", "wald", "ll", "logrank"),
+          #omit.stat=c("ser","f", "ll"),
           column.sep.width = "-10pt",
@@
           header = FALSE,
+          initial.zero = FALSE,
           omit = c("teacherid"),
-          omit.labels = c("Teacher fixed effect"),
+          omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student sim. 1",
@@
                              "Persistence"),
-          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA", "Course size"))
+          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
 ```
@@
 studentfirstsimilaritymodelhb <- lm(data = expdatahb,
-                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                  formula = s1_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondsimilaritymodelhb <- lm(data = expdatahb,
-                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                   formula = s2_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
                                      na.action = na.omit)
 teachersecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     formula = t2_tsr ~ treatment + teacherid,
                                      na.action = na.omit)
 grademodelhb <- lm(data = expdatahb,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
                                      na.action = na.omit)
 examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
                                 data = expdatahb,
@@
 enrollmentfallmodelhb <- glm(data = expdatahb,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
                            family = "binomial",
@@
           ci = TRUE,
+          initial.zero = FALSE,
           float.env = "sidewaystable",
@@
           omit = c("teacherid"),
-          omit.labels = c("Teacher fixed effect"),
+          omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student sim. 1",
@@
 studentfirstsimilaritymodel1g <- lm(data = expdata1g,
-                                  formula = s1_sim ~ treatment + n_course + teacherid,
+                                  formula = s1_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondsimilaritymodel1g <- lm(data = expdata1g,
-                                   formula = s2_sim ~ treatment + n_course + teacherid,
+                                   formula = s2_sim ~ treatment + teacherid,
                                      na.action = na.omit)
 studentsecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + teacherid,
+                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
                                      na.action = na.omit)
 teachersecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = t2_tsr ~ treatment + n_course + teacherid,
+                                     formula = t2_tsr ~ treatment + teacherid,
                                      na.action = na.omit)
 grademodel1g <- lm(data = expdata1g,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
                                      na.action = na.omit)
 examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + teacherid,
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
+                                     na.action = na.omit)
+teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
                                 data = expdata1g,
@@
 enrollmentfallmodel1g <- glm(data = expdata1g,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + n_course + teacherid,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
                            family = "binomial",
@@
           align = TRUE,
+          initial.zero = FALSE,
           ci = TRUE,
@@
           no.space = TRUE,
+          initial.zero = FALSE,
           omit.stat=c("ser","f"),
@@
           omit = c("teacherid"),
-          omit.labels = c("Teacher fixed effect"),
+          omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student ISR 1",
@@
 studentsecondrelationshipmodel_ <- lm(data = expdata,
-                                     formula = s2_tsr ~ treatment + s2_sim + t2_sim1 + n_course + teacherid,
+                                     formula = s2_tsr ~ treatment + s2_sim + t2_sim1 + n_course,
                                      na.action = na.omit)
 teachersecondrelationshipmodel_ <- lm(data = expdata,
-                                     formula = t2_tsr ~ treatment + s2_sim + t2_sim1 + n_course + teacherid,
+                                     formula = t2_tsr ~ treatment + s2_sim + t2_sim1 + n_course,
                                      na.action = na.omit)
 grademodel_ <- lm(data = expdata,
-                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                 formula = grade ~ teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                      na.action = na.omit)
 examgrademodel_ <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                     formula = t2_finalexam ~  teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                      na.action = na.omit)
 enrollmentfallmodel_ <- glm(data = expdata,
-                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + teacherid,
+                           formula = f17_enrolled ~  teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                            family = "binomial",
@@
 
-Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.66, 95% CI[0.631,0.688]), and instructor similiarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.502, 95% CI[0.472,0.533]).
+Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.733, 95% CI[0.706,0.76]), and instructor similarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.512, 95% CI[0.483,0.541]).
 Additionally, course size is negatively related to student ISR perception ($\beta$ = -0.102, 95% CI[-0.178,-0.026]).
@@
           ci = TRUE,
-          #float.env = "sidewaystable",
+          initial.zero = FALSE,
           no.space = TRUE,
@@
           omit = c("teacherid"),
-          omit.labels = c("Teacher fixed effect"),
+          omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student ISR 2",
@@
 Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
-As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception.
-However, the expected increase is only 0.107 on a scale of 1 to 5, which is even lower than in the replication models.
+As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.107, 95% CI[0.035,0.179]).
+However, the coefficient is even lower than in the replication models.
 There is still no significant relationship between treatment and student end-of-term ISR perception.
@@
 
-TODO: Shiny.
-
 # Discussion
 
-TODO: why did @citeoriginal pick their models? What do I think about the selection?
-
-TODO: causal chain: intervention -> increased similarity perception -> improved ISR/ISR perception -> improved grades/persistence
-
-TODO: Inefficacy of weak measures in the face of divergent student/instructor demographics and massive class sizes.
-
-TODO: Who perceives similarity?
-
-TODO: What does it mean for relationships and therefore academic success to be built on similarity.
+The results of this replication study align with @citeoriginal's findings.
+When undergraduate students and their instructors learn about similarities they share, there is a small improvement in student similarity perception.
+However, the treatment has no significant impact on student ISR perception, instructor perception, or student outcomes.
+These results stand when student gender, prior GPA, course size, matching student-instructor traits, and instructor fixed effects are controlled for.
+
+To understand what was and wasn't impacted by the treatment, we can consider the desired causal sequence in this study, shown in Figure \@ref(fig:causaldiagram).
+The intervention refers to informing students and instructors about shared traits.
+Similarity perception is measured by the two perceived similarity scales.
+ISR, or instructor-student relationship, cannot be directly measured -- only perception can be measured, through the perceived ISR scales.
+The student outcomes, grades and persistence, are factual and more easily measured.
+The diagram shows the complexity of succeeding with this type of intervention: the treatment must affect perception, which must then affect relationships, which must then affect student outcomes.
+
+```{r causaldiagram, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Diagram of desired causal sequence", dev='png', out.width="100%"}
+DiagrammeR::grViz("digraph {
+  graph [layout = dot, rankdir = LR]
+
+  node [shape = diamond, color = MediumSpringGreen, penwidth = 2]         
+  treatment [label = 'Intervention']
+
+  node [shape = rectangle, color = MediumSlateBlue]        
+  ssp [label = 'Student\nSimilarity\nPerception']
+  isp [label = 'Instructor\nSimilarity\nPerception']
+  isr [label =  'ISR']
+
+  node [shape = oval, color = HotPink]
+  grades [label = 'Grades']
+  persistence [label = 'Persistence']
+
+  # edge definitions with the node IDs
+  treatment -> ssp [ label = '1']
+  treatment -> isp [ label = '2']
+  ssp -> isr [ label = '3']
+  isp -> isr [ label = '4']
+  isr -> grades [ label = '5']
+  isr -> persistence [ label = '6']
+  grades -> persistence [ label = '7']
+  }",
+  height = 564,
+  width = 2700)
+```
+
+This study did successfully demonstrate relationship 1: the treatment caused a statistically significant increase in student similarity perception at the beginning and end of the term.
+Relationship 2 failed to materialize, as there was no significant difference in instructor similarity perception in the treatment group.
+To understand why the treatment effect was small for students and insignificant for instructors, we can consider the specifics of the treatment.
+What similarities were students and instructors informed of?
+[Appendix D] contains a complete list.
+Students and instructors were presented with generalities such as "An instructor’s most important characteristic?
+You both say being knowledgeable" or "You are cut from the same cloth–you both described yourself as thoughtful!" Such facts may not be unusual enough to be memorable, reducing the treatment's impact.
+It is also possible that similarities declared by an anonymous third party are simply not as impactful as similarities that are personally observed and experienced.
+
+The exploratory modeling in Table \@ref(tab:noncausaltable) indicates that relationships 3 and 4 may exist: increases in similarity perception were associated with statistically significant increases in ISR perception.
+However, this relationship was not the focus of the study, and we cannot draw any firm conclusions.
+We also cannot determine causality: it is possible that an improved instructor-student relationship results in an increase in perceived similarity, not vice-versa.
+Additionally, much of the research about similarity and ISRs comes from elementary and secondary school contexts.
+The exact relationship between perceived similarity and ISRs could be different at the college level.
+
+The exploratory modeling in Table \@ref(tab:noncausaltable) also indicates that ISR, or at least ISR perception, is related to student outcomes, as in relationships 5 and 6 in Figure \@ref(fig:causaldiagram).
+Again, we cannot conclude anything about causality.
+It is possible that students engage in behaviors that both earn higher grades and build better relationships with their instructors.
+It should also be noted that the quality of the ISR cannot be measured directly.
+It is also possible that instructor-student relationships can improve student outcomes, but the particular perceptions that are measured in this case are not the relevant factors, or perceived ISR is less important that the actions within an ISR.
+
+The environment of @citeoriginal's experiment is also relevant.
+They aimed to build ISRs in a context where the average class size is 36.4, students typically take five classes at a time, and instructors teach multiple classes.
+The intervention is appealingly simple, quick, and inexpensive, but such a small change is likely insignificant in the face of a wider campus culture that makes connections challenging.
+
+Further, the statistics found in the study indicate that the instructor and student populations are demographically different.
+Compared to the group of instructors, the student body had 3.5 times as many individuals who identified as Hispanic.
+Compared to students, the instructor population had 3.1 times as many individuals who identified as White.
+This is only a hint of the differences that may exist between the student and instructor populations.
+If students feel fundamentally different from their instructors, an intervention that provides superficial similarities is insufficient to bridge that gap.
+
+If colleges are considering investing in interventions that try to enhance similarity perception, we need to ask what it means to perceive similarity.
+Do some similarities matter more than others in ISRs?
+Should positive ISRs be contingent on similarity at all?
+As most students will go on to be instructors, we can expect, at the very least, that many students will not share life goals and trajectories with their instructors.
+There are other ways to cultivate ISRs that may make more sense and offer more equitable outcomes.
 
@@
 
-TODO: does this study have external validity?
-
-TODO: very long surveys pose limitations on accuracy
+While the study is internally valid, due to successful randomly-assigned treatment and control groups, the external validity is more questionable.
+The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate and followed through.
+The sample of instructors that would be interested in the study would not necessarily be equal to the overall population: they could be instructors who care more about supporting the work of other academics, instructors for whom a $150 giftcard was a significant draw, or instructors who are particularly interested in the role of ISRs.
+
+The instructors who participated were part of both the treatment and control group.
+That is, they received similarities for some students but not others.
+If a significant effect size had been found, this would be concerning, as the instructor's attention could be shifted from students in the control group to students in the treatment group, potentially exaggerating the effect.
+However, as no effect was found, this limitation is less concerning.
+
+As with all survey studies, there is a risk that self-reported information is inaccurate.
+In particular, the surveys used by @citeoriginal had many Likert scale questions about very similar topics.
+Due to survey fatigue, this may reduce the meaningfulness of the survey answers, especially those later in the survey.
+
+## Future directions
+
+Given the lack of significant findings, future studies could take a wide variety of approaches.
+First, more intensive, but similar, interventions to increase perceived similarity could be tested.
+Second, different methods of increasing perceived similarity, perhaps based in actual instructor-student interaction, could be tested.
+Third, interventions that aim to improve ISR without increasing perceived similarity could be considered.
 
@@
 demotab3[2:3,4] <- c("Male", "Female")
-demotab3[13:14,4] <- c("No", "Yes")
+demotab3[10:11,4] <- c("No", "Yes")
 
@@
   kableExtra::kable_styling(latex_options = c("scale_down", "striped", "hold_position"),
-                            stripe_index =c(2:3, 13:14, 16))
-
+                            stripe_index =c(2:3, 10:11, 13),
+                            stripe_color = "#f6f0fb")
 ```
@@
 
+# Appendix D
+
+```{r similarities, echo=FALSE, warning=FALSE, message=FALSE}
+tibble(Similarities = unique(expdata$sentence1)) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence2))) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence3))) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence4))) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence5))) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence6))) %>%
+  bind_rows(tibble(Similarities = unique(expdata$sentence7))) %>%
+  unique() %>%
+  arrange(Similarities) %>%
+  filter(Similarities != " ") %>%
+  knitr::kable(caption = "Similarities presented to students and instructors during treatment",
+               booktabs = TRUE, linesep = "", longtable = TRUE
+               ) %>%
+  column_spec(1, width = "47em") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
+                            stripe_color = "#f6f0fb")
+```
+
+\newpage
+
 # References
-
-TODO: look at the references that bookdown is generating and edit bibtex accordingly.

### 2021-04-20 outputs/paper/paper.Rmd
add april 19 draft

@@
 The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment.
-While the scalability and affordability of the intervention are desirable, there are no results in any of the targeted measures: those that affect and reflect college retention.
+The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no results in any of the targeted measures.
 These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
@@
 @citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance.
-In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships (ISRs) and measures of student success.
+In this field experiment, they evaluated the relationships between perceived similarity, instructor-student relationships (ISRs), and measures of student success.
 Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
@@
 These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence.
-These conclusions stand when matching instructor-student traits are considered, and when vulnerable student groups are evaluated independent of the wider student population.
-The lack of significant effects reflects the complicated task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.
+These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population.
+The lack of significant effects reflects the complicated task of increasing perceived similarity, hopefully to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.
 
@@
           no.space = TRUE,
-          keep.stat = c("rsq", "adj.rsq", "aic", "lr", "wald", "ll", "logrank"),
-          #omit.stat=c("ser","f", "ll"),
+          omit.stat=c("ser","f"),
           column.sep.width = "-10pt",

### 2021-04-26 outputs/exploratory.Rmd
shiny complete, train/test complete, first round of content edits done.

@@
 ```
+
+```{r}
+ddata <- expdata %>%
+  dplyr::select(treatment,
+                s1_sim, s1_tsr, 
+                s2_sim, s2_tsr, 
+                t2_sim1, t2_tsr,
+                grade, std_grade, 
+                t2_finalexam, f17_enrolled,
+                obj_exam,
+                s_firstgen,
+                s_hisp_black,
+                year,
+                s_firstgen) %>%
+  mutate(all = 1)
+
+write_csv(ddata, here("takingittothenextlevel/ddata.csv"))
+```
+
+```{r}
+teststudentfirstsimilaritymodel <- lm(data = expdata,
+                                  formula = s1_sim ~ treatment + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teststudentsecondsimilaritymodel <- lm(data = expdata,
+                                   formula = s2_sim ~ treatment + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teststudentsecondrelationshipmodel <- lm(data = expdata,
+                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+testteachersecondrelationshipmodel <- lm(data = expdata,
+                                     formula = t2_tsr ~ treatment + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+testgrademodel <- lm(data = expdata,
+                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+testexamgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                     na.action = na.omit)
+testteachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                                data = expdata,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+testenrollmentfallmodel <- glm(data = expdata,
+                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid + s_female + s_race + s_firstgen + ir_f16_gpa + year,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+```{r testmodeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(teststudentfirstsimilaritymodel,
+          teststudentsecondsimilaritymodel,
+          teststudentsecondrelationshipmodel,
+          testteachersecondrelationshipmodel,
+          testgrademodel,
+          testexamgrademodel,
+          testteachersecondsimilarity,
+          testenrollmentfallmodel,
+          title = "Replication model results",
+          align = TRUE,
+          ci = TRUE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          omit.stat=c("ser","f"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltable",
+          type = "latex",
+          header = FALSE,
+          initial.zero = FALSE,
+          omit = c("teacherid"),
+          omit.labels = c("Instructor fixed effect"))
+```
+
+```{r}
+ols_plot_resid_qq(studentfirstsimilaritymodel)
+```
+
+```{r}
+ols_plot_resid_fit(studentfirstsimilaritymodel)
+```
+
+```{r}
+ols_plot_resid_hist(studentfirstsimilaritymodel)
+```

### 2021-04-26 outputs/paper/paper.Rmd
shiny complete, train/test complete, first round of content edits done.

@@
   \usepackage{dcolumn}
+  \usepackage{float}
   \usepackage{rotating}
+  \usepackage{pdflscape}
+  \usepackage{booktabs}
 output:
@@
 subtitle: "Replicating ‘Taking It to the Next Level’"
-abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationship.
-The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment.
-The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no results in any of the targeted measures.
-These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
+abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationships.
+The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity, or instructor-student relationship.
+The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no measurable effects on any of the student outcome measures, including grades and reenrollment.
+These results are consistent with the original paper.\\par \\textbf{Keywords:} replication study, instructor-student relationships, college, similarity, college persistence"
 thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
@@
 library(DiagrammeR) # for causal diagram
+library(olsrr)
+library(tidymodels) # for train test
 
@@
 
-TODO: Shiny.
-Recreate Figure \@ref(fig:continuousoutcomes)(add graphs for standardized grade and persistence).
-Make an interactive element where you can filter the graph for a subset of the data -- specific racial s_race, firstgen/not, specific year of study.
-
 # Introduction
@@
 The causes are complex: factors that affect student persistence include student characteristics, organizational factors, peer environment, and student experiences.
-Interventions to improve student retention can target any of these aspects.
-
-@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance.
-In this field experiment, they evaluated the relationships between perceived similarity, instructor-student relationships (ISRs), and measures of student success.
-Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result.
+Interventions to improve student persistence can target any of these aspects.
+
+@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college student persistence and performance.
+In this field experiment, they evaluated the links between perceived similarity, instructor-student relationships (ISRs), and student success.
+Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive impact on measurable student outcomes.
 The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not.
@@
 Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.
-Finally,
 
@@
 These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population.
-The lack of significant effects reflects the complicated task of increasing perceived similarity, hopefully to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.
+The lack of significant effects reflects the complicated multi-stage task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.
 
@@
 
-Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, @citeoriginal was directly inspired by two previous works in middle and high schools.
-@Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships are associated with changes in student perceived self-efficacy, but not scholastic performance.
+Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, 'Taking It to the Next Level' was directly inspired by two previous works in middle and high schools.
+@Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships were associated with changes in student perceived self-efficacy, but not scholastic performance.
 Using a randomized controlled trial and surveys, @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015 found that when high school students were presented with similarities they shared with their teachers, they subsequently perceived greater similarity with their teachers.
@@
 
-@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship, course grade, and re-enrollment.
+@citeoriginal conducted a randomized controlled trial, with surveys, to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship perception, grades, and reenrollment.
 
@@
 The study took place in the 2017 spring term at a large Californian University.
-The study included 120 instructors and their 2,749 students.
+The study included 148 instructors and their 3,355 students.
 The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized.
 Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors.
-The initial sample of 147 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements.
+The initial sample of 148 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements.
 145 instructors consented and took the initial survey, resulting in an corresponding sample of 3,352 students.
@@
 
-All students participated in a survey immediately following the treatment or the placebo. They were surveyed again at the end of the course. Instructors were surveyed only at the end of the course.
+At the beginning of the term, the instructors were surveyed, to collect potential similarities.
+Then, students were surveyed, to collect potential similarities.
+Immediately, students were either given the treatment or the placebo, and then given further survey questions that asked about similarity and ISR.
+Students were surveyed again at the end of the term, regarding similarity and ISR.
+Instructors were surveyed about similarity and ISR only at the end of the term.
 
@@
 @citeoriginal identify key measures. Full descriptions of all measures are available in [Appendix A].
-
-Some are extracted from the student survey at the beginning of the term:
 
@@
 Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5.
-These responses were averaged to create a student perceived similarity scale.
+These responses were averaged to create a *student perceived similarity scale*.
   2.
 Immediately after the treatment or placebo, students answered seven questions about their anticipated instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create a student anticipated ISR scale.
-
-Others are extracted from the student survey at the end of the term:
-
-  3. At the end of the term, students answered the student perceived similarity scale questions again.
+These responses were averaged to create a *student anticipated ISR scale*.
+  3. At the end of the term, students answered the *student perceived similarity scale* questions again.
   4.
 At the end of the term, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create a student perceived ISR scale.
-
-Others come from the instructor survey at the end of the term:
-
-  5. Instructor similarity perception: At the end of the term, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
+These responses were averaged to create a *student perceived ISR scale*.
+  5. At the end of the term, instructors answered only one question about *instructor perceived similarity* with the student, on a scale of 1 to 5.
   6.
-Instructor perceived ISR scale: At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
-These responses were averaged to create an ISR scale.
-  7. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
-
-Finally, some are extracted from the university's internal records:
-
-  8. The grade that the student received in the course.
-  9. The student's course grade, standardized against other grades in the course.
-  10. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
+At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5.
+These responses were averaged to create an *instructor ISR perception* scale.
+  7. Instructors were asked to report the student's *grade on their final exam*, paper, or project.
+  8. The *grade that the student received in the course* was pulled from internal records.
+  9. The student's course grade was *standardized* against other grades in the course.
+  10. The student's *status as of Fall term 2017*, either enrolled or not, was pulled from internal records.
 
@@
 Test-retest reliability, where a scale is given to a group at different times, is a common way to assess this [@Coolican_2014].
-In our data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.
+In this data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.
 
@@
                             stripe_color = "#f6f0fb")
-
 ```
@@
 Finally, the student sample is 9.8% multiracial, but the instructor sample is 52.1% multiracial.
-Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classed differently as a result.
+Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classified differently as a result.
 For this reason, it is very difficult to draw conclusions.
@@
 In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories.
-More students identified as Hispanic, while more instructors identified as White and multi-racial.
+More students identified as Hispanic, while more instructors identified as White and selected multiple options.
 
@@
 
-```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(5,3), fig.cap="Histogram illustrating class sizes", out.width="60%", fig.align="center"}
-expdata %>%
+```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(9,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
+n_instructor <- expdata %>%
   dplyr::select(teacherid, n_course) %>%
-  unique() %>%
+  unique()
+
+n_by_instructor <- n_instructor %>%
   ggplot(aes(x = n_course)) +
   geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
-  labs(x = "Class size", y = "Frequency", title = "Some class sizes are very large") +
+  geom_vline(xintercept = mean(n_instructor %>% pull(n_course), na.rm = TRUE),
+             col = "#868cf7",
+             lwd = .5) +
+  labs(x = "Class size", y = "Number of classes") +
   theme_minimal()
-```
-
-## Outcomes of interest
+
+n_student <- expdata %>%
+  dplyr::select(id, n_course) %>%
+  unique()
+
+n_by_student <- n_student %>%
+  ggplot(aes(x = n_course)) +
+  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
+  geom_vline(xintercept = mean(n_student %>% pull(n_course), na.rm = TRUE),
+             col = "#868cf7",
+             lwd = .5) +
+  labs(x = "Class size", y = "Number of students in study") +
+  theme_minimal()
+
+combined <- plot_grid(n_by_instructor, n_by_student,
+          ncol = 2)
+
+title <- ggdraw() + 
+  draw_label(
+    "Some classes are very large (means marked)",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+plot_grid(title,
+          combined,
+          legend,
+          ncol = 1,
+          rel_heights = c(.1, 2))
+```
+
+## Outcomes of interest [^1]
+
+[^1]: Interactive graphs and tables showing outcomes of interest are available [here](https://amycfarrow.shinyapps.io/takingittothenextlevel/).
 
@@
 
-129 eligible instructors completed the first survey, but 6 of those instructors did not complete the second survey.
-Corresponding to the remaining 123 instructors, 2,801 students completed the first survey (1,392 in the treatment group, 1,409 in the control group), but 682 of those students did not complete the second survey (328 in the treatment group, 354 in the control group).
-The attrition in the treatment and control groups was comparable (24% in the treatment group, 25% in the control group).
+140 eligible instructors completed the first survey, and 2,808 of their students completed the first survey.
+21 of those instructors did not complete the second survey.
+2,808 students completed the first survey (1,396 in the treatment group, 1,412 in the control group), but 755 of those students did not complete the second survey (364 in the treatment group, 391 in the control group).
+The attrition in the treatment and control groups was comparable (26.1% in the treatment group, 27.7% in the control group).
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many instructors participated? 148
+isrdata %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# How many students could theoretically participate, based on their instructor participating? 3355
+isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many instructors consented to the first survey? 145
+isrdata %>%
+  filter(t1_consent == 1) %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# How many students does that correspond to? 3352
+isrdata %>%
+  filter(t1_consent == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# How many of those students consented to survey 1? 3229
+isrdata %>%
+  filter(t1_consent == 1, s1_consent == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# So total participants in the study: 145 instructors and 3229 students.
+# Now reasons for removal from analysis.
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# If we consider non-eligibility, how many instructors are left? 140
+isrdata %>%
+  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students") %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# And students? 3112
+isrdata %>%
+  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# Check for instructors completing the first survey. 140
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t1_complete == 1) %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# The number of students that corresponds to. 3112
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# Then, how many students completed the first survey? 2808
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+## Treatment: 1396
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & treatment == "Treatment") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+## Control 1412
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1 
+         & treatment == "Control") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many instructors completed the second survey? 119
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t1_complete == 1
+         & t2_complete == 1) %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# How many students does that correspond to? 2701
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# Students that completed the second survey: 2053
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & s2_complete) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# Treatment 1032
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & s2_complete
+         & treatment == "Treatment") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# Control 1021
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & s2_complete 
+         & treatment == "Control") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many teachers spent enough time on the survey? 119.
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t1_complete == 1
+         & t2_complete == 1
+         & t1_timer_consent_3 > 1
+         & s1_timer_consent_3 > 1
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & t2_timer > 10) %>%
+  dplyr::select(teacherid) %>%
+  unique() %>%
+  count()
+
+# Students that corresponds to: 2560
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & t1_timer_consent_3 > 1
+         & s1_timer_consent_3 > 1
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & t2_timer > 10) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+
+# Students that spent enough time on the surveys: 2273
+isrdata %>%
+  filter(t1_consent == 1 
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & s1_consent == 1
+         & t1_complete == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & t1_timer_consent_3 > 1
+         & s1_timer_consent_3 > 1
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & t2_timer > 10
+         & s1_timer_gtky1_3 > 10
+         & s1_timer_gtky2_3 > 10
+         & s1_timer_gtky3_3 > 10
+         & s1_timerpre1_3 > 10
+         & s1_timerpre2_3 > 10
+         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
+         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
+         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
+         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
+         & s1_timerpost1_3 > 10
+         & s1_timerpost2_3 > 10
+         & s1_timerdemo1_3 > 10
+         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
+         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
+         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
+         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
+         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
+         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
+         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count()
+```
 
@@
 From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis.
-Units of study were excluded because the instructor did not participate in the study (33,486); the student did not consent (123); the student did not spend more than a second reading the consent page (1); the initial student survey, initial instructor survey, or end-of-term instructor survey was not complete (428); the instructor mistakenly administered the wrong survey (30); the course was online (20); the course was for graduate students (50); or a participant did not spend more than ten seconds reading and answering a page with five or more questions (427).
-The ten-second time limit is somewhat arbitrary, and a different choice could lead to different study results.
+Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend at more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428).
+The time limits are somewhat arbitrary, and different choices could have different analysis results.
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many students were discarded because the instructor did not participate? 33175
+all_students <- isrdata %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+all_students - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many were excluded because the course wasn't actually eligible? 96
+instructor_participated <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+instructor_participated - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many were excluded because the instructor didn't adminster the correct survey? 31
+eligible <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+eligible - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many instructors did not complete survey 1? 0
+eligible_correct <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+eligible_correct - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many students did not consent? 116
+teacher_complete <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+teacher_complete - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many students did not complete survey 1? 304
+student_consent <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+student_consent - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1
+         & s1_fullsurvey == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many were discarded because instructors didn't complete the second survey? 107
+student_complete <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1
+         & s1_fullsurvey == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+student_complete - isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# How many were discarded because someone didn't spend enough time? 428
+instructor_complete <- isrdata %>%
+  filter(!is.na(teacherid)) %>%
+  filter(t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+
+instructor_complete - isrdata %>%
+  filter(!is.na(teacherid)
+         & t1_consent == 1
+         & t_drop_reasons != "Taught online course"
+         & t_drop_reasons != "Administered to only graduate students"
+         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
+         & t1_consent == 1
+         & t1_complete == 1
+         & s1_consent == 1
+         & s1_fullsurvey == 1
+         & t2_complete == 1
+         & t1_timer_consent_3 > 1
+         & s1_timer_consent_3 > 1
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & t2_timer > 10
+         & s1_timer_gtky1_3 > 10
+         & s1_timer_gtky2_3 > 10
+         & s1_timer_gtky3_3 > 10
+         & s1_timerpre1_3 > 10
+         & s1_timerpre2_3 > 10
+         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
+         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
+         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
+         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
+         & s1_timerpost1_3 > 10
+         & s1_timerpost2_3 > 10
+         & s1_timerdemo1_3 > 10
+         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
+         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
+         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
+         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
+         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
+         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
+         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))) %>%
+  dplyr::select(id) %>%
+  unique() %>%
+  count() %>%
+  pull()
+```
 
@@
 
-$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA).
+$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study).
 ISR anticipation is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes.
@@
 
-$X_{2j}$ is a vector of instructor-level covariates (course size, teacher ID).
-
-$\epsilon_{ij}$ is a clustered residual.
-
-$\beta_{0}$, $beta_{1}$, $\Gamma_{1}$, $\Gamma_{2}$, and $a_{k}$ are coefficients on the resulting models.
+$X_{2j}$ is a vector of instructor-level covariates (class size).
+
+$\epsilon_{ij}$ is a standard error.
+As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard error and p-values when compared with the original paper.
+This does not affect the regression coefficients.
+
+$\beta_{0}$, $\beta_{1}$, $\Gamma_{1}$, $\Gamma_{2}$, and $a_{k}$ are coefficients on the resulting models.
 
@@
 
-When creating additional models, I focused on three key outcomes: end-of-term student perceived similarity scale, end-of-term student perceived ISR scale, and standardized course grade.
-I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance) and perceptions that could improve student outcomes (like student ISR perception), and this particular intervention tries to accomplish these goals by improving similarity perception.
+When creating additional models, I focused on three key outcomes: standardized course grade, end-of-term student perceived ISR scale, and end-of-term student perceived similarity scale.
+I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes than outcomes from subsequent terms; and because previous studies show more detectable impact on student perception than instructor perception.
 Equation \@ref(eq:mylinmodel) shows the linear model used.
@@
 studentfirstsimilaritymodel <- lm(data = expdata,
-                                  formula = s1_sim ~ treatment + teacherid,
+                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondsimilaritymodel <- lm(data = expdata,
-                                   formula = s2_sim ~ treatment + teacherid,
+                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondrelationshipmodel <- lm(data = expdata,
-                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 teachersecondrelationshipmodel <- lm(data = expdata,
-                                     formula = t2_tsr ~ treatment + teacherid,
+                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 grademodel <- lm(data = expdata,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
+                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
+                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                 data = expdata,
@@
 enrollmentfallmodel <- glm(data = expdata,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
+                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
                            family = "binomial",
@@
 
-In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.189, 95% CI[0.134,0.245]).
-Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.118, 95% CI[0.051,0.184]).
+In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]).
+Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]).
 However, both coefficients are minimal on a scale of 1 to 5.
@@
 Model 3 shows no significant relationship between treatment and student end-of-term ISR perception.
-However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.566, 95% CI[0.52,0.612]).
-For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.565.
+However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]).
+For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671.
 
 Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-term ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent term.
+
+Models 1 through 8 were also repeated with additional student covariates (gender, race, first-generation status, CGPA, and year), and the results still showed a treatment effect on student similarity perception but no other effect.
+
+Model validation for the two primary significant models, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].
 
@@
           float.env = "sidewaystable",
+          notes.label = "",
           no.space = TRUE,
-          omit.stat=c("ser","f"),
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
           column.sep.width = "-10pt",
@@
           initial.zero = FALSE,
-          omit = c("teacherid"),
-          omit.labels = c("Instructor fixed effect"),
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student sim. 1",
@@
                              "Persistence"),
-          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
-```
+          covariate.labels = c("Treatment", "Student ISR 1", "Course size", "Student female", "CGPA", "Year of study"))
+```
+
+Because some students are more commonly targeted by persistence interventions, the subset of Hispanic and/or Black college student participants is considered separately from the rest of the student population.
+While the effect of treatment on student similarity perception is larger than for the general student population ($\beta$ = 0.199, 95% CI [0.124, 0.274] for initial perception; $\beta$ = 0.116, 95% CI [0.021, 0.211] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+Full results are in [Appendix E].
+
+Similarly, the subset of first-generation college student participants is considered.
+While the effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+Full results are in [Appendix F].
+
+### Exploratory replication results
 
@@
 
-expdatahb <- expdata %>%
-  filter(s_hisp_black == 1)
-
-studentfirstsimilaritymodelhb <- lm(data = expdatahb,
-                                  formula = s1_sim ~ treatment + teacherid,
-                                     na.action = na.omit)
-studentsecondsimilaritymodelhb <- lm(data = expdatahb,
-                                   formula = s2_sim ~ treatment + teacherid,
-                                     na.action = na.omit)
-studentsecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
-                                     na.action = na.omit)
-teachersecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = t2_tsr ~ treatment + teacherid,
-                                     na.action = na.omit)
-grademodelhb <- lm(data = expdatahb,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                                     na.action = na.omit)
-examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
-                                data = expdatahb,
-                                Hess = TRUE,
-                                     na.action = na.omit)
-enrollmentfallmodelhb <- glm(data = expdatahb,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                           family = "binomial",
-                                     na.action = na.omit)
-```
-
-Table \@ref(tab:modeltablehb) shows the same models as in Table \@ref(tab:modeltable), created only using the subset of Hispanic and/or Black college student participants.
-While the effect of treatment on student similarity perception is larger than for the general student population ($\beta$ = 0.21, 95% CI[0.137,0.282] for initial perception; $\beta$ = 0.148, 95% CI[0.06,0.237] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
-
-```{r modeltablehb, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
-# Make table to display model results
-stargazer(studentfirstsimilaritymodelhb,
-          studentsecondsimilaritymodelhb,
-          studentsecondrelationshipmodelhb,
-          teachersecondrelationshipmodelhb,
-          grademodelhb,
-          examgrademodelhb,
-          teachersecondsimilarityhb,
-          enrollmentfallmodelhb,
-          title = "Replication model results: Hispanic and Black students only",
-          align = TRUE,
-          ci = TRUE,
-          initial.zero = FALSE,
-          float.env = "sidewaystable",
-          no.space = TRUE,
-          omit.stat=c("ser","f"),
-          column.sep.width = "-10pt",
-          label = "tab:modeltablehb",
-          type = "latex",
-          header = FALSE,
-          omit = c("teacherid"),
-          omit.labels = c("Instructor fixed effect"),
-          dep.var.labels = c("Student sim. 1",
-                             "Student sim. 2",
-                             "Student ISR 2",
-                             "Instructor ISR 2",
-                             "Course grade",
-                             "Final grade",
-                             "Instructor sim. 2",
-                             "Persistence"),
-          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Make all replication models
-
-expdata1g <- expdata %>%
-  filter(s_firstgen == 1)
-
-studentfirstsimilaritymodel1g <- lm(data = expdata1g,
-                                  formula = s1_sim ~ treatment + teacherid,
-                                     na.action = na.omit)
-studentsecondsimilaritymodel1g <- lm(data = expdata1g,
-                                   formula = s2_sim ~ treatment + teacherid,
-                                     na.action = na.omit)
-studentsecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = s2_tsr ~ treatment + s1_tsr + teacherid,
-                                     na.action = na.omit)
-teachersecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = t2_tsr ~ treatment + teacherid,
-                                     na.action = na.omit)
-grademodel1g <- lm(data = expdata1g,
-                 formula = grade ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                                     na.action = na.omit)
-examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                                     na.action = na.omit)
-teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + teacherid,
-                                data = expdata1g,
-                                Hess = TRUE,
-                                     na.action = na.omit)
-enrollmentfallmodel1g <- glm(data = expdata1g,
-                           formula = f17_enrolled ~ treatment + s1_female + ir_f16_gpa + teacherid,
-                           family = "binomial",
-                                     na.action = na.omit)
-```
-
-Table \@ref(tab:modeltable1g) shows the same models as in Table \@ref(tab:modeltable), created only using the subset of first-generation college student participants.
-While the effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.272, 95% CI[0.188,0.357] for initial perception; $\beta$ = 0.165, 95% CI[0.062,0.268] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
-
-```{r modeltable1g, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
-# Make table to display model results
-stargazer(studentfirstsimilaritymodel1g,
-          studentsecondsimilaritymodel1g,
-          studentsecondrelationshipmodel1g,
-          teachersecondrelationshipmodel1g,
-          grademodel1g,
-          examgrademodel1g,
-          teachersecondsimilarity1g,
-          enrollmentfallmodel1g,
-          title = "Replication model results: first-generation students only",
-          align = TRUE,
-          initial.zero = FALSE,
-          ci = TRUE,
-          float.env = "sidewaystable",
-          no.space = TRUE,
-          omit.stat=c("ser","f"),
-          column.sep.width = "-10pt",
-          label = "tab:modeltable1g",
-          type = "latex",
-          header = FALSE,
-          omit = c("teacherid"),
-          omit.labels = c("Instructor fixed effect"),
-          dep.var.labels = c("Student sim. 1",
-                             "Student sim. 2",
-                             "Student ISR 2",
-                             "Instructor ISR 2",
-                             "Course grade",
-                             "Final grade",
-                             "Instructor sim. 2",
-                             "Persistence"),
-          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
-```
-
-### Exploratory replication results
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Make all replication models
-
 studentfirstisrmodel1g <- lm(data = expdata,
-                                  formula = s1_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
+                                  formula = s1_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 studentsecondrelationshipmodel1g <- lm(data = expdata,
-                                     formula = s2_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
+                                     formula = s2_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 teachersecondrelationshipmodel1g <- lm(data = expdata,
-                                     formula = t2_tsr ~ s_firstgen + ir_f16_gpa + grade + teacherid,
-                                     na.action = na.omit)
-studentfirstisrmodelhb <- lm(data = expdata,
-                                  formula = s1_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
-                                     na.action = na.omit)
-studentsecondrelationshipmodelhb <- lm(data = expdata,
-                                     formula = s2_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
-                                     na.action = na.omit)
-teachersecondrelationshipmodelhb <- lm(data = expdata,
-                                     formula = t2_tsr ~ s_hisp_black + ir_f16_gpa + grade + teacherid,
-                                     na.action = na.omit)
-
-```
-
-Table \@ref(tab:modeltablerel) disregards the treatment condition, instead considering the relationship between student identity and similarity/ISR.
+                                     formula = t2_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+```
+
+Table \@ref(tab:modeltablerel1) disregards the treatment condition, instead considering the relationship between student identity and similarity/ISR.
 While we cannot make causal inferences, we can explore relationships between the variables in this particular study.
-If a student identifies as first generation, or as Hispanic and/or Black, can we anticipate a different student or instructor ISR perception?
-There is a significant negative relationship between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.1, 95% CI[-0.158,-0.041]).
-
-```{r modeltablerel, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+If a student identifies as first generation, can we anticipate a different student or instructor ISR perception?
+In this case, there is no significant relationship.
+
+```{r modeltablerel1, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
 # Make table to display model results
@@
           teachersecondrelationshipmodel1g,
-          studentfirstisrmodelhb,
+          title = "Replication model results: first-generation status and ISR",
+          align = TRUE,
+          ci = TRUE,
+          no.space = TRUE,
+          initial.zero = FALSE,
+          notes.label = "",
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          column.sep.width = "-10pt",
+          label = "tab:modeltablerel1",
+          type = "latex",
+          header = FALSE,
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
+          dep.var.labels = c("Student ISR 1",
+                             "Student ISR 2",
+                             "Instructor ISR 2"),
+          covariate.labels = c("Student first gen.", "Student female", "CGPA", "Course grade", "Year of study", "Course size"))
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+studentfirstisrmodelhb <- lm(data = expdata,
+                                  formula = s1_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+studentsecondrelationshipmodelhb <- lm(data = expdata,
+                                     formula = s2_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+teachersecondrelationshipmodelhb <- lm(data = expdata,
+                                     formula = t2_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+
+```
+
+Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative relationship between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.096, 95% CI [-0.166, -0.026]).
+
+```{r modeltablerel2, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstisrmodelhb,
           studentsecondrelationshipmodelhb,
           teachersecondrelationshipmodelhb,
-          title = "Replication model results: first-generation status, Hispanic and/or Black students, and ISR",
+          title = "Replication model results: Hispanic and/or Black students and ISR",
           align = TRUE,
           ci = TRUE,
-          float.env = "sidewaystable",
           no.space = TRUE,
           initial.zero = FALSE,
-          omit.stat=c("ser","f"),
+          notes.label = "",
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
           column.sep.width = "-10pt",
-          label = "tab:modeltablerel",
+          label = "tab:modeltablerel2",
           type = "latex",
           header = FALSE,
-          omit = c("teacherid"),
-          omit.labels = c("Instructor fixed effect"),
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student ISR 1",
                              "Student ISR 2",
-                             "Instructor ISR 2",
-                             "Student ISR 1",
-                             "Student ISR 2",
                              "Instructor ISR 2"),
-          covariate.labels = c("Student first gen.", "Student Hispanic and/or Black", "CGPA", "Course grade"))
+          covariate.labels = c("Student Hispanic and/or Black", "Student female", "CGPA", "Course grade", "Year of study", "Course size"))
 ```
@@
 grademodel_ <- lm(data = expdata,
-                 formula = grade ~ teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
+                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 examgrademodel_ <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~  teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
+                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 enrollmentfallmodel_ <- glm(data = expdata,
-                           formula = f17_enrolled ~  teacherid + treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
+                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
                            family = "binomial",
@@
 
-Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.733, 95% CI[0.706,0.76]), and instructor similarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.512, 95% CI[0.483,0.541]).
-Additionally, course size is negatively related to student ISR perception ($\beta$ = -0.102, 95% CI[-0.178,-0.026]).
-
-Models 3 and 4 show that student ISR perception and instructor ISR perception are both significantly related to course grade and final grade.
-Instructor perception is a stronger predictor ($\beta$ = 0.395, 95% CI[0.343,0.447] for course grade; $\beta$ = 0.432, 95% CI[0.299,0.565] for final grade) than student perception ($\beta$ = 0.076, 95% CI[0.023,0.129] for course grade; $\beta$ = 0.144, 95% CI[0.017,0.271] for final grade).
+Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.512, 95% CI [0.483, 0.541]).
+
+Models 3 and 4 show that student ISR perception is significantly related to course grade, while instructor ISR perception is significantly related to course grade and final grade.
+Instructor perception is a stronger predictor ($\beta$ = 0.339, 95% CI [0.294, 0.384] for course grade; $\beta$ = 0.643, 95% CI [0.527, 0.758] for final grade) than student perception ($\beta$ = 0.105, 95% CI [0.055, 0.155] for course grade).
 
@@
           no.space = TRUE,
-          omit.stat=c("ser","f"),
+          notes.label = "",
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
           column.sep.width = "-10pt",
@@
           header = FALSE,
-          omit = c("teacherid"),
-          omit.labels = c("Instructor fixed effect"),
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
           dep.var.labels = c("Student ISR 2",
@@
                              "Student female",
-                             "CGPA"))
-```
-
-TODO: clustered residuals (low priority)
+                             "CGPA",
+                             "Year"))
+```
 
@@
 mystudentsecondsimilaritymodel <- lm(data = expdata,
-                                     formula = s2_sim ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     formula = s2_sim ~ treatment + n_course + race_match + gend_match + age_dif + nsims,
                                      na.action = na.omit)
 mystudentsecondrelationshipmodel <- lm(data = expdata,
-                                     formula = s2_tsr ~ treatment + n_course + race_match + gend_match + age_dif,
+                                     formula = s2_tsr ~ treatment + n_course + race_match + gend_match + age_dif + nsims,
                                      na.action = na.omit)
 mystdgrademodel <- lm(data = expdata,
-                      formula = grade ~ treatment + ir_f16_gpa + n_course + race_match + gend_match + age_dif,
+                      formula = grade ~ treatment + ir_f16_gpa + n_course + race_match + gend_match + age_dif + nsims,
                                      na.action = na.omit)
@@
 Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
-As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.107, 95% CI[0.035,0.179]).
+As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.108, 95% CI [0.036, 0.179]).
 However, the coefficient is even lower than in the replication models.
@@
           no.space = TRUE,
+          notes.label = "",
           initial.zero = FALSE,
-          omit.stat=c("ser","f"),
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
           column.sep.width = "-10pt",
@@
                                "Course size",
-                               "Matching racial self-ID", "Matching gender self-ID", "Age difference")
+                               "Matching racial self-ID", "Matching gender self-ID", "Age difference", "Number of similarities")
           )
@@
 However, the treatment has no significant impact on student ISR perception, instructor perception, or student outcomes.
-These results stand when student gender, prior GPA, course size, matching student-instructor traits, and instructor fixed effects are controlled for.
 
@@
 The intervention refers to informing students and instructors about shared traits.
+It does not target actual instructor-student similarity -- it simply aims to heighten awareness of any existing similarities.
 Similarity perception is measured by the two perceived similarity scales.
 ISR, or instructor-student relationship, cannot be directly measured -- only perception can be measured, through the perceived ISR scales.
-The student outcomes, grades and persistence, are factual and more easily measured.
+The student outcomes, grades and persistence, are more easily measured.
 The diagram shows the complexity of succeeding with this type of intervention: the treatment must affect perception, which must then affect relationships, which must then affect student outcomes.
@@
 What similarities were students and instructors informed of?
-[Appendix D] contains a complete list.
+[Appendix G] contains a complete list.
 Students and instructors were presented with generalities such as "An instructor’s most important characteristic?
@@
 It is possible that students engage in behaviors that both earn higher grades and build better relationships with their instructors.
-It should also be noted that the quality of the ISR cannot be measured directly.
-It is also possible that instructor-student relationships can improve student outcomes, but the particular perceptions that are measured in this case are not the relevant factors, or perceived ISR is less important that the actions within an ISR.
+Because the quality of the ISR cannot be measured directly, it is also possible that instructor-student relationships can improve student outcomes, but the particular perceptions that are measured in this case are not the relevant factors, or perceived ISR is less important that the real behaviors within an ISR.
 
 The environment of @citeoriginal's experiment is also relevant.
-They aimed to build ISRs in a context where the average class size is 36.4, students typically take five classes at a time, and instructors teach multiple classes.
+They aimed to build ISRs in a context where a student's average class size is 49.5, students typically take five classes at a time, and instructors teach multiple classes, sometimes at multiple campuses.
 The intervention is appealingly simple, quick, and inexpensive, but such a small change is likely insignificant in the face of a wider campus culture that makes connections challenging.
 
-Further, the statistics found in the study indicate that the instructor and student populations are demographically different.
+Further, the study indicates that the instructor and student populations are demographically different.
 Compared to the group of instructors, the student body had 3.5 times as many individuals who identified as Hispanic.
@@
 This is only a hint of the differences that may exist between the student and instructor populations.
-If students feel fundamentally different from their instructors, an intervention that provides superficial similarities is insufficient to bridge that gap.
+As most students will not go on to be instructors, we can expect, at the very least, that many students will not share life goals and trajectories with their instructors.
+If students feel fundamentally different from their instructors, based on life experience or socio-economic status, an intervention that provides superficial similarities is insufficient to bridge that gap.
 
@@
 Do some similarities matter more than others in ISRs?
-Should positive ISRs be contingent on similarity at all?
-As most students will go on to be instructors, we can expect, at the very least, that many students will not share life goals and trajectories with their instructors.
+If demographics traits, which are already part of an inequitable campus environment, are important for perceived similarity, Should positive ISRs be contingent on similarity at all?
 There are other ways to cultivate ISRs that may make more sense and offer more equitable outcomes.
@@
 
-While the study is internally valid, due to successful randomly-assigned treatment and control groups, the external validity is more questionable.
-The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate and followed through.
-The sample of instructors that would be interested in the study would not necessarily be equal to the overall population: they could be instructors who care more about supporting the work of other academics, instructors for whom a $150 giftcard was a significant draw, or instructors who are particularly interested in the role of ISRs.
-
+The study's internal validity comes from the successfully randomly-assigned treatment and control groups, but there is one concern.
 The instructors who participated were part of both the treatment and control group.
 That is, they received similarities for some students but not others.
-If a significant effect size had been found, this would be concerning, as the instructor's attention could be shifted from students in the control group to students in the treatment group, potentially exaggerating the effect.
-However, as no effect was found, this limitation is less concerning.
+If more significant effect sizes, or larger effect sizes, had been found, this potential lack of separation between the treatment and control groups would be a problem, because the instructor's attention could have simply been shifted from students in the control group to students in the treatment group, potentially exaggerating the effect.
+However, as little treatment effect was found, this limitation is less concerning.
+
+The external validity is more questionable.
+The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate and followed through.
+The sample of instructors that would be interested in the study would not necessarily be comparable to the overall population: they could be instructors who care more about supporting the work of other academics, instructors for whom a $150 giftcard was a significant draw, or instructors who are particularly interested in the role of ISRs.
 
@@
 In particular, the surveys used by @citeoriginal had many Likert scale questions about very similar topics.
-Due to survey fatigue, this may reduce the meaningfulness of the survey answers, especially those later in the survey.
+Survey fatigue may have reduced the meaningfulness of the survey answers, especially those later in the survey.
 
@@
 First, more intensive, but similar, interventions to increase perceived similarity could be tested.
+These interventions could include more specific or important similarities between students and instructors.
 Second, different methods of increasing perceived similarity, perhaps based in actual instructor-student interaction, could be tested.
+Such interventions are less desirable from the college's point of view, because they would be more costly and demanding to implement.
 Third, interventions that aim to improve ISR without increasing perceived similarity could be considered.
+Similarity is only one of many theorized methods for improved ISRs.
+These would be entirely different lines of inquiry, however.
 
@@
 
-* Initial student survey:
-  1. Student perceived similarity scale (s1_sim)
+* Initial student survey
+  1. Student perceived similarity scale
       - Overall, how similar to your instructor's values do you think your values are?
@@
       - Overall, how similar do you think you and your instructor are?
-  2. Student anticipated ISR scale (s1_tsr)
+  2. Student anticipated ISR scale
       - How much do you think you will enjoy learning from this instructor?
@@
       - Overall, how much do you think you will learn from this instructor?
-* End of term student survey:
-  3. Student perceived similarity scale (s2_sim)
-  4. Student perceived ISR scale (s2_tsr)
+* End of term student survey
+  3. Student perceived similarity scale
+  4. Student perceived ISR scale
       - How much do you enjoy learning from this professor?
@@
       - Overall, how much do you think you have learned from this professor?
-* End of term instructor survey:
-  5. Instructor similarity perception (t2_sim1):
+* End of term instructor survey
+  5. Instructor similarity perception
       - Overall, how similar do you think you and STUDENTNAME are?
-  6. Instructor perceived ISR scale (t2_tsr):
+  6. Instructor perceived ISR scale
       - How much did you enjoy helping STUDENTNAME learn?
@@
       - Overall, how much did STUDENTNAME learn from you?
-  8. Final grade (t2_finalexam): Instructors were asked to report the student's grade on their final exam, paper, or project.
+  8. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.
 * University internal records:
-  9. Course grade (grade): The final grade that the student received in the course.
-  10. Standardized course grade (std_grade): The student's final grade, standardized against other grades in the course.
-  11. Persistence (f17_enrolled): The student's status as of Fall term 2017: not enrolled or enrolled.
+  9. Course grade: The final grade that the student received in the course.
+  10. Standardized course grade: The student's final grade, standardized against other grades in the course.
+  11. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.
 
@@
 
+```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+set.seed(1847)
+expdata_split <- rsample::initial_split(expdata, prop = 0.80)
+expdata_train <- rsample::training(expdata_split)
+expdata_test  <-  rsample::testing(expdata_split)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+studentfirstsimilaritymodelv <- lm(data = expdata_train,
+                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+studentsecondsimilaritymodelv <- lm(data = expdata_train,
+                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+```
+
+Table \@ref(tab:modelvtable) shows the two primary significant models, redone using a random train/test split (80% training, 20% testing).
+Table \@ref(tab:traintesttable) shows metrics to assess the performance of the two models on the training and testing datasets.
+
+```{r modelvtable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstsimilaritymodelv,
+          studentsecondsimilaritymodelv,
+          title = "Train replication model results",
+          align = TRUE,
+          ci = TRUE,
+          no.space = TRUE,
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          column.sep.width = "-10pt",
+          label = "tab:modelvtable",
+          notes.label = "",
+          type = "latex",
+          header = FALSE,
+          initial.zero = FALSE,
+          keep = c("treatment", "n_course", "s_female", "ir_f16_gpa", "year", "Constant"),
+          #omit = c("teacherid"),
+          #mit.labels = c("Instructor fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2"),
+          covariate.labels = c("Treatment", "Course size", "Student female", "CGPA", "Year of study"))
+```
+
+```{r traintesttable, echo=FALSE, warning=FALSE, message=FALSE}
+expdata_train <- expdata_train %>%
+  mutate(pred1 = predict(studentfirstsimilaritymodelv, expdata_train),
+         pred2 = predict(studentsecondsimilaritymodelv, expdata_train))
+
+expdata_test <- expdata_test %>%
+  mutate(pred1 = predict(studentfirstsimilaritymodelv, expdata_test),
+         pred2 = predict(studentsecondsimilaritymodelv, expdata_test))
+
+metrics(expdata_train, s1_sim, pred1) %>%
+  bind_cols(metrics(expdata_test, s1_sim, pred1) %>% dplyr::select(.estimate)) %>%
+  bind_rows(metrics(expdata_train, s2_sim, pred2) %>%
+              bind_cols(metrics(expdata_test, s2_sim, pred2) %>% dplyr::select(.estimate))) %>%
+kable(caption = "Metrics for student similarity perception models",
+        booktabs = TRUE,
+        col.names = c("Metric", "Estimator", "Training data", "Testing data")) %>%
+  pack_rows("Initial", 1, 3) %>%
+  pack_rows("End-of-term", 4, 6) %>%
+  kable_styling()
+```
+
+\newpage
+
+# Appendix E
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+
+expdatahb <- expdata %>%
+  filter(s_hisp_black == 1)
+
+studentfirstsimilaritymodelhb <- lm(data = expdatahb,
+                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+studentsecondsimilaritymodelhb <- lm(data = expdatahb,
+                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+studentsecondrelationshipmodelhb <- lm(data = expdatahb,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondrelationshipmodelhb <- lm(data = expdatahb,
+                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+grademodelhb <- lm(data = expdatahb,
+                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                data = expdatahb,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+enrollmentfallmodelhb <- glm(data = expdatahb,
+                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+```{r modeltablehb, results=c('asis','hold'), echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstsimilaritymodelhb,
+          studentsecondsimilaritymodelhb,
+          studentsecondrelationshipmodelhb,
+          teachersecondrelationshipmodelhb,
+          grademodelhb,
+          examgrademodelhb,
+          teachersecondsimilarityhb,
+          enrollmentfallmodelhb,
+          title = "Replication model results: Hispanic and Black students only",
+          align = TRUE,
+          ci = TRUE,
+          initial.zero = FALSE,
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          notes.label = "",
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          column.sep.width = "-12pt",
+          label = "tab:modeltablehb",
+          font.size = "footnotesize",
+          type = "latex",
+          header = FALSE,
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Instructor sim. 2",
+                             "Persistence"),
+          covariate.labels = c("Treatment", "Student ISR 1", "Course size", "Student female", "CGPA", "Year of study"))
+```
+
+\newpage
+
+# Appendix F
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all replication models
+
+expdata1g <- expdata %>%
+  filter(s_firstgen == 1)
+
+studentfirstsimilaritymodel1g <- lm(data = expdata1g,
+                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+studentsecondsimilaritymodel1g <- lm(data = expdata1g,
+                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+studentsecondrelationshipmodel1g <- lm(data = expdata1g,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondrelationshipmodel1g <- lm(data = expdata1g,
+                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+grademodel1g <- lm(data = expdata1g,
+                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
+                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                data = expdata1g,
+                                Hess = TRUE,
+                                     na.action = na.omit)
+enrollmentfallmodel1g <- glm(data = expdata1g,
+                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                           family = "binomial",
+                                     na.action = na.omit)
+```
+
+```{r modeltable1g, results=c('asis','hold'), echo=FALSE, warning=FALSE, message=FALSE}
+# Make table to display model results
+stargazer(studentfirstsimilaritymodel1g,
+          studentsecondsimilaritymodel1g,
+          studentsecondrelationshipmodel1g,
+          teachersecondrelationshipmodel1g,
+          grademodel1g,
+          examgrademodel1g,
+          teachersecondsimilarity1g,
+          enrollmentfallmodel1g,
+          title = "Replication model results: first-generation students only",
+          align = TRUE,
+          initial.zero = FALSE,
+          ci = TRUE,
+          notes.label = "",
+          float.env = "sidewaystable",
+          no.space = TRUE,
+          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          column.sep.width = "-12pt",
+          font.size = "footnotesize",
+          label = "tab:modeltable1g",
+          type = "latex",
+          header = FALSE,
+          #omit = c("teacherid"),
+          #omit.labels = c("Instructor fixed effect"),
+          dep.var.labels = c("Student sim. 1",
+                             "Student sim. 2",
+                             "Student ISR 2",
+                             "Instructor ISR 2",
+                             "Course grade",
+                             "Final grade",
+                             "Instructor sim. 2",
+                             "Persistence"),
+          covariate.labels = c("Treatment", "Student ISR 1", "Course size", "Student female", "CGPA", "Year of study"))
+```
+
+\newpage
+
+# Appendix G
+
 ```{r similarities, echo=FALSE, warning=FALSE, message=FALSE}

### 2021-04-26 takingittothenextlevel/app.R

@@ -1,0 +1,244 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(tidyverse)
+library(cowplot)
+library(finalfit)
+library(here)
+
+ddata <- read_csv("ddata.csv")
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+
+    # Application title
+    titlePanel("Improving Instructor-Student Relationships by Establishing Similarity"),
+
+    # Sidebar with a slider input for number of bins 
+    sidebarLayout(
+        
+        sidebarPanel(
+            selectInput('subset', 
+                        "Which subset of students would you like to see results for?",
+                        choices = c("All students",
+                                    "First-generation college students",
+                                    "Hispanic and/or Black students"))
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+            tabsetPanel(
+                tabPanel("Plot", plotOutput("distPlot")), 
+                tabPanel("Table", tableOutput("table")), 
+                tabPanel("Information", htmlOutput("summary"))
+            )
+           
+        )
+    )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+
+    output$distPlot <- renderPlot({
+        
+        bin = 30
+
+        if (input$subset == "All students") {
+            data <- ddata
+        } else if (input$subset == "First-generation college students") {
+            data <- ddata %>%
+                filter(s_firstgen == 1)
+        } else {
+            data <- ddata %>%
+                filter(s_hisp_black == 1)
+        }
+        
+        s1_simplot <- data %>% 
+            ggplot(aes(x = s1_sim, fill = treatment)) +
+            geom_histogram(position = "dodge", bins = bin) +
+            labs(x = "Initial student instructor-student similarity rating", y = "Frequency", fill = "Group") +
+            scale_x_continuous(limits = c(.8,5.2)) +
+            scale_y_continuous(limits = c(0,155)) +
+            scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
+            theme_minimal()
+        
+        s1_tsrplot <- data %>% 
+            ggplot(aes(x = s1_tsr, fill = treatment)) +
+            geom_histogram(position = "dodge", bins = bin) +
+            labs(x = "Student anticipated instructor-student relationship rating", y = "Frequency") +
[…]

### 2021-04-26 outputs/paper/paper.Rmd
update repo organization, edit paper.

@@
 author: "Amy Farrow"
-date: "April 19th, 2021"
+date: "April 26th, 2021"
 header-includes:
@@
 library(DiagrammeR) # for causal diagram
-library(olsrr)
 library(tidymodels) # for train test
+library(performance) # for model performance evaluation
 
@@
 # This remakes the variable t_race, for consistency
+
 expdata <- expdata %>%
@@
   mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
-                         "Mixed race", 
+                         "Mixed race/ethnicity", 
                          ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
-                                ifelse(t1_race_1 == 1, "White",
-                                       ifelse(t1_race_2 == 1, "Black",
-                                              ifelse(t1_race_3 == 1, "Hispanic",
-                                                     ifelse(t1_race_4, "Asian",
-                                                            ifelse(t1_race_5, "American Indian",
+                                ifelse(t1_race_1 == 1, "White/Caucasian",
+                                       ifelse(t1_race_2 == 1, "Black or African American",
+                                              ifelse(t1_race_3 == 1, "Hispanic American or Latino/a",
+                                                     ifelse(t1_race_4, "Asian or Pacific Islander",
+                                                            ifelse(t1_race_5, "American Indian or Alaskan",
                                                                    ifelse(t1_race_6, "Middle Eastern",
-                                                                          "Other")))))))))
+                                                                          "Other Race")))))))))
 
@@
   mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
-                         "Mixed race", 
+                         "Mixed race/ethnicity", 
                          ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
-                                ifelse(s1_race_1 == 1, "White",
-                                       ifelse(s1_race_2 == 1, "Black",
-                                              ifelse(s1_race_3 == 1, "Hispanic",
-                                                     ifelse(s1_race_4, "Asian",
-                                                            ifelse(s1_race_5, "American Indian",
+                                ifelse(s1_race_1 == 1, "White/Caucasian",
+                                       ifelse(s1_race_2 == 1, "Black or African American",
+                                              ifelse(s1_race_3 == 1, "Hispanic American or Latino/a",
+                                                     ifelse(s1_race_4, "Asian or Pacific Islander",
+                                                            ifelse(s1_race_5, "American Indian or Alaskan",
                                                                    ifelse(s1_race_6, "Middle Eastern",
-                                                                          "Other"))))))))) %>%
+                                                                          "Other Race"))))))))) %>%
   mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))
@@
 `ltm` was used to check scale validity [@citeltm].
+`tidymodels` [@citetidymodels] and `performance` [@citeperformance] were used in assessing model validity and performance.
 Finally, the diagram was made using `DiagrammeR` [@citeDiagrammeR].
@@
 
-Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.
+Most key demographic traits come from the initial surveys given to students and instructors.
+For gender, they were asked to identify as male, female, or other, and given the option to enter text.
+For race/ethnicity, they were allowed to check any number of a range of options ('White/Caucasian', 'Black or African American', 'Hispanic American or Latino/a', 'Asian or Pacific Islander', 'American Indian or Alaskan', 'Middle Eastern', and 'Other Race').
+They were also asked if they were the first generation in their family to attend college (no or yes).
+Finally, they were asked to enter their age in years.
+The student's year of study and cumulative CGPA as of the previous term and the class size came from the college's internal records.
+
+Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples.
+There are no significant differences between the treatment and control groups, as indicated by p-values which are all above 0.4.
 
@@
   dplyr::select(treatment, 
-                s_female, 
+                s1_female, 
                 s_race, 
@@
   mutate(treatment = ff_label(treatment, "Group"),
-         s_female = ff_label(s_female, "Student gender"), 
+         s1_female = ff_label(s1_female, "Student gender"), 
          s1_age = ff_label(s1_age, "Student age"),
-         s_race = ff_label(s_race, "Student race"), 
+         s_race = ff_label(s_race, "Student race/ethnicity"), 
          ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
@@
   summary_factorlist("treatment", 
-                     c("s_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
+                     c("s1_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
                      p = TRUE, 
@@
          t_female = ff_label(t_female, "Instructor gender"),
-         t_race = ff_label(t_race, "Instructor race"),
+         t_race = ff_label(t_race, "Instructor race/ethnicity"),
          t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"),
-         n_course = ff_label(n_course, "Course size")) %>%
+         n_course = ff_label(n_course, "Class size")) %>%
   summary_factorlist("teacherid",
@@
 ```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8,4), fig.cap="Bar chart illustrating racial identification proprtions of participants"}
+
 instructorrace <- expdata %>%
@@
   unique() %>%
-  dplyr::summarize(White = sum(t1_race_1, na.rm = TRUE)/n(),
-            Black = sum(t1_race_2, na.rm = TRUE)/n(),
-            Hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
-            Asian = sum(t1_race_4, na.rm = TRUE)/n(),
-            "American Indian" = sum(t1_race_5, na.rm = TRUE)/n(),
-            "Middle Eastern" = sum(t1_race_6, na.rm = TRUE)/n(),
-            Other = sum(t1_race_7, na.rm = TRUE)/n(),
-            Multiple = sum(t_multi_race > 1, na.rm = TRUE)/n())
+  dplyr::summarize('White/Caucasian' = sum(t1_race_1, na.rm = TRUE)/n(),
+            'Black or African American' = sum(t1_race_2, na.rm = TRUE)/n(),
+            'Hispanic American or Latino/a' = sum(t1_race_3, na.rm = TRUE)/n(),
+            'Asian or Pacific Islander' = sum(t1_race_4, na.rm = TRUE)/n(),
+            'American Indian or Alaskan' = sum(t1_race_5, na.rm = TRUE)/n(),
+            'Middle Eastern' = sum(t1_race_6, na.rm = TRUE)/n(),
+            'Other Race' = sum(t1_race_7, na.rm = TRUE)/n(),
+            'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())
 studentrace <- expdata %>%
@@
   unique() %>%
-  dplyr::summarize(White = sum(s1_race_1, na.rm = TRUE)/n(),
-            Black = sum(s1_race_2, na.rm = TRUE)/n(),
-            Hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
-            Asian = sum(s1_race_4, na.rm = TRUE)/n(),
-            "American Indian" = sum(s1_race_5, na.rm = TRUE)/n(),
-            "Middle Eastern" = sum(s1_race_6, na.rm = TRUE)/n(),
-            Other = sum(s1_race_7, na.rm = TRUE)/n(),
-            Multiple = sum(s_multi_race > 1, na.rm = TRUE)/n())
+  dplyr::summarize('White/Caucasian' = sum(s1_race_1, na.rm = TRUE)/n(),
+            'Black or African American' = sum(s1_race_2, na.rm = TRUE)/n(),
+            'Hispanic American or Latino/a' = sum(s1_race_3, na.rm = TRUE)/n(),
+            'Asian or Pacific Islander' = sum(s1_race_4, na.rm = TRUE)/n(),
+            'American Indian or Alaskan' = sum(s1_race_5, na.rm = TRUE)/n(),
+            'Middle Eastern' = sum(s1_race_6, na.rm = TRUE)/n(),
+            'Other Race' = sum(s1_race_7, na.rm = TRUE)/n(),
+            'Multiple' = sum(s_multi_race > 1, na.rm = TRUE)/n())
 
@@
   mutate(nms...1 = fct_relevel(nms...1, 
-                               c("Hispanic", "White", "Asian",  "Middle Eastern",
-                                 "Black", "American Indian", "Other", "Multiple")),
+                               c('Hispanic American or Latino/a',
+                                 'White/Caucasian',
+                                 'Asian or Pacific Islander',
+                                 'Middle Eastern',
+                                 'Black or African American', 
+                                 'American Indian or Alaskan',
+                                 'Other Race',
+                                 'Multiple')),
          group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
@@
   geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
-  labs(x = "Self-identified race", y = "Proportion of sample", 
+  labs(x = "Self-identified race/ethnicity", y = "Proportion of sample", 
        fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
   scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
   theme_minimal()
@@
 
-```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(9,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
+```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
 n_instructor <- expdata %>%
@@
   draw_label(
-    "Some classes are very large (means marked)",
+    "Some classes are very large (class size means marked)",
     x = 0,
@@
           ncol = 1,
-          rel_heights = c(.1, 2))
+          rel_heights = c(.09, 2))
 ```
@@
 Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values.
-There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception.
-Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 and 3.5 versus 3.6).
+There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception).
+Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for intial perception and 3.5 versus 3.6 for end-of-term perception).
 
@@
 $X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study).
-ISR anticipation is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes.
+The original study used gender and CGPA are used as controls for grade based outcomes only.
 @citeoriginal state this is because females earn higher grades generally.
+This paper uses them as controls for the majority of outcomes.
+ISR anticipation is used as a control only for end-of-term student ISR perception.
 
 $X_{2j}$ is a vector of instructor-level covariates (class size).
+The original study consider instructor fixed effects instead.
+Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both.
+This paper selected class size because of the larger potential for transferable insights.
 
@@
 
-Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception (t2_sim1).
+Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.
+Because the instructors were only asked to answer one five-point scale question about similarity, this outcome cannot be treated the same way as the full-scale outcomes.
 
@@
 
-Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017 (f17_enrolled).
+Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017.
+Students are either not enrolled (0), enrolled (1), or they have graduated and are excluded from the model.
 
@@
 studentfirstsimilaritymodel <- lm(data = expdata,
-                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondsimilaritymodel <- lm(data = expdata,
-                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondrelationshipmodel <- lm(data = expdata,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 teachersecondrelationshipmodel <- lm(data = expdata,
-                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 grademodel <- lm(data = expdata,
-                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
-                                     na.action = na.omit)
-teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                 data = expdata,
@@
 enrollmentfallmodel <- glm(data = expdata,
-                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                            family = "binomial",
@@
 
-In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]).
-Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]).
-However, both coefficients are minimal on a scale of 1 to 5.
-
-Model 3 shows no significant relationship between treatment and student end-of-term ISR perception.
-However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]).
+In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]).
+Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller.
+Both coefficients are minimal on a scale of 1 to 5.
+The adjusted R-squared values, which indicate the proportion of variation in the outcome that can be explained by the treatment and covariates, are low (0.032 and 0.022).
+These models are helpful for analyzing treatment effect, but they do not aim to be and are not useful predictors of student similarity perception, not even within the sample.
+
+Model 3 shows no evidence that treatment impacted student end-of-term ISR perception.
+However, it does show a significant association between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]).
 For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671.
-
-Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-term ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent term.
-
-Models 1 through 8 were also repeated with additional student covariates (gender, race, first-generation status, CGPA, and year), and the results still showed a treatment effect on student similarity perception but no other effect.
-
-Model validation for the two primary significant models, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].
+Models 4, 5, 6, 7, and 8 show no significant treatment effect on end-of-term instructor ISR perception, course grade, objectively graded exam grade, end-of-term instructor similarity perception, or enrollment in the subsequent term.
+
+Model performance and validation for the two models with significant treatment effects, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].
 
@@
           notes.label = "",
+          font.size = "small",
           no.space = TRUE,
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
@@
                              "Persistence"),
-          covariate.labels = c("Treatment", "Student ISR 1", "Course size", "Student female", "CGPA", "Year of study"))
+          covariate.labels = c("Treatment", "Student ISR 1", "Class size", "Student female", "CGPA", "Year of study"))
 ```
@@
 Similarly, the subset of first-generation college student participants is considered.
-While the effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+The effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), but there is still no effect on ISR perception, instructor perception, or student outcomes.
 Full results are in [Appendix F].
@@
 studentfirstisrmodel1g <- lm(data = expdata,
-                                  formula = s1_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
+                                  formula = s1_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 studentsecondrelationshipmodel1g <- lm(data = expdata,
-                                     formula = s2_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
+                                     formula = s2_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 teachersecondrelationshipmodel1g <- lm(data = expdata,
-                                     formula = t2_tsr ~ s_firstgen + s_female + ir_f16_gpa + grade + year + n_course,
-                                     na.action = na.omit)
-```
-
-Table \@ref(tab:modeltablerel1) disregards the treatment condition, instead considering the relationship between student identity and similarity/ISR.
-While we cannot make causal inferences, we can explore relationships between the variables in this particular study.
+                                     formula = t2_tsr ~ s_firstgen + s1_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+```
+
+Table \@ref(tab:modeltablerel1) disregards the treatment condition, instead considering the association between student identity and similarity/ISR.
+While we cannot make causal inferences, this is an exploratory analysis that could shape future intervention studies.
 If a student identifies as first generation, can we anticipate a different student or instructor ISR perception?
-In this case, there is no significant relationship.
+In this case, there is no significant association.
 
@@
           notes.label = "",
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
           label = "tab:modeltablerel1",
+          font.size = "small",
           type = "latex",
@@
                              "Instructor ISR 2"),
-          covariate.labels = c("Student first gen.", "Student female", "CGPA", "Course grade", "Year of study", "Course size"))
+          covariate.labels = c("Student first gen.", "Student female", "CGPA", "Course grade", "Year of study", "Class size"))
 ```
@@
 studentfirstisrmodelhb <- lm(data = expdata,
-                                  formula = s1_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
+                                  formula = s1_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 studentsecondrelationshipmodelhb <- lm(data = expdata,
-                                     formula = s2_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
+                                     formula = s2_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
                                      na.action = na.omit)
 teachersecondrelationshipmodelhb <- lm(data = expdata,
-                                     formula = t2_tsr ~ s_hisp_black + s_female + ir_f16_gpa + grade + year + n_course,
-                                     na.action = na.omit)
-
-```
-
-Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative relationship between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.096, 95% CI [-0.166, -0.026]).
+                                     formula = t2_tsr ~ s_hisp_black + s1_female + ir_f16_gpa + grade + year + n_course,
+                                     na.action = na.omit)
+
+```
+
+Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative association between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.096, 95% CI [-0.166, -0.026]).
 
@@
           notes.label = "",
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
           label = "tab:modeltablerel2",
+          font.size = "small",
           type = "latex",
@@
                              "Instructor ISR 2"),
-          covariate.labels = c("Student Hispanic and/or Black", "Student female", "CGPA", "Course grade", "Year of study", "Course size"))
+          covariate.labels = c("Student Hispanic and/or Black", "Student female", "CGPA", "Course grade", "Year of study", "Class size"))
 ```
@@
 grademodel_ <- lm(data = expdata,
-                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
+                 formula = grade ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                      na.action = na.omit)
 examgrademodel_ <- lm(data = expdata %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
+                     formula = t2_finalexam ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                                      na.action = na.omit)
 enrollmentfallmodel_ <- glm(data = expdata,
-                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa + year,
+                           formula = f17_enrolled ~ treatment + s2_tsr + t2_tsr + s1_female + ir_f16_gpa,
                            family = "binomial",
@@
 
-Table \@ref(tab:noncausaltable) considers the relationship between similarity perception and ISR perception and the relationship between ISR perception and student outcomes.
+Table \@ref(tab:noncausaltable) considers the association between similarity perception and ISR perception and the relationship between ISR perception and student outcomes.
 Again, because only the treatment was randomized, we cannot make causal inferences.
 
-In Models 1 and 2, we can see that student and instructor similarity perception are both significantly related to student and instructor ISR perception.
-
-Unsurprisingly, student similarity perception relates more strongly to student ISR perception ($\beta$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception relates more strongly to instructor ISR perception ($\beta$ = 0.512, 95% CI [0.483, 0.541]).
-
-Models 3 and 4 show that student ISR perception is significantly related to course grade, while instructor ISR perception is significantly related to course grade and final grade.
-Instructor perception is a stronger predictor ($\beta$ = 0.339, 95% CI [0.294, 0.384] for course grade; $\beta$ = 0.643, 95% CI [0.527, 0.758] for final grade) than student perception ($\beta$ = 0.105, 95% CI [0.055, 0.155] for course grade).
-
-Model 5 indicates that ISR perception is not significantly related to persistence.
+In Models 1 and 2, we can see that student and instructor similarity perception are both significantly associated with student and instructor ISR perception.
+Unsurprisingly, student similarity perception is more strongly associated with student ISR perception ($\beta$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception is more strongly associated with instructor ISR perception ($\beta$ = 0.512, 95% CI [0.483, 0.541]).
+
+Models 3 and 4 show that student ISR perception is significantly associated course grade, while instructor ISR perception is significantly associated with course grade and final grade.
+Instructor perception is more strongly associated ($\beta$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\beta$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\beta$ = 0.115, 95% CI [0.064, 0.164] for course grade).
+
+Model 5 indicates that ISR perception is not significantly associated with persistence.
 
@@
           no.space = TRUE,
+          font.size = "small",
           notes.label = "",
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
@@
                              "Instructor sim. 2",
-                             "Course size",
+                             "Class size",
                              "Student ISR 2",
@@
                              "Student female",
-                             "CGPA",
-                             "Year"))
+                             "CGPA"))
 ```
@@
                              0)) %>%
-  mutate(gend_match = ifelse(t_female == s_female,
+  mutate(gend_match = ifelse(t_female == s1_female,
                              1,
@@
 Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
-As with the replication results, there is a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.108, 95% CI [0.036, 0.179]).
+As with the replication results, there is a measurable treatment effect for initial student similarity perception ($\beta$ = 0.108, 95% CI [0.036, 0.179]).
 However, the coefficient is even lower than in the replication models.
-There is still no significant relationship between treatment and student end-of-term ISR perception.
-Additionally, there is no significant relationship between treatment and standardized course grade.
+There is still no treatment effect for student end-of-term ISR perception or standardized course grade.
 
@@
           initial.zero = FALSE,
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          font.size = "small",
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
@@
                                "CGPA", 
-                               "Course size",
+                               "Class size",
                                "Matching racial self-ID", "Matching gender self-ID", "Age difference", "Number of similarities")
@@
 Do some similarities matter more than others in ISRs?
-If demographics traits, which are already part of an inequitable campus environment, are important for perceived similarity, Should positive ISRs be contingent on similarity at all?
+As indicated by the results in Table \@ref(tab:modeltablerel2, some marginalized student identities may be associated with measurably lower ISR perceptions.
+If demographics traits, which are already part of an inequitable campus environment, are important for perceived similarity, should positive ISRs be contingent on similarity at all?
 There are other ways to cultivate ISRs that may make more sense and offer more equitable outcomes.
@@
 
+Any measurable effects could also be specific to this particular college campus. Interventions may not be transferable due to different demographics, cultures, and institutional structures.
+
 As with all survey studies, there is a risk that self-reported information is inaccurate.
@@
 Survey fatigue may have reduced the meaningfulness of the survey answers, especially those later in the survey.
+While responses that were given overly quickly were removed from the analysis, the fact that 428 cases had to be removed for this reason indicates a common problem with survey data: we cannot know if the participants actually considered their answers.
 
@@
 Similarity is only one of many theorized methods for improved ISRs.
-These would be entirely different lines of inquiry, however.
 
@@
   mutate(treatment = ff_label(treatment, "Group"),
-         t_female = ff_label(t_female, "Teacher gender"), 
-         t_age = ff_label(t_age, "Teacher age"),
-         t_race = ff_label(t_race, "Teacher race"), 
-         t_firstgen = ff_label(t_firstgen, "Teacher first-gen status"), 
-         n_course = ff_label(n_course, "Course size")) %>%
+         t_female = ff_label(t_female, "Instructor gender"), 
+         t_age = ff_label(t_age, "Instructor age"),
+         t_race = ff_label(t_race, "Instructor race/ethnicity"), 
+         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"), 
+         n_course = ff_label(n_course, "Class size")) %>%
   summary_factorlist("treatment", 
@@
 
+Linear models are based on a number of assumptions: a linear relationship, homoscedasticity (homogeneity of variance), a lack of correlation between the independent variables, and normally-distributed residuals.
+Figure  \@ref(fig:model1performance) summarizes tests of these assumptions for the initial student similarity perception model.
+All assumptions seem to hold for this model.
+Figure  \@ref(fig:model2performance) shows the same tests for the end-of-term student similarity perception model.
+
+```{r model1performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="Initial student similarity perception model performance"}
+performance::check_model(studentfirstsimilaritymodel)
+```
+
+```{r model2performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="End-of-term student similarity perception model performance"}
+performance::check_model(studentsecondsimilaritymodel)
+```
+
 ```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
@@
 studentfirstsimilaritymodelv <- lm(data = expdata_train,
-                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondsimilaritymodelv <- lm(data = expdata_train,
-                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
@@
 Table \@ref(tab:modelvtable) shows the two primary significant models, redone using a random train/test split (80% training, 20% testing).
+The treatment effect coefficient for initial student similarity perception is close to the coefficient in the first version of the model ($\beta$ = .202, 95% CI [.137, .267]), but the coefficient for end-of-term student similarity perception is more different ($\beta$ = .137, 95% CI [.055,.219]).
 Table \@ref(tab:traintesttable) shows metrics to assess the performance of the two models on the training and testing datasets.
+The root mean squared errors and mean absolute errors are slightly larger for the testing data, which makes sense, as the model was fit to the training data but not the testing data.
+From these numbers, we can conclude that minimal overfitting in the model has occurred.
 
@@
           no.space = TRUE,
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-10pt",
           label = "tab:modelvtable",
+          font.size = "small",
           notes.label = "",
@@
           initial.zero = FALSE,
-          keep = c("treatment", "n_course", "s_female", "ir_f16_gpa", "year", "Constant"),
+          keep = c("treatment", "n_course", "s1_female", "ir_f16_gpa", "year", "Constant"),
           #omit = c("teacherid"),
@@
                              "Student sim. 2"),
-          covariate.labels = c("Treatment", "Course size", "Student female", "CGPA", "Year of study"))
+          covariate.labels = c("Treatment", "Class size", "Student female", "CGPA", "Year of study"))
 ```
@@
               bind_cols(metrics(expdata_test, s2_sim, pred2) %>% dplyr::select(.estimate))) %>%
+  mutate(.metric = c("Root mean square error", "R-squared", "Mean absolute error",
+                     "Root mean square error", "R-squared", "Mean absolute error")) %>%
 kable(caption = "Metrics for student similarity perception models",
@@
 studentfirstsimilaritymodelhb <- lm(data = expdatahb,
-                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondsimilaritymodelhb <- lm(data = expdatahb,
-                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 teachersecondrelationshipmodelhb <- lm(data = expdatahb,
-                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 grademodelhb <- lm(data = expdatahb,
-                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 examgrademodelhb <- lm(data = expdatahb %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
-                                     na.action = na.omit)
-teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarityhb <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                 data = expdatahb,
@@
 enrollmentfallmodelhb <- glm(data = expdatahb,
-                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                            family = "binomial",
@@
           notes.label = "",
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-12pt",
           label = "tab:modeltablehb",
-          font.size = "footnotesize",
+          font.size = "small",
           type = "latex",
@@
 studentfirstsimilaritymodel1g <- lm(data = expdata1g,
-                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondsimilaritymodel1g <- lm(data = expdata1g,
-                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 studentsecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
+                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 teachersecondrelationshipmodel1g <- lm(data = expdata1g,
-                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 grademodel1g <- lm(data = expdata1g,
-                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                      na.action = na.omit)
 examgrademodel1g <- lm(data = expdata1g %>% filter(obj_exam == 1),
-                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
-                                     na.action = na.omit)
-teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
+                                     na.action = na.omit)
+teachersecondsimilarity1g <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                 data = expdata1g,
@@
 enrollmentfallmodel1g <- glm(data = expdata1g,
-                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
+                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                            family = "binomial",
@@
           no.space = TRUE,
-          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
+          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
           column.sep.width = "-12pt",
-          font.size = "footnotesize",
+          font.size = "small",
           label = "tab:modeltable1g",

### 2021-04-27 outputs/paper/paper.Rmd
final edits? i hope.

@@
 In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories.
-More students identified as Hispanic, while more instructors identified as White and selected multiple options.
+While the tables display race/ethnicity as a single variable (it is either a single identity, mixed race/ethnicity, or unknown), this graph, and the subsequent analysis, deal with multiple racial/ethnic identities that are either selected or not.
+On the graph, we can see that more students identified as Hispanic, while more instructors identified as White and selected multiple options.
+Viewing only the table, where race is shown as a one-dimensional variable, does not show the nuance of multiple racial/ethnic identities.
+Still, these categories are limiting, and do not necessarily truly reflect the social or cultural identities of study participants.
 
@@
 Corresponding to Table \@ref(tab:summarytable), the only visible differences in distributions are in the initial student similarity perception and end-of-term student similarity perception.
+
+Interestingly, there is a large difference between how students rate perceived similarity and how instructors rate perceived similarity.
+Students, on average, rate end of term similarity at 3.6, while instructors give an average rating of 2.7.
+Students give an average end-of-term ISR rating of 4.1, while instructors give an average of only 3.5.
+This could be due to more positive student perceptions, or due to students giving higher ratings for equivalent perceptions.
+Both groups rate ISR more highly than similarity, but it is difficult to compare the weight of these scores as the scales are not necessarily measuring in the same units.
 
@@
 $X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study).
+Different covariates are used to model different outcomes.
 The original study used gender and CGPA are used as controls for grade based outcomes only.
@@
 $X_{2j}$ is a vector of instructor-level covariates (class size).
+Different covariates are used to model different outcomes.
 The original study consider instructor fixed effects instead.
@@
 
-$\beta_{0}$, $\beta_{1}$, $\Gamma_{1}$, $\Gamma_{2}$, and $a_{k}$ are coefficients on the resulting models.
+In the resulting models, $\beta_{0}$ and $a_{k}$ are constant terms. $\beta_{1}$ is the coefficient that indicates treatment effect. $\Gamma_{1}$ and $\Gamma_{2}$ are coefficients for the covariates.
 
@@
 (\#eq:linmodel)
-Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij}
+Outcome_{ij} = \beta_{0} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij}
 \end{equation}
@@
 (\#eq:ordlogmodel)
-prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij} > k
+prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij} > k
 \end{equation}
@@
 (\#eq:logmodel)
-prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + X_{1i}\Gamma_{1} + X_{2j}\Gamma_{2} + \epsilon_{ij} > k
+prob(outcome_{ij}) = a_{k} + \beta_{1}treatment_{i} + \Gamma_{1}X_{1i} + \Gamma_{2}X_{2j} + \epsilon_{ij} > k
 \end{equation}
@@
 
+All results are reported using the treatment coefficient, $\beta_{1}$, and covariate coefficients, which are represented with $\Gamma$.
+95% confidence intervals for those coefficients are shown in brackets below the coefficients.
+Significance levels are indicated with stars.
+Initial perceptions are indicated with 1, and end-of-term perceptions with 2.
+ISR refers to instructor-student relationship perception, and sim.
+refers to similarity perception.
+Final grade refers to the grade given for an objectively graded final exam in the course, and Std.
+course grade refers to a standardized final course grade.
+
 ## Replication results
@@
 
-In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]).
-Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller.
+In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta_{1}$ = 0.194, 95% CI [0.136, 0.253]).
+Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller.
 Both coefficients are minimal on a scale of 1 to 5.
@@
 Model 3 shows no evidence that treatment impacted student end-of-term ISR perception.
-However, it does show a significant association between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]).
+However, it does show a significant association between student ISR anticipation and student end-of-term ISR perception ($\Gamma$ = 0.671, 95% CI [0.624, 0.717]).
 For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671.
 Models 4, 5, 6, 7, and 8 show no significant treatment effect on end-of-term instructor ISR perception, course grade, objectively graded exam grade, end-of-term instructor similarity perception, or enrollment in the subsequent term.
+
+All models in Table \@ref(tab:modeltable) use gender, prior GPA, and year of study as student-level covariates.
+Instructor-level covariates are limited to the class size.
+This is based on the assumption that student traits will be more likely to affect student experiences, and therefore student outcomes.
+Race/ethnicity is not included because that covariate is further explored in separate models.
 
@@
 Because some students are more commonly targeted by persistence interventions, the subset of Hispanic and/or Black college student participants is considered separately from the rest of the student population.
-While the effect of treatment on student similarity perception is larger than for the general student population ($\beta$ = 0.199, 95% CI [0.124, 0.274] for initial perception; $\beta$ = 0.116, 95% CI [0.021, 0.211] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
+While the effect of treatment on student similarity perception is larger than for the general student population ($\beta_{1}$ = 0.199, 95% CI [0.124, 0.274] for initial perception; $\beta_{1}$ = 0.116, 95% CI [0.021, 0.211] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes.
 Full results are in [Appendix E].
@@
 Similarly, the subset of first-generation college student participants is considered.
-The effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), but there is still no effect on ISR perception, instructor perception, or student outcomes.
+The effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta_{1}$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta_{1}$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), but there is still no effect on ISR perception, instructor perception, or student outcomes.
 Full results are in [Appendix F].
@@
 
-Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative association between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\beta$ = -0.096, 95% CI [-0.166, -0.026]).
+Table \@ref(tab:modeltablerel2), however, shows that there is a significant negative association between student identification as Hispanic and/or Black and end-of-term instructor ISR perception ($\Gamma$ = -0.096, 95% CI [-0.166, -0.026]).
 
@@
 Again, because only the treatment was randomized, we cannot make causal inferences.
+Course size is used as a covariate for ISR perception because it directly shapes how students and instructors relate to one another.
+Student gender and prior GPA are used as covariates for student outcomes for the reasons previously outlined from @citeoriginal.
 
 In Models 1 and 2, we can see that student and instructor similarity perception are both significantly associated with student and instructor ISR perception.
-Unsurprisingly, student similarity perception is more strongly associated with student ISR perception ($\beta$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception is more strongly associated with instructor ISR perception ($\beta$ = 0.512, 95% CI [0.483, 0.541]).
+Unsurprisingly, student similarity perception is more strongly associated with student ISR perception ($\Gamma$ = 0.733, 95% CI [0.706, 0.76]), and instructor similarity perception is more strongly associated with instructor ISR perception ($\Gamma$ = 0.512, 95% CI [0.483, 0.541]).
 
 Models 3 and 4 show that student ISR perception is significantly associated course grade, while instructor ISR perception is significantly associated with course grade and final grade.
-Instructor perception is more strongly associated ($\beta$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\beta$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\beta$ = 0.115, 95% CI [0.064, 0.164] for course grade).
+Instructor perception is more strongly associated ($\Gamma$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\Gamma$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\Gamma$ = 0.115, 95% CI [0.064, 0.164] for course grade).
 
@@
 Table \@ref(tab:mymodeltable) shows the results of models that consider matching student-instructor traits.
-As with the replication results, there is a measurable treatment effect for initial student similarity perception ($\beta$ = 0.108, 95% CI [0.036, 0.179]).
+As with the replication results, there is a measurable treatment effect for initial student similarity perception ($\beta_{1}$ = 0.108, 95% CI [0.036, 0.179]).
 However, the coefficient is even lower than in the replication models.
 There is still no treatment effect for student end-of-term ISR perception or standardized course grade.
+Because the matching ID covariates were used, student covariates like gender and race were not included.
+Prior GPA and class size were still included, because they do not have a "matching" trait.
 
@@
 Table \@ref(tab:modelvtable) shows the two primary significant models, redone using a random train/test split (80% training, 20% testing).
-The treatment effect coefficient for initial student similarity perception is close to the coefficient in the first version of the model ($\beta$ = .202, 95% CI [.137, .267]), but the coefficient for end-of-term student similarity perception is more different ($\beta$ = .137, 95% CI [.055,.219]).
+The treatment effect coefficient for initial student similarity perception is close to the coefficient in the first version of the model ($\beta_{1}$ = .202, 95% CI [.137, .267]), but the coefficient for end-of-term student similarity perception differs more ($\beta_{1}$ = .137, 95% CI [.055,.219]).
 Table \@ref(tab:traintesttable) shows metrics to assess the performance of the two models on the training and testing datasets.

### 2021-04-27 scripts/02_make_shiny_data.R

@@ -1,0 +1,112 @@
+### Preamble ###
+# Purpose: Process and save data for use with shiny app
+# Author: Amy Farrow
+# Contact: amy.farrow@mail.utoronto.ca
+# Date: 2021-04-25
+# Pre-requisites: 01_import_data.R
+# to-dos:
+
+expdata <- isrdata %>%
+  filter(t1_consent == 1 # Only include if instructor consented to first survey
+         & t1_complete == 1 # Only included if instructor completed first survey
+         & t1_timer_consent_3 > 1 # Only include if instructor read consent
+         & s1_consent == 1 # Only include if student consented to first survey
+         & s1_timer_consent_3 > 1 # Only include if student read consent
+         & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
+         & t2_complete == 1 # Only include if the instructor completed the whole second survey
+         # The rest of the lines make sure that the survey taker looked at each page for more than three seconds
+         & t1_timer_gtky1_3 > 10
+         & t1_timer_gtky2_3 > 10
+         & t1_timer_gtky3_3 > 10
+         & t1_timer_demos1_3 > 10
+         & t1_timer_demos2_3 > 10
+         & s1_timer_gtky1_3 > 10
+         & s1_timer_gtky2_3 > 10
+         & s1_timer_gtky3_3 > 10
+         & s1_timerpre1_3 > 10
+         & s1_timerpre2_3 > 10
+         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
+         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
+         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
+         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
+         & s1_timerpost1_3 > 10
+         & s1_timerpost2_3 > 10
+         & s1_timerdemo1_3 > 10
+         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
+         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
+         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
+         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
[…]

### 2021-04-27 outputs/paper/paper.Rmd
actual last edits.

@@
          & t2_complete == 1 # Only include if the instructor completed the whole second survey
-         # The rest of the lines make sure that the survey taker looked at each page for more than three seconds
+         # Make sure that the survey taker looked at each page for more than three seconds
          & t1_timer_gtky1_3 > 10
@@
          & t2_timer > 10
+         # Remove specific drop reasons
          & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
@@
          )
-
-# Alternatively: expdata <- isrdata %>% filter(use == 1) %>% # Select only the cases the original study used.
 
@@
 
+# This creates a variable that is the final grade if the final was an objectively graded exam
+# and otherwise NA
 expdata <- expdata %>%
@@
 ```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# Conduct tests for scale reliability
 s1_sim_scale <- expdata %>%
@@
 The dataset contains 36,838 observations of 653 variables.
+While there are many additional variables, including a wider range of survey questions targeting interest level, enjoyment, and confidence, this paper focuses on the covariates and key outcomes that are discussed in the original paper.
 
@@
 
+# Summarise instructor race/ethnicity data
 instructorrace <- expdata %>%
@@
             'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())
+
+# summarise student race/ethnicity data
 studentrace <- expdata %>%
@@
 
+# bind the two summaries and present them in a bar chart
 bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
@@
 ```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
+
+# Summarise course sizes from instructor POV
 n_instructor <- expdata %>%
@@
 
+# Plot instructor POV course sizes
 n_by_instructor <- n_instructor %>%
@@
 
+# Summarise course sizes from a student POV
 n_student <- expdata %>%
@@
 
+# Plot student POV summary
 n_by_student <- n_student %>%
@@
 
+# Combine the plots in a grid
 combined <- plot_grid(n_by_instructor, n_by_student,
           ncol = 2)
-
 title <- ggdraw() + 
@@
   )
-
 plot_grid(title,
@@
           ncol = 1,
-          rel_heights = c(.09, 2))
+          rel_heights = c(.1, 2))
 ```
@@
 
-Interestingly, there is a large difference between how students rate perceived similarity and how instructors rate perceived similarity.
-Students, on average, rate end of term similarity at 3.6, while instructors give an average rating of 2.7.
+Interestingly, there is a large difference between how students and instructors rate perceived similarity and ISR.
+Students, on average, rate end-of-term similarity at 3.6, while instructors give an average rating of 2.7.
 Students give an average end-of-term ISR rating of 4.1, while instructors give an average of only 3.5.
@@
 ```{r continuousoutcomes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height= 10, fig.cap="Histograms illustrating continous and ordinal key outcomes"}
+# Plot initial student similarity perception
 s1_simplot <- expdata %>% 
@@
 
+# Plot initial student ISR perception
 s1_tsrplot <- expdata %>% 
@@
 
+# Plot end-of-term student similarity perception
 s2_simplot <- expdata %>% 
@@
 
+# PLot end of term student ISR perception
 s2_tsrplot <- expdata %>% 
@@
 
+# Plot end of term instructor similarity perception
 t2_sim1plot <- expdata %>% 
@@
 
+# Plot end of term instructor ISR perception
 t2_tsrplot <- expdata %>% 
@@
 
+# Plot grade
 gradeplot <- expdata %>% 
@@
 
+# plot standardized grades
 std_gradeplot <- expdata %>% 
@@
 
+# Plot final exam grades
 t2_finalexamplot <- expdata %>% 
@@
 
+# Combine plots in a grid
 combined <- plot_grid(s1_simplot + theme(legend.position = "none"), s1_tsrplot + theme(legend.position = "none"),
@@
           ncol = 2)
-
 legend <- get_legend(s1_simplot + 
@@
                        theme(legend.position = "bottom"))
-
 title <- ggdraw() + 
@@
 From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis.
-Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend at more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428).
+Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428).
 The time limits are somewhat arbitrary, and different choices could have different analysis results.
@@
 Different covariates are used to model different outcomes.
-The original study used gender and CGPA are used as controls for grade based outcomes only.
+The original study used gender and CGPA as controls for grade based outcomes only.
 @citeoriginal state this is because females earn higher grades generally.
 This paper uses them as controls for the majority of outcomes.
+This is because of the potential link between student performance and perceptions of similarity and ISR.
 ISR anticipation is used as a control only for end-of-term student ISR perception.
@@
 $X_{2j}$ is a vector of instructor-level covariates (class size).
-Different covariates are used to model different outcomes.
-The original study consider instructor fixed effects instead.
+The original study considered instructor fixed effects instead.
 Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both.
 This paper selected class size because of the larger potential for transferable insights.
+These particular instructors are not of general interest, but the effect of class size may be.
 
 $\epsilon_{ij}$ is a standard error.
-As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard error and p-values when compared with the original paper.
+As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard errors and p-values when compared with the original paper.
 This does not affect the regression coefficients.
@@
 Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes.
-This include complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception.
-Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, they can be treated as continuous variables and modeled using linear functions.
+This includes complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception.
+Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, and they can be treated as continuous variables and modeled using linear functions.
 It also includes grade-based outcomes: course grade and objectively graded exam grade, both of which are shown on a 4.0 GPA scale.
@@
 
-Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.
-Because the instructors were only asked to answer one five-point scale question about similarity, this outcome cannot be treated the same way as the full-scale outcomes.
+Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale.
+Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced.
+Because this variable is measured differently than the two student similarity perception scales, and therefore modeled differently, direct comparisons between student and instructor perception of similarity should be made with caution.
 
@@
 When creating additional models, I focused on three key outcomes: standardized course grade, end-of-term student perceived ISR scale, and end-of-term student perceived similarity scale.
-I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes than outcomes from subsequent terms; and because previous studies show more detectable impact on student perception than instructor perception.
+I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes as opposed to outcomes in subsequent terms; and because previous studies focus more on student perception than instructor perception.
 Equation \@ref(eq:mylinmodel) shows the linear model used.
@@
 In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta_{1}$ = 0.194, 95% CI [0.136, 0.253]).
-Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller.
+Similarly, Model 2 shows that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller.
 Both coefficients are minimal on a scale of 1 to 5.
@@
 
-Models 3 and 4 show that student ISR perception is significantly associated course grade, while instructor ISR perception is significantly associated with course grade and final grade.
+Models 3 and 4 show that student ISR perception is significantly associated with course grade, while instructor ISR perception is significantly associated with course grade and final grade.
 Instructor perception is more strongly associated ($\Gamma$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\Gamma$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\Gamma$ = 0.115, 95% CI [0.064, 0.164] for course grade).
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# create variables to indicate matching student-instructor traits
 expdata <- expdata %>%
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Make all models for matching traits
 mystudentsecondsimilaritymodel <- lm(data = expdata,
@@
 ```{r causaldiagram, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Diagram of desired causal sequence", dev='png', out.width="100%"}
+# Make a diagram to show the causal sequence that is desired in this intervention
 DiagrammeR::grViz("digraph {
@@
 It is also possible that similarities declared by an anonymous third party are simply not as impactful as similarities that are personally observed and experienced.
+Finally, because the instructor similarity perception was measured on one five-point scale, rather than a full scale as student similarity perception was, perhaps the measurement was less sensitive to change.
+@citeoriginal note that they used a single question to reduce the burden placed on instructors, who had to answer surveys for many students.
+However, this courtesy created a dissimilarity in measurment for outcomes that we would like to compare.
 
@@
 
-The exploratory modeling in Table \@ref(tab:noncausaltable) also indicates that ISR, or at least ISR perception, is related to student outcomes, as in relationships 5 and 6 in Figure \@ref(fig:causaldiagram).
+The exploratory modeling in Table \@ref(tab:noncausaltable) also indicates that ISR, or at least ISR perception, is related to student grade outcomes, as in relationship 5 in Figure \@ref(fig:causaldiagram).
 Again, we cannot conclude anything about causality.
@@
 Because the quality of the ISR cannot be measured directly, it is also possible that instructor-student relationships can improve student outcomes, but the particular perceptions that are measured in this case are not the relevant factors, or perceived ISR is less important that the real behaviors within an ISR.
-
-The environment of @citeoriginal's experiment is also relevant.
+Even in this exploratory analysis, ISR perception was not associated with persistence, and relationship 6 did not appear.
+
+The environment of @citeoriginal's experiment is relevant.
 They aimed to build ISRs in a context where a student's average class size is 49.5, students typically take five classes at a time, and instructors teach multiple classes, sometimes at multiple campuses.
@@
 Do some similarities matter more than others in ISRs?
-As indicated by the results in Table \@ref(tab:modeltablerel2, some marginalized student identities may be associated with measurably lower ISR perceptions.
+As indicated by the results in Table \@ref(tab:modeltablerel2), some marginalized student identities may be associated with measurably lower instructor ISR perceptions.
 If demographics traits, which are already part of an inequitable campus environment, are important for perceived similarity, should positive ISRs be contingent on similarity at all?
@@
 The external validity is more questionable.
-The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate and followed through.
+The study participants were chosen in two main ways: the instructor was interested in the study and completed all the steps successfully, and the student agreed to participate.
+Data is only available for the second student survey for students who chose to follow through.
 The sample of instructors that would be interested in the study would not necessarily be comparable to the overall population: they could be instructors who care more about supporting the work of other academics, instructors for whom a $150 giftcard was a significant draw, or instructors who are particularly interested in the role of ISRs.
+Because the instructors administered the surveys to students, students who actively engage with their instructors' requests could be overrepresented in the student sample.
 
@@
 While responses that were given overly quickly were removed from the analysis, the fact that 428 cases had to be removed for this reason indicates a common problem with survey data: we cannot know if the participants actually considered their answers.
+Additionally, survey-takers may feel pressure to give the answers they believe they should give, rather than answers that are more honest.
+This could be especially true in this context, where students are usually being graded on their performance and answers to multiple-choice questions.
 
@@
 
-Given the lack of significant findings, future studies could take a wide variety of approaches.
+The data collected by @citeoriginal is wider than is shown in this paper or the original.
+As mentioned previously, there were other survey questions targeting interest level, enjoyment, and confidence, amoungst other additional information.
+This data could be used in further exploratory analysis to look at associations between similarity, ISR, interest, enjoyment, and confidence.
+
+Given the minimal treatment effects found in this particular intervention, future studies could take a wide variety of approaches.
 First, more intensive, but similar, interventions to increase perceived similarity could be tested.
@@
 ```{r missingdata, echo=FALSE, warning=FALSE, message=FALSE, fig.height=9, fig.width=9}
+# Plot missing data patterns for key covariates
 explanatory = c("s2_tsr", "s1_tsr", "ir_f16_gpa", "n_course", "s1_female")
@@
 
-```{r model1performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="Initial student similarity perception model performance"}
+```{r model1performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="Initial student similarity perception model assumptions"}
+# Plot Student sim. 1 model assumptions
 performance::check_model(studentfirstsimilaritymodel)
@@
 
-```{r model2performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="End-of-term student similarity perception model performance"}
+```{r model2performance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height= 10, fig.cap="End-of-term student similarity perception model assumptions"}
+# Plot student sim. 2 model assumptions
 performance::check_model(studentsecondsimilaritymodel)
@@
 ```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+# Split train and test data
 set.seed(1847)
@@
 
-```{r modelvtable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
+```{r modelvtable, results=c('asis', 'hold'), echo=FALSE, warning=FALSE, message=FALSE}
 # Make table to display model results
@@
 ```{r traintesttable, echo=FALSE, warning=FALSE, message=FALSE}
+# Use the model to predict training data outcomes
 expdata_train <- expdata_train %>%
@@
 
+# Use the model to predict testing data outcomes
 expdata_test <- expdata_test %>%
@@
 
+# Calculate metrics to assess model performance
 metrics(expdata_train, s1_sim, pred1) %>%
@@
                      "Root mean square error", "R-squared", "Mean absolute error")) %>%
+# put in a table
 kable(caption = "Metrics for student similarity perception models",
@@
   pack_rows("End-of-term", 4, 6) %>%
-  kable_styling()
+  kable_styling(latex_options = "hold_position")
 ```
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Make all replication models
-
+# Filter to data that only includes students who identify as hispanic or black
 expdatahb <- expdata %>%
@@
 
+# Make all replication models
 studentfirstsimilaritymodelhb <- lm(data = expdatahb,
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Make all replication models
-
+# Filter so we only have data for first-generation college students
 expdata1g <- expdata %>%
@@
 
+# Make all replication models
 studentfirstsimilaritymodel1g <- lm(data = expdata1g,
@@
 ```{r similarities, echo=FALSE, warning=FALSE, message=FALSE}
+# Collect all similarities that were presented to students and instructors
 tibble(Similarities = unique(expdata$sentence1)) %>%
@@
   filter(Similarities != " ") %>%
+# put in a table
   knitr::kable(caption = "Similarities presented to students and instructors during treatment",

### 2021-04-27 scripts/02_make_shiny_data.R

@@ -6,6 +6,9 @@
 # Pre-requisites: 01_import_data.R
 # to-dos:
 
+# this script replicates the data selection and cleaning that is done in the paper.Rmd file.
+# the data is then filtered to include only what is necessary for the shiny app and saved.
+
 expdata <- isrdata %>%
   filter(t1_consent == 1 # Only include if instructor consented to first survey
          & t1_complete == 1 # Only included if instructor completed first survey
@@ -14,7 +17,7 @@ expdata <- isrdata %>%
          & s1_timer_consent_3 > 1 # Only include if student read consent
          & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
          & t2_complete == 1 # Only include if the instructor completed the whole second survey
-         # The rest of the lines make sure that the survey taker looked at each page for more than three seconds
+         # Make sure that the survey taker looked at each page for more than three seconds
          & t1_timer_gtky1_3 > 10
          & t1_timer_gtky2_3 > 10
          & t1_timer_gtky3_3 > 10
@@ -40,6 +43,7 @@ expdata <- isrdata %>%
          & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
          & (s2_timer4_3 > 10 | is.na(s2_timer4_3))
          & t2_timer > 10
+         # Remove specific drop reasons
          & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
          & t_drop_reasons != "Taught online course"
          & t_drop_reasons != "Administered to only graduate students"
@@ -52,6 +56,10 @@ expdata <- expdata %>%
          f17_enrolled = as.factor(f17_enrolled),
          s1_female = as.factor(s1_female),
          t_female = as.factor(t_female))
+
+# The original coding of race contained errors -- ex. a professor who selected White and Asian in the survey,
+# and was marked as Asian for most cases, but white for one random case.
+# This remakes the variable t_race, for consistency
 
 expdata <- expdata %>%
   mutate(t1_race_1 = replace_na(as.numeric(t1_race_1), 0),
@@ -92,9 +100,12 @@ expdata <- expdata %>%
                                                                           "Other Race"))))))))) %>%
   mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))
 
+# This creates a variable that is the final grade if the final was an objectively graded exam
+# and otherwise NA
[…]

### 2021-04-27 outputs/paper/paper.Rmd
why didn't it push the correctly-knitted version before

@@
 In this case, the scales should measure similarity perception and instructor-student relationship (ISR) perception.
-@citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eight-grade American students and their teachers.
+@citeoriginal used modified versions of the scales from @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015, where they were used with ninth-grade American students and their teachers; the scales were originally were presented in @Gehlbach_Brinkworth_Harris_2012, and they were developed for use with sixth-, seventh-, and eighth-grade American students and their teachers.
 Given the fact that these scales were developed for children, it is worth considering what the scales actually measure.
@@
 There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception).
-Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for intial perception and 3.5 versus 3.6 for end-of-term perception).
+Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for initial perception and 3.5 versus 3.6 for end-of-term perception).
 
@@
 
-Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale.
+Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.
+The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale.
 Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced.
@@
 @citeoriginal note that they used a single question to reduce the burden placed on instructors, who had to answer surveys for many students.
-However, this courtesy created a dissimilarity in measurment for outcomes that we would like to compare.
+However, this courtesy created a dissimilarity in measurement for outcomes that we would like to compare.
 
